### 2017-01-30 report/paper.tex
started writing the paper

\documentclass[conference,10pt]{IEEEtran}
%\documentclass[conference,draft,onecolumn]{IEEEtran}
% useful packages, copy and paste from diff sources

\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{cite,url,color} % Citation numbers being automatically sorted and properly "compressed/ranged".
\usepackage{graphics,amsfonts}
\usepackage{epstopdf}
\usepackage[pdftex]{graphicx}
\usepackage[cmex10]{amsmath}
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally does.
\usepackage[utf8]{inputenc}
% Useful for displaying quotations
%\usepackage{csquotes}
% Compact lists
%\let\labelindent\relax
\usepackage{enumitem}

%tikz figures
\usepackage{tikz}
\usetikzlibrary{automata,positioning,chains,shapes,arrows}
\usepackage{pgfplots}
\usetikzlibrary{plotmarks}
\newlength\fheight
\newlength\fwidth
\pgfplotsset{compat=newest}
\pgfplotsset{plot coordinates/math parser=false}

\usepackage{array}
% http://www.ctan.org/tex-archive/macros/latex/required/tools/
%\usepackage{mdwmath}
%\usepackage{mdwtab}
%mdwtab.sty	-- A complete ground-up rewrite of LaTeX's `tabular' and  `array' environments.  Has lots of advantages over
%		   the standard version, and over the version in `array.sty'.
% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
\usepackage{subfig}

\usepackage[top=1.5cm, bottom=2cm, right=1.6cm,left=1.6cm]{geometry}
\usepackage{indentfirst}

\usepackage{times}
% make sections titles smaller to save space
%\usepackage{sectsty}
%\sectionfont{\large}
% enable the use of 'compactitem', a smaller 'itemize'
%\usepackage{paralist}

% MP
% to split equations using dmath env
\usepackage{breqn}
% nice rules in tables
\usepackage{booktabs}

%\setlength\parindent{0pt}
\linespread{1}

% MC
\newcommand{\MC}[1]{\textit{\color{red}MC says: #1}}
\newcommand{\AZ}[1]{\textit{\color{blue}AZ says: #1}}
\newcommand{\MP}[1]{\textit{\color{green}MP says: #1}}

\usepackage{placeins}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{LaTeX Template for a Scientific Paper}

\author{\IEEEauthorblockN{Andrea Pittaro, Matteo Maso }
\IEEEauthorblockA{Department of Information Engineering, University of Padova -- Via Gradenigo, 6/b, 35131 Padova, Italy\\Email: {\tt\{pittaroa,masomatt\}@dei.unipd.it}
}}

\maketitle

\begin{abstract}
In this paper we present an application of the Software Defined Networks, focusing on the integration of the actual network with the SDN one.
We will show how it's possible to implement a QoS mechanism to keep the network up and to grant the users the possibility of using it outside
a laboratory.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \textit{\textbf{SDN}, Software Defined Network} is a new approach to create a network of computer and networks, which is based on the
most famous OpenFlow protocol. The key concepts of this type of networks are the hierarchy, the flow control and the statistics.
The hierarchy allows a network administrator to set up a controller which will choose how to make the network evolve setting up
how the routing will switch the traffic and doing so, how the single switch will behave.
The flow control is done by the controller that, running the algorithm the network administrator created, choose the path for a
type of traffic that occurs between a source and a destinaion host.
The most important and interesting thing about those type of networks is the statistics: every switching unit collects data from its
interfaces (e.g. the number of packets going in or out through a specific port or interface). This allow the controller to instantly
how the network is evolving and, if necessary, switch a traffic from congestioned links to other that, in that moment, are almost fully loaded.
Exploiting the statistic capabilities is the key to manage wisely a network and, since there aren't many papers about these potentialities,
we decided to focus on that. In this paper we will show how we handled the contemporary traffic between multiple host, both inside SDN network
nd between the SDN one and the external internet. Starting from traffic analysis and going through the SDN state of the art, we implemented some function
to exploit better the SDN POX controller and we reach the end of this paper showing how we found three different types of traffic and
for each one, we provided different metrics and different path to optimize the load. We realized the controller using POX, a python-based controller
and we tested it with the switches that were provided us in the Communication Networks Lab at DEI.
\begin{itemize}
	\item What are we talking about: description of the addressed problem. - done 
\item Motivation: why the problem is important. done
\item Novelty: how you contribute to advance the state of the art. done
\item Results: summary of the main findings  done
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}\label{sec:sota}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Related Work section contains an analysis of the most relevant related literature (remarking the shortcomings that are addressed in your work)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System Model}\label{sec:symo}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The System model is a description of your operating assumptions with related motivation and justification

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec:res}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The Results section contains a selection of the most relevant results with the explanation of their meaning. Please, not that you do NOT have to describe the shape of the curves that can be seen in the figures, but the reasons WHY such curves have that shape!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}\label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"

\end{document}

### 2017-01-30 report/paper.tex
added related work

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The Related Work section contains an analysis of the most relevant related literature (remarking the shortcomings that are addressed in your work)
+A good introduction about the SDN can be the  "Simulation in an SDN network scenario using the
+POX Controller" where has been presented the SDN environment and the mininet simulation
+tool that creates a topology where to simulate the switches, the traffic between hosts and
+allow the controller to communicate with the switches. The work presented in
+"Efficient topology discovery in OpenFlow-based Software Defined Networks" by F. Pakzad [put cit]
+has been implemented to improve the controller speed and efficiency as it make it more reactive.
+The introduction to the statistics power in a SDN network using the OpenFlow protocol has been found
+in the paper called "Getting traffic statistics from network devices in an
+SDN environment using OpenFlow" by D.J. Hamad & others. [ put note here]
+For traffic analysis we haven't found any paper describing the characteristics of
+some types of flows, so we used our PCs to sniff the traffic. Traffic analysis will be discussed in section \ref{sec:symo}.
 
@@
 The System model is a description of your operating assumptions with related motivation and justification
+
+\subsection{Traffic analysis}
+We used laptop connected via
+GigabitEthernet interface and Ethernet cable cat.5a to receive the traffic from the destination host.
+Our laptop has been connected via Wi-Fi to our access point via protocol 802.11n.
+No other traffic sources were active during the testing, so the router was handling only
+the traffic from laptop when sniffing traffic. The connection used was an ADSL2+ via copper.
+We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
+#Mb for the gaming traffic, #Mb for the VoIP traffic and 3GB for the torrent traffic.
+The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP)
 

### 2017-01-30 report/paper.tex
Paper relation   - added traffic analysis TODO: insert data   - added Intro and OpenFlow messages

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{LaTeX Template for a Scientific Paper}
+\title{Quality of Service with Software Defined Network.}
 
@@
 and we tested it with the switches that were provided us in the Communication Networks Lab at DEI.
+\subsection{Known Limitations}
+SDN networks are known to be underperforming with low traffic, which means that if nodes tries to connect with each other
+to exchange little traffic, the network will have a higher delay and will perform worst than a
+traditional network. We tried to reduce this underperformance with the usage of the wildcards:
+Every rule needs at least the source ip or the destination ip. If one of them is in a network
+(e.g. the internet traffic is in the network 0.0.0.0/0 excluded 192.168.10.0/24 ),
+we can redirect all the traffic from an internal host to the internet with a predefined
+path. This means that only one PacketIn will be sent to the controller if a host is surfing the network.
+
+Another limitation is the no default routing from a source to a destination: when an host
+is plugged to a network of SDN switches, it won't know the destinatio. The only way it can
+discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
+slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
+the network disposition, but doesn't know each host how much traffic is doing. For that the administrator need to put a firewall.
+The last one isn't needed, as the controller can set a switch to drop all the packets from/to a destination.
+
+\subsection{Openflow 1.0 messages}
+Openflow defines a number of messages between a controller and a switch. Now there will be presented some messages we exploited:
+\begin{itemize}
+	\item[Hello] this message is exchanged between a switch who is looking for a controller after the
+	switch was turned on. After receiving it, the controller will send the FeatureReq packet to that switch.
+	\item[FeatureReq] this is the message that the controller sends to a switch to know its abilities. The switch
+	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
+	With the featureRes the controller will know how many ports a switch has and their properties (e.g. if they are connected, link speed,...), the flow table capabilities and the supported statistics.
+	\item[PortStatus] this message is sent from a switch to the controller, to notify that some ports has changed their status.
+		It's useful to check if a port has someone connected
+	\item[StatsReq] The controller will send this message to request the statistics. The payload of the message contains the type of statistic the controller want to know.
+		The switch then will answer with the Stats reply, if it supports that type of stat. It will send a Bad_Request message if not.
+		\item[PacketIn] This message is sent from the switch to the controller whenever a packet arrives and it doesn't match
+		any rule. The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
+		a flow rule for packets like this using a FlowMod.
+		\item[PacketOut] This message is sent from the controller to forward a packet, which is the payload of the PacketOut message, to some
+		port(s).
+		\item[FlowMod] This message set, delete or modify a rule inside a switch. Is the message that allows routing.
+\end{itemize}
+
 \begin{itemize}
@@
 
+We developed a controller that can handle a mesh network, which has loops, that avoid
+the traffic flooding. The controller stores the switch and the connection between them;
+knows which hosts are connected to each switch port, with the hypothesys that only one
+host is connected at each port, unless the port is the one connected to the internet.
+With the use of the statistics, the controller get the status of the network and, if necessary,
+tries to adjust the traffic and put in the link with the higher path loss the torrent traffic,
+make the connection between two gaming host the one with less delay and tries to redistribute all the traffic in a way
+such that all the nodes have an average throughput as lowest as possible.
+
 \subsection{Traffic analysis}
@@
 The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP)
+
+The traffic analysis of the torrent protocol shows that the average packet size is the MTU
+(i.e. 1500bytes) and the protocol tryies to exploit all the channel capabilities. The traffic
+is made via UDP packets with the port range in the ephimeral ports (2000_- 65535).
+Moreover, the node is linked to a lot of IPs with multiple connection: this means that
+if someone wants to find a p2p traffic, it can look to the statistics and see if the traffic is loading all the channel,
+to how many IPs, an host is connected to and if it uses UDP.
+For the VoIP we saw that the protocol uses UDP and the traffic, after a transition time, goes only between two hosts.
+The average packet size is the MTU and the protocol tryies to exploit all the channel for a better QoE.
+On another test, we tried the VoIP while a download was ongoing, and we saw that the quality
+decreased to allow communication. The traffic is between the UDP port 80 ##TODO## TOCHECK ##
+
+The last traffic analysis we made was with a DOTA[explain what it is] web match.
+We saw that for that type of gaming session, the packet size is about # bytes and involves both TCP and UDP ports.
 

### 2017-01-30 report/paper.tex
added openflow messages

@@
 \begin{itemize}
-	\item[Hello] this message is exchanged between a switch who is looking for a controller after the
-	switch was turned on. After receiving it, the controller will send the FeatureReq packet to that switch.
+	\item[Hello] This message is sent both from the controller and the switch who is looking for a controller after it
+	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
 	\item[FeatureReq] this is the message that the controller sends to a switch to know its abilities. The switch

### 2017-01-30 pox/ext/graphUpdater.py
Need some changes:   - the host object should have, in the connectedTo the     1. destination host     2. type of traffic (tcp, udp, none)     3. the path that connects the host to the destination host   - the graph updater should let the changes for only one type of traffic   (i.e. tcp or udp) and should remove old rules.   No other changes happen in my mind right now

@@ -1,0 +1,11 @@
+import my_topo_SDN as topo
+import my_controller
+import stats_handler as sh
+import pox.openflow.libopenflow_01 as of
+from pox.lib.addresses import IPAddr, EthAddr
+
+def checkChanges():
+    for h in topo.hosts:
+        if h.gaming: # redirect tcp through the least delay
+            for dstH,path in h.connectedTo:
+                sw_list = nx.dijkstra_path(get_gf(DELAY_OPT), source=h.ip, target=dstH.ip, weight='weight')

### 2017-01-30 pox/ext/my_topo_SDN.py

@@ -308,7 +308,7 @@ class Host():
             raise("Invalid argument. ip address must be an IPAddr object")
         self.isGaming=False
         self.traffic = False
-        self.connectedTo = []
+        self.connectedTo = {}
         self.switch = ( dpid , portN) # tuple for the
         self.ip=ipAddr
         self.mac=macAddr
@@ -324,16 +324,18 @@ class Host():
     def getTraffic():
         return self.traffic
 
-    def addConnection(ip,path=None):
+    def addConnection(host,path=None):
         #update timer if ip exist
         if(ip in self.connectedTo):
-            for p,tup in self.connectedTo:
-                self.connectedTo[p][1] = datetime.datetime.now()
+            for host,tup in self.connectedTo:
+                self.connectedTo[host][1] = datetime.datetime.now()
         #add ip
-        self.connectedTo.append([])
-        self.connectedTo[-1].append((ip,datetime.datetime.now()))
+        self.connectedTo.[host] = []
+        self.connectedTo[host].append(datetime.datetime.now())
         if path:
-            self.connectedTo[-1].append(path) # add the path to reach the node
+            for i,p in enumerate(path):
+                if connectedTo.path
+            self.connectedTo[host].append(path) # add the path to reach the node
 
     def isConnected(ip):
         """if is connected return the time since when it was connected [datetime.datetime]

### 2017-01-30 pox/ext/stats_handler.py

@@ -9,7 +9,7 @@ log = core.getLogger()
 
 # from enum import Enum
 # from graphUpdater import GraphUpdater
-UPD_GRAPH = 1 # every 10 seconds update the graph weight
+UPD_GRAPH = 10 # every 10 seconds update the graph weight
 GAMING_THRES = 10 #bytes per packet
 
 dpid = list()

### 2017-01-30 pox/ext/my_controller.py
Need some changes:   - the host object should have, in the connectedTo the     1. destination host     2. type of traffic (tcp, udp, none)     3. the path that connects the host to the destination host   - the graph updater should let the changes for only one type of traffic   (i.e. tcp or udp) and should remove old rules.   No other changes happen in my mind right now

@@ -246,6 +246,7 @@ def _show_topo():
     job_for_another_core = multiprocessing.Process(target=topo.save_graph,args=()) #chiama la funzione save_graph in un processo separato
     job_for_another_core.start()
 
+
 def launch(__INSTANCE__=None, **kw):
     """
     start:
@@ -279,3 +280,4 @@ def launch(__INSTANCE__=None, **kw):
 
     Timer(5, _show_topo, recurring=True) #every 2 seconds execute _show_topo
     Timer(30, topo.ipCleaner, recurring = True) # every 30s clean the old connection ip
+    Timer(5, _checkChanges, recurring = True) # change the graph if something happened

### 2017-01-30 pox/ext/stats_handle.py

@@ -1,101 +1,0 @@
-from pox.core import core
-from pox.lib.util import dpid_to_str
-import pox.openflow.libopenflow_01 as of
-import time
-from pox.lib.recoco import Timer  #per eseguire funzioni ricorsivamente
-import my_topo_SDN as myTopo
-log = core.getLogger()
-from enum import Enum
-
-stats = list() # create the stats
-stats.append({})
-stats.append({})
-stats.append({})
-stats.append({})
-# def saluta():
-#     log.debug("\n\n ciao ciao \n\n")
-
-PORT=1
-TABLE=2
-FLOW=3
-QUEUE=0
-
-def _saveStats(sType,dpid, new_stats):
-    if not (dpid in stats[sType]):
-        stats[sType].update({dpid:""}) # add the dpid dictionary
-        stats[sType][dpid]= {} #create the space to save the stat in this
-
-    stats[sType][dpid]=new_stats # overwrite the stats
-    log.debug("added the stats for dpid %i",dpid)
-
-def getStats(sType, dpid, port=None, table=None,flow=None):
-    """ Try to get the stats for the specified dpid. If no stats is found, return None
-    """
-    if sType is None:
-        raise ValueError("type cannot be None, use a type from StatsType")
-    if dpid is None:
-        raise ValueError("dpid cannot be None, use a valid dpid")
-
-    if sType == PORT:
-        return _getPortStat(dpid,port)
-    if sType == TABLE:
-        return _getTableStat(dpid,table)
-    if sType == QUEUE:
-        return _getQueueStat(dpid,port)
-    if sType == FLOW:
-        return _getQueueStat(dpid,flow)
-    # should never reach there
-    else:
-            raise ValueError("stats type was not recognized, please choose one from StatsType")
-
-def _getQueueStat(dpid, port):
-    if not dpid in stats[StatsType.QUEUE]: # check if I have some stats for this dpid
-        return None;
-    if port is None: # check if all stats are needed
-        return stats[StatsType.QUEUE][dpid] # list of dictionary
-    if not port in stats[StatsType.QUEUE][dpid]: # if no stats is available for that port
-        return None
-    return stats[StatsType.QUEUE][dpid][port-1] # dictionary with stats of that port
-
-def _getTableStat(dpid,table):
-    if not dpid in stats[StatsType.TABLE]: # check if I have some stats for this dpid
-        return None;
-    if table is None: # check if all stats are needed
-        return stats[StatsType.TABLE][dpid] # list of dictionary
[…]

### 2017-01-31 pox/ext/my_controller.py
Added internal network support:   - Internal IPs flooded if no other better rule is satisfied   - TCP and UDP traffic are differentiated   TODO:   - set removePath   - set modifyFlow   - drop other private IPs   - flood the network for priate IPs (this or the above)   - test

@@ -40,7 +40,7 @@ log = core.getLogger()
 
 # SDN_network = "10.0.0.0/24"
 # SDN_NETMASK = "255.255.255.0"
-SDN_network = "192.168.0.0/16"
+SDN_network = ""#"192.168.0.0/16"
 PCK_ERROR_OPT = 1
 DELAY_OPT     = 2
 DEFAULT_OPT   = 3
@@ -96,7 +96,12 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
             #probably useless
             sw.port_mac[port.port_no]=port.hw_addr
             sw.mac_port[port.hw_addr]=port.port_no
+            #add the link to the external
+            topo.add_link(event.dpid,port.port_no,topo.hosts[0].ip,port.port_no,isHost=True)
 
+        #default rules
+        if SDN_network != "": # I set the default network
+            topo.add_default_rules(event.connection.dpid, SDN_network)
 
 
     #verificare che sua uno switch openflow
@@ -169,27 +174,33 @@ def _handle_ip_packet(event):
         if ip_src == hst.ip:
             alreadySrc=True
             log.debug("src host already found.")
-            for dst in hst.connectedTo:
-                if ip_dst in dst[0][0]:
-                    log.debug("connection found before")
-                    alreadyDst =True
+            if hst.isConnected(ip_dst,Host.TRANSP_BOTH):
+                log.debug("connection found before")
+                alreadyDst =True
+
 
     if not alreadySrc and not (IPAddr(ip_src) == IPAddr('100.100.100.0') or IPAddr(ip_src) == IPAddr('100.100.100.1')):
         #add host to host list
         h = topo.Host(event.dpid,event.in_port, ipAddr=ip_src,macAddr=src_mac)
         h.addConnection(ip_dst)
         topo.hosts.append(h)
-    elif not alreadyDst and not (ip_dst.inNetwork(SDN_network, netmask=SDN_NETMASK)):
+    elif not alreadyDst and not ip_dst.inNetwork(SDN_network):#, netmask=SDN_NETMASK)):
         for h in topo.hosts:
             if h.ip == ip_src:
                 if h.isGaming:
                     log.debug("Adding link with the lowest delay path")
-                    #TODO add with delay graph
+                    path = topo.get_path(topo.get_gf(topo.DELAY_OPT),h.ip, ip_dst)
+                    h.addConnection(ip_dst,path)
[…]

### 2017-01-31 pox/ext/my_topo_SDN.py

@@ -26,6 +26,8 @@ hosts=[] # list of object of type Host
 __DEFAULT_RULES_PRIORITY = 50
 __DEFAULT_ARP_PATH = 150
 __DEFAULT_IP_PATH = 1000
+__DEFAULT_EXT_NET_RULE=10
+__DEFAULT_INT_NET_RULE= 30
 
 IP_TIMEOUT = 60 #seconds
 PCK_ERROR_OPT = 1
@@ -118,7 +120,7 @@ def save_graph():
     plt.clf()                        #elimina l'immagine corrente dalla libreria
 
 
-def add_link(dpid1, port1, dpid2, port2):
+def add_link(dpid1, port1, dpid2, port2, isHost=False):
     """
     Inside add_link function, add switch and link
     """
@@ -126,8 +128,9 @@ def add_link(dpid1, port1, dpid2, port2):
     add_switch(dpid2)
     switch[dpid1].port_dpid[port1] = dpid2
     switch[dpid1].dpid_port[dpid2] = port1
-    switch[dpid2].port_dpid[port2] = dpid1
-    switch[dpid2].dpid_port[dpid1] = port2
+    if not isHost: # do the reverse only if the other is a switch
+        switch[dpid2].port_dpid[port2] = dpid1
+        switch[dpid2].dpid_port[dpid1] = port2
     grafo.add_edge(dpid1, dpid2)
     pck_error_gf.add_edge(dpid1, dpid2, weight=0)
     delay_gf.add_edge(dpid1, dpid2, weight=1)
@@ -186,7 +189,7 @@ def get_gf(option):
     elif option == DEFAULT_OPT:
         return grafo
 
-def add_path_to_gw(ip_int, ip_ext, option):
+def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
 
     path = nx.dijkstra_path(get_gf(option), source=ip_int, target=ip_ext, weight='weight')
     #devo fare in modo di riconoscere uno switch come gateway
@@ -230,10 +233,10 @@ def add_path(ip_src, ip_dst, option):
         msg.actions.append(of.ofp_action_output(port = pt_pre_hope ))
         core.openflow.sendToDPID(sw_list[i], msg)
 
-def add_default_rules(dpid):
+def add_default_rules(dpid, net = None):
     """
     add default rules on new switch
-    arp request flooding
+    arp request flooding and in/out network traffic
     """
     msg = of.ofp_flow_mod()
[…]

### 2017-01-31 pox/ext/my_topo_SDN.py
Added flow modication

@@ -189,9 +189,109 @@ def get_gf(option):
     elif option == DEFAULT_OPT:
         return grafo
 
-def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
+def get_path(ip_int, ip_ext, option):
+    return nx.dijkstra_path(get_gf(option), source=ip_int, target=ip_ext, weight='weight')
 
-    path = nx.dijkstra_path(get_gf(option), source=ip_int, target=ip_ext, weight='weight')
+def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
+    newPath=get_path(ip_int,hosts[0].ip,option):
+    h = [host.ip == ip_int for host in hosts ]
+    oldPath;
+    if option == PCK_ERROR_OPT:
+        oldPath = h.isConnected(ip_ext)
+        if oldPath is False:
+        #add path as it's the first time:
+            for i in range (1, len(newPath) - 2):
+                #install fluxes from ip_src to ip_dst
+                msg = of.ofp_flow_mod()
+                msg.priority = __DEFAULT_IP_PATH
+                msg.match.nw_dst = IPAddr(str(ip_dst))
+                msg.match.nw_proto = 17 # UDP
+                msg.match.dl_type = 0x800 #ip
+                pt_next_hop = switch[sw_list[i]].dpid_port[sw_list[i+1]]
+                msg.actions.append(of.ofp_action_output(port = pt_next_hop ))
+                core.openflow.sendToDPID(sw_list[i], msg) #switch i-esimo
+            #the reverse
+            for i in range (2, len(newPath) - 1):
+                #install fluxes from ip_dst to ip_src
+                msg = of.ofp_flow_mod()
+                msg.priority = __DEFAULT_IP_PATH
+                msg.match.nw_dst = IPAddr(str(ip_src))
+                msg.match.dl_type = 0x800 #ip
+                msg.match.nw_proto = 17 #UDP
+                pt_pre_hop = switch[sw_list[i]].dpid_port[sw_list[i-1]]
+                msg.actions.append(of.ofp_action_output(port = pt_pre_hop ))
+                core.openflow.sendToDPID(sw_list[i], msg) #switch i-esimo
+            h.addConnection(hosts[0],path,UDP)
+        else:
+            for i in range(2,max(len(oldPath,newPath))-2):
+                if oldPath[i-1] == newPath[i-1]: #same switch
+                    if oldPath[i] == newPath[i]: # also the same next hop
+                        if oldPath[i+1] == newPath[i+1]:
[…]

### 2017-02-01 pox/ext/my_controller.py
merge andrea

@@ -109,8 +109,6 @@ def _handle_ip_packet(event):
         pass
     elif (IP.IPv4Address(ip_src) in IP.IPv4Network(SDN_network)):
         """ sorgente e' nella sotto rete SDN """
-        #verifico se gia' presente nel grafo
-        #log.debug("sorgente nella rete SDN")
         #log.debug("is_src %s, il_log: %s", ip_src, topo.is_logged(ip_src) )
         if (topo.is_logged(ip_src)): #se non e' presente lo aggiungo
             pass

### 2017-02-01 pox/ext/my_controller.py
Corrected type errors in code. need to be tested. continuing until code is completed.

@@ -35,7 +35,7 @@ import ipaddress as IP
 import multiprocessing #multiprocess
 
 import my_topo_SDN as topo #new class
-
+import graphUpdater as gu
 log = core.getLogger()
 
 # SDN_network = "10.0.0.0/24"
@@ -100,8 +100,8 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
             topo.add_link(event.dpid,port.port_no,topo.hosts[0].ip,port.port_no,isHost=True)
 
         #default rules
-        if SDN_network != "": # I set the default network
-            topo.add_default_rules(event.connection.dpid, SDN_network)
+        # if SDN_network != "": # I set the default network
+        #     topo.add_default_rules(event.connection.dpid, SDN_network)
 
 
     #verificare che sua uno switch openflow
@@ -209,7 +209,7 @@ def _handle_ip_packet(event):
         """ pacchetti di controllo """
         #non esegue nulla in quanto se ne occupa network_performance
         pass
-    elif (ip_src.in_network(SDN_network):#, netmask=SDN_NETMASK)):
+    elif (ip_src.in_network(SDN_network)):#, netmask=SDN_NETMASK)):
         """ sorgente e' nella sotto rete SDN """
         #verifico se gia' presente nel grafo
         #log.debug("sorgente nella rete SDN")
@@ -278,7 +278,7 @@ def launch(__INSTANCE__=None, **kw):
                 topo.hosts.append(h) # add the host as the first one
                 # print("added ext host" )
                 log.info("added the external host and ready for finding the switch/port to which forward all external packets")
-        elif(k.find "net")>-1:
+        elif k.find("net")>-1 :
             log.debug("parsing network address")
             if len(v) >=9 and len(v)<=18 : # "192.168.240.240/24" is the net address form
                 n = v.split('/') # divide the cidr notation
@@ -298,4 +298,4 @@ def launch(__INSTANCE__=None, **kw):
 
     Timer(5, _show_topo, recurring=True) #every 2 seconds execute _show_topo
     Timer(30, topo.ipCleaner, recurring = True) # every 30s clean the old connection ip
-    Timer(5, _checkChanges, recurring = True) # change the graph if something happened
+    Timer(5, gu.checkChanges, recurring = True) # change the graph if something happened

### 2017-02-01 pox/ext/my_topo_SDN.py

@@ -193,7 +193,10 @@ def get_path(ip_int, ip_ext, option):
     return nx.dijkstra_path(get_gf(option), source=ip_int, target=ip_ext, weight='weight')
 
 def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
-    newPath=get_path(ip_int,hosts[0].ip,option):
+    if isDpid:
+        #TODO dpid to internet
+        log.debug("")
+    newPath=get_path(ip_int,hosts[0].ip,option)
     h = [host.ip == ip_int for host in hosts ]
     oldPath;
     if option == PCK_ERROR_OPT:
@@ -207,7 +210,7 @@ def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
                 msg.match.nw_dst = IPAddr(str(ip_dst))
                 msg.match.nw_proto = 17 # UDP
                 msg.match.dl_type = 0x800 #ip
-                pt_next_hop = switch[sw_list[i]].dpid_port[sw_list[i+1]]
+                pt_next_hop = switch[sw_list[i]].dpid_port[sw_list[i+1]] #TODO
                 msg.actions.append(of.ofp_action_output(port = pt_next_hop ))
                 core.openflow.sendToDPID(sw_list[i], msg) #switch i-esimo
             #the reverse
@@ -218,7 +221,7 @@ def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
                 msg.match.nw_dst = IPAddr(str(ip_src))
                 msg.match.dl_type = 0x800 #ip
                 msg.match.nw_proto = 17 #UDP
-                pt_pre_hop = switch[sw_list[i]].dpid_port[sw_list[i-1]]
+                pt_pre_hop = switch[sw_list[i]].dpid_port[sw_list[i-1]] #TODO
                 msg.actions.append(of.ofp_action_output(port = pt_pre_hop ))
                 core.openflow.sendToDPID(sw_list[i], msg) #switch i-esimo
             h.addConnection(hosts[0],path,UDP)
@@ -267,7 +270,7 @@ def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
                                 pt_next_hope = switch[newPath[i]].dpid_port[newPath[i+1]]
                                 msg.actions.append(of.ofp_action_output(port = pt_next_hope ))
                                 core.openflow.sendToDPID(newPath[i], msg) #switch i-esimo
-                            h.addConnection(host[0],newPath,UDP)
[…]

### 2017-02-01 pox/ext/graphUpdater.py
OK with IP handling. rules checked graph checked stats: need merge to be done flow checked

@@ -7,5 +7,367 @@ from pox.lib.addresses import IPAddr, EthAddr
 def checkChanges():
     for h in topo.hosts:
         if h.gaming: # redirect tcp through the least delay
-            for dstH,path in h.connectedTo:
-                sw_list = nx.dijkstra_path(get_gf(DELAY_OPT), source=h.ip, target=dstH.ip, weight='weight')
+            for dstH,oldPath in h.connectedTo:
+                newPath = nx.dijkstra_path(get_gf(DELAY_OPT), source=h.ip, target=dstH.ip, weight='weight')
+                if len(oldPath)!= len(newPath):
+                    change_path(oldPath,newPath,topo.TCP)
+                    h.addConnection(dstH,newPath)
+                else:
+                    for i in range (1,len(oldPath)):
+                        if(oldPath[i]==newPath[i]):
+                            continue
+                        else:
+                            change_path(oldPath,newPath,topo.TCP)
+                            h.addConnection(dstH,newPath)
+        elif h.traffic:
+            for dstH,oldPath in h.connectedTo:
+                newPath = nx.dijkstra_path(get_gf(PCK_ERROR_MAX_OPT), source=h.ip, target=dstH.ip, weight='weight')
+                if len(oldPath)!= len(newPath):
+                    change_path(oldPath,newPath,topo.UDP)
+                    h.addConnection(dstH,newPath)
+                else:
+                    for i in range (1,len(oldPath)):
+                        if(oldPath[i]==newPath[i]):
+                            continue
+                        else:
+                            change_path(oldPath,newPath,topo.UDP)
+                            h.addConnection(dstH,newPath)
+
+        else: # default rule
+            for dstH,oldPath in h.connectedTo:
+                newPath = nx.dijkstra_path(get_gf(DEFAULT_OPT), source=h.ip, target=dstH.ip, weight='weight')
+                if len(oldPath)!= len(newPath):
+                    change_path(oldPath,newPath,topo.BOTH)
+                    h.addConnection(dstH,newPath)
+                else:
+                    for i in range (1,len(oldPath)):
+                        if(oldPath[i]==newPath[i]):
+                            continue
+                        else:
+                            change_path(oldPath,newPath,topo.BOTH)
+                            h.addConnection(dstH,newPath)
+
+
+def change_path(oldPath,newPath,opt):
+    if opt == topo.TCP:
[…]

### 2017-02-01 pox/ext/my_controller.py

@@ -71,17 +71,21 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
         if(port.curr & of.ofp_port_features_rev_map["OFPPF_10MB_HD"] or
                 port.curr & of.ofp_port_features_rev_map["OFPPF_10MB_FD"]):
             topo.switch[event.connection.dpid].port_capacity[port.port_no] = 10; #TODO constants
+            topo.capacity_gf(event.dpid,topo.switch[event.dpid].port_dpid[port.port_no],10)
             log.info("port %i is a 10Mbps",port.port_no)
         elif(port.curr & of.ofp_port_features_rev_map["OFPPF_100MB_HD"] or
             port.curr & of.ofp_port_features_rev_map["OFPPF_100MB_FD"]):
             topo.switch[event.connection.dpid].port_capacity[port.port_no] = 100; #TODO constants
+            topo.capacity_gf(event.dpid,topo.switch[event.dpid].port_dpid[port.port_no],100)
             log.info("port %i is a 100Mbps",port.port_no)
         elif(port.curr & of.ofp_port_features_rev_map["OFPPF_1GB_HD"] or
             port.curr & of.ofp_port_features_rev_map["OFPPF_1GB_FD"]):
             topo.switch[event.connection.dpid].port_capacity[port.port_no] = 1000; #TODO constants
+            topo.capacity_gf(event.dpid,topo.switch[event.dpid].port_dpid[port.port_no],1000)
             log.info("port %i is a 1Gbps",port.port_no)
         elif(port.curr & of.ofp_port_features_rev_map["OFPPF_10GB_FD"]):
             topo.switch[event.connection.dpid].port_capacity[port.port_no] = 10000; #TODO constants
+            topo.capacity_gf(event.dpid,topo.switch[event.dpid].port_dpid[port.port_no],10000)
             log.info("port %i is a 10Gbps",port.port_no)
 
         #Find the port where is connected the standard router
@@ -168,87 +172,118 @@ def _handle_ip_packet(event):
 
     ip_src = IPAddr(ip_pck.srcip) #ip sorgente
     ip_dst = IPAddr(ip_pck.dstip) #ip destinatario
-    alreadySrc = False
-    alreadyDst=False
-    for hst in topo.hosts:
-        if ip_src == hst.ip:
-            alreadySrc=True
-            log.debug("src host already found.")
-            if hst.isConnected(ip_dst,Host.TRANSP_BOTH):
[…]

### 2017-02-01 pox/ext/my_topo_SDN.py

@@ -23,16 +23,18 @@ ip_to_switch = {} #dizionario in cui l'ip solo le chiavi e i valori gli elementi
 mac_to_ip = {} #per gli host
 hosts=[] # list of object of type Host
 
-__DEFAULT_RULES_PRIORITY = 50
-__DEFAULT_ARP_PATH = 150
+__DEFAULT_RULES_PRIORITY = 500
+__DEFAULT_ARP_PATH = 50
 __DEFAULT_IP_PATH = 1000
 __DEFAULT_EXT_NET_RULE=10
 __DEFAULT_INT_NET_RULE= 30
 
 IP_TIMEOUT = 60 #seconds
-PCK_ERROR_OPT = 1
+PCK_ERROR_MIN_OPT = 1
+PCK_ERROR_MAX_OPT = 4
 DELAY_OPT     = 2
 DEFAULT_OPT   = 3
+LOAD_OPT = 5
 
 def add_host(dpid, mac, port, ip):
     """
@@ -195,12 +197,26 @@ def get_path(ip_int, ip_ext, option):
 def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
     if isDpid:
         #TODO dpid to internet
+
         log.debug("")
     newPath=get_path(ip_int,hosts[0].ip,option)
     h = [host.ip == ip_int for host in hosts ]
     oldPath;
+    add_path(ip_int,ip_dst, option,isExt=True)
+    #devo fare in modo di riconoscere uno switch come gateway
+    #capire nella realta' come arrivano i dpid
+
+def add_path(ip_src, ip_dst, option, isExt=False):
+    if isExt:
+        newPath=get_path(ip_int,hosts[0].ip,option)
+    else:
+        newPath=get_path(ip_src,ip_dst,option)
+
     if option == PCK_ERROR_OPT:
-        oldPath = h.isConnected(ip_ext)
+        h.addConnection(ip_ext,newPath,UDP)
+    else:
+        h.addConnection(ip_ext,newPath,TCP)
+    oldPath = h.isConnected(ip_ext)
         if oldPath is False:
         #add path as it's the first time:
             for i in range (1, len(newPath) - 2):
@@ -208,25 +224,27 @@ def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
                 msg = of.ofp_flow_mod()
                 msg.priority = __DEFAULT_IP_PATH
                 msg.match.nw_dst = IPAddr(str(ip_dst))
[…]

### 2017-02-01 pox/ext/stats_handler.py

@@ -10,7 +10,8 @@ log = core.getLogger()
 # from enum import Enum
 # from graphUpdater import GraphUpdater
 UPD_GRAPH = 10 # every 10 seconds update the graph weight
-GAMING_THRES = 10 #bytes per packet
+GAMING_THRES = 200 #bytes per packet
+TRAFFIC_THRES = 0.8 #
 
 dpid = list()
 DESC_STATS = 1
@@ -181,8 +182,15 @@ def  _setAvgPktSize(stat,dpid):
             if(avgPktS<GAMING_THRES): #byte for packet
                 log.debug("on dpid %s there's an host who is gaming. pktSize is %.3f",dpid_to_str(dpid),avgPktS)
                 myTopo.switch[dpid].host_gaming[port["Pnum"]] = True
+                for h in myTopo.hosts:
+                    if h.ip == myTopo.switch[dpid].port_dpid[port["Pnum"]]:
+                        h.isGaming = True
             else:
                 myTopo.switch[dpid].host_gaming[port["Pnum"]] = False
+                for h in myTopo.hosts:
+                    if h.ip == myTopo.switch[dpid].port_dpid[port["Pnum"]]:
+                        h.isGaming = False
+
 
 
 def _setTraffic(flow_stat,dpid):
@@ -201,9 +209,18 @@ def _setTraffic(flow_stat,dpid):
     #set the traffic load [0;1] 0= no traffic. 1= link is full
         utilization = avgTrafPort/myTopo.switch[dpid].port_capacity[1]
         log.debug("Link utilization is %.4f",utilization)
-        myTopo.link_capacity(dpid, dpid2, utilization)
-
-
+        myTopo.link_load(dpid, dpid2, utilization)
+        if(utilization>TRAFFIC_THRES): #byte for packet
+            log.debug("on dpid %s there's an host who is congestioning. Throughput/Capacity=%.3f",dpid_to_str(dpid),utilization)
+        myTopo.switch[dpid].host_traffic[port["Pnum"]] = True
+        for h in myTopo.hosts:
+            if h.ip == myTopo.switch[dpid].port_dpid[port["Pnum"]]:
+                h.traffic = True
+            else:
+                myTopo.switch[dpid].host_traffic[port["Pnum"]] = False
+                for h in myTopo.hosts:
+                    if h.ip == myTopo.switch[dpid].port_dpid[port["Pnum"]]:
+                        h.traffic = False
 
 def launch():
     core.openflow.addListenerByName("FlowStatsReceived", _handle_flow_stats)

### 2017-02-01 pox/ext/my_controller.py
deleted possibly useless code

@@ -192,99 +192,7 @@ def _handle_ip_packet(event):
             if(h.ip == ip.src):
                 h.addConnection(ip_dst)
         log.debug("\n %s aggiunto nella rete", ip_src)
-    #
-    # for hst in topo.hosts:
-    #     if ip_src == hst.ip:
-    #         alreadySrc=True
-    #         log.debug("src host already found.")
-    #         srcH = hst
-    #         if hst.isConnected(ip_dst,Host.TRANSP_BOTH):
-    #             log.debug("connection found before")
-    #     if ip_dst== hst.ip:
-    #         dstH = hst
-    #         alreadyDst = True
-    #
-    #
-    #
-    # if not alreadySrc and not alreadyDst and ip_src.inNetwork(SDN_network):
-    #     #add host to host list
-    #     h = topo.Host(event.dpid,event.in_port, ipAddr=ip_src,macAddr=src_mac)
-    #     topo.add_host(event.connection.dpid,src_mac,event.port,ip_src) # add ip to graph
-    #     h.addConnection(ip_dst)
-    #     topo.hosts.append(h)
-    #     alreadySrc =True # now I know the SRC
-    # if not alreadySrc and alreadyDst: # internal
-    #     h = topo.Host(event.dpid,event.in_port, ipAddr=ip_src,macAddr=src_mac)
-    #     topo.add_host(event.connection.dpid,src_mac,event.port,ip_src) # add ip to graph
-    #     topo.add_path(ip_src,ip_dst,LOAD_OPT)
-    #     h.addConnection()
-    if alreadySrc and not alreadyDst and not ip_dst.inNetwork(SDN_network):# def rule 4
-        for h in topo.hosts:
-            if h.ip == ip_src:
-                if h.isGaming:
-                    log.debug("Adding link with the lowest delay path")
-                    # path = topo.get_path(topo.get_gf(topo.DELAY_OPT),h.ip, host[0].ip)
-                    topo.add_path_through_gw(h.ip, ip_dst, topo.DELAY_OPT)
-
-                elif h.traffic:
-                    log.debug("Adding link with the worst graph")
-                    # path = topo.get_path(topo.get_gf(topo.PCK_ERROR_MAX_OPT))
-                    # h.addConnection(ip_dst,path)
-                    topo.add_path_through_gw(h.ip, ip_dst, topo.PCK_ERROR_MAX_OPT)
-                else:
-                    # path = topo.get_path(topo.get_gf(topo.LOAD_OPT))
[…]

### 2017-02-01 pox/ext/my_controller.py
add min_pck error

@@ -85,7 +85,7 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
             log.info("port %i is a 10Gbps",port.port_no)
 
         #Find the port where is connected the standard router
-        if port.hw_addr == topo.hosts[0].mac:
+        if port.hw_addr == topo.hosts[0].mac: #is the switch with route to internet
             log.info("found port to the internet")
             EXTERNAL = (event.dpid,port.port_no)
             topo.hosts[0].switch=EXTERNAL
@@ -98,6 +98,8 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
             sw.mac_port[port.hw_addr]=port.port_no
             #add the link to the external
             topo.add_link(event.dpid,port.port_no,topo.hosts[0].ip,port.port_no,isHost=True)
+            #all ip_dst == external through port
+            topo.add_default_ext_roules(event.connection.dpid, SDN_network, port)
 
         #default rules
         if SDN_network != "": # I set the default network
@@ -168,47 +170,55 @@ def _handle_ip_packet(event):
 
     ip_src = IPAddr(ip_pck.srcip) #ip sorgente
     ip_dst = IPAddr(ip_pck.dstip) #ip destinatario
-    alreadySrc = False
-    alreadyDst=False
-    for hst in topo.hosts:
-        if ip_src == hst.ip:
-            alreadySrc=True
-            log.debug("src host already found.")
-            if hst.isConnected(ip_dst,Host.TRANSP_BOTH):
-                log.debug("connection found before")
-                alreadyDst =True
 
-
-    if not alreadySrc and not (IPAddr(ip_src) == IPAddr('100.100.100.0') or IPAddr(ip_src) == IPAddr('100.100.100.1')):
-        #add host to host list
-        h = topo.Host(event.dpid,event.in_port, ipAddr=ip_src,macAddr=src_mac)
-        h.addConnection(ip_dst)
-        topo.hosts.append(h)
-    elif not alreadyDst and not ip_dst.inNetwork(SDN_network):#, netmask=SDN_NETMASK)):
-        for h in topo.hosts:
-            if h.ip == ip_src:
-                if h.isGaming:
-                    log.debug("Adding link with the lowest delay path")
-                    path = topo.get_path(topo.get_gf(topo.DELAY_OPT),h.ip, ip_dst)
[…]

### 2017-02-01 pox/ext/my_topo_SDN.py

@@ -15,7 +15,8 @@ log = core.getLogger()
 switch = {} #dizionario di switch dpid e' la chiave
 grafo = nx.Graph()          #grafo con vari attributi
 switch_gf = nx.Graph()      #grafo con solo gli switch
-pck_error_gf = nx.Graph()    #grafo pesato secondo il pathloss
+pck_error_min_gf = nx.Graph()    #grafo pesato secondo il pathloss
+pck_error_max_gf = nx.Graph()    #grafo pesato secondo il complementarepathloss
 delay_gf = nx.Graph()       #delay del link
 capacity_gf = nx.Graph()    #capacita' max link
 load_gf = nx.Graph()        #percentuale del caricamento del link in base alla sua capacita' max
@@ -30,9 +31,11 @@ __DEFAULT_EXT_NET_RULE=10
 __DEFAULT_INT_NET_RULE= 30
 
 IP_TIMEOUT = 60 #seconds
-PCK_ERROR_OPT = 1
+PCK_ERROR_MIN_OPT = 1
+PCK_ERROR_MAX_OPT = 4
 DELAY_OPT     = 2
 DEFAULT_OPT   = 3
+LOAD_OPT      = 5
 
 def add_host(dpid, mac, port, ip):
     """
@@ -45,12 +48,14 @@ def add_host(dpid, mac, port, ip):
         mac_to_ip[mac] = ip
         ip_to_switch[ip] = switch[dpid]
         grafo.add_node(ip)
-        pck_error_gf.add_node(ip)
+        pck_error_max_gf.add_node(ip)
+        pck_error_min_gf.add_node(ip)
         capacity_gf.add_node(ip)
         load_gf.add_node(ip)
 
         grafo.add_edge(dpid, ip)
-        pck_error_gf.add_edge(dpid, ip, weight=0)
+        pck_error_min_gf.add_edge(dpid, ip, weight=0)
+        pck_error_max_gf.add_edge(dpid, ip, weight=0)
         capacity_gf.add_edge(dpid, ip, weight=0)
         load_gf.add_edge(dpid, ip, weight=0)
         log.debug("add host %s", ip)
@@ -66,7 +71,7 @@ def add_switch(dpid):
     else:
         switch[dpid] = my_Switch(dpid)
         grafo.add_node(dpid)
-        pck_error_gf.add_node(dpid)
+        pck_error_min_gf.add_node(dpid)
         delay_gf.add_node(dpid)
[…]

### 2017-02-01 pox/ext/graphUpdater.py
added stats for both minimum and maximum packet error

@@ -11,6 +11,7 @@ def checkChanges():
                 newPath = nx.dijkstra_path(get_gf(DELAY_OPT), source=h.ip, target=dstH.ip, weight='weight')
                 if len(oldPath)!= len(newPath):
                     change_path(oldPath,newPath,topo.TCP)
+                    change_path(oldPath,newPath,topo.UDP)
                     h.addConnection(dstH,newPath)
                 else:
                     for i in range (1,len(oldPath)):
@@ -18,6 +19,7 @@ def checkChanges():
                             continue
                         else:
                             change_path(oldPath,newPath,topo.TCP)
+                            change_path(oldPath,newPath,topo.UDP)
                             h.addConnection(dstH,newPath)
         elif h.traffic:
             for dstH,oldPath in h.connectedTo:
@@ -35,7 +37,7 @@ def checkChanges():
 
         else: # default rule
             for dstH,oldPath in h.connectedTo:
-                newPath = nx.dijkstra_path(get_gf(DEFAULT_OPT), source=h.ip, target=dstH.ip, weight='weight')
+                newPath = nx.dijkstra_path(get_gf(PCK_ERROR_MIN_OPT), source=h.ip, target=dstH.ip, weight='weight')
                 if len(oldPath)!= len(newPath):
                     change_path(oldPath,newPath,topo.BOTH)
                     h.addConnection(dstH,newPath)

### 2017-02-01 pox/ext/stats_handler.py

@@ -152,7 +152,7 @@ def _setPktLoss(stat,dpid):
     for i,PER in enumerate(pErrRate):
         try: #there isn't a link yet
             dpid2=myTopo.switch[dpid].port_dpid[i+1] # get the dpid connected to that port
-            myTopo.link_pathloss(dpid,dpid2,PER) # Update the packet error rate
+            myTopo.link_pck_err(dpid,dpid2,PER) # Update the packet error rate
         except:
             pass;
 

### 2017-02-01 pox/ext/graphUpdater.py
Solved flood error, but now no ping packet work. solving tomorrow.

@@ -6,12 +6,14 @@ from pox.lib.addresses import IPAddr, EthAddr
 
 def checkChanges():
     for h in topo.hosts:
-        if h.gaming: # redirect tcp through the least delay
-            for dstH,oldPath in h.connectedTo:
+        if h.isGaming: # redirect tcp through the least delay
+            for i,val in enumerate(h.connectedToTCP):
+                oldPath=h.connectedToTCP[h.connectedToTCP[val]][1]
+                if isinstance(oldPath,datetime.datetime):                    continue
+                dstH=val
                 newPath = nx.dijkstra_path(get_gf(DELAY_OPT), source=h.ip, target=dstH.ip, weight='weight')
                 if len(oldPath)!= len(newPath):
                     change_path(oldPath,newPath,topo.TCP)
-                    change_path(oldPath,newPath,topo.UDP)
                     h.addConnection(dstH,newPath)
                 else:
                     for i in range (1,len(oldPath)):
@@ -19,10 +21,28 @@ def checkChanges():
                             continue
                         else:
                             change_path(oldPath,newPath,topo.TCP)
+                            h.addConnection(dstH,newPath)
+            for i,val in enumerate(h.connectedToUDP):
+                oldPath=h.connectedToUDP[h.connectedToUDP[val]][1]
+                if isinstance(oldPath,datetime.datetime):                    continue
+                dstH=val
+                newPath = nx.dijkstra_path(get_gf(DELAY_OPT), source=h.ip, target=dstH.ip, weight='weight')
+                if len(oldPath)!= len(newPath):
+                    change_path(oldPath,newPath,topo.UDP)
+                    h.addConnection(dstH,newPath)
+                else:
+                    for i in range (1,len(oldPath)):
+                        if(oldPath[i]==newPath[i]):
+                            continue
+                        else:
                             change_path(oldPath,newPath,topo.UDP)
                             h.addConnection(dstH,newPath)
+
         elif h.traffic:
-            for dstH,oldPath in h.connectedTo:
+            for i,val in enumerate(h.connectedToUDP):
+                oldPath=h.connectedToUDP[h.connectedToUDP[val]][1]
+                dstH=val
+                if isinstance(oldPath,datetime.datetime):                    continue
                 newPath = nx.dijkstra_path(get_gf(PCK_ERROR_MAX_OPT), source=h.ip, target=dstH.ip, weight='weight')
                 if len(oldPath)!= len(newPath):
[…]

### 2017-02-01 pox/ext/my_controller.py

@@ -33,6 +33,7 @@ from pox.lib.addresses import IPAddr, EthAddr
 from pox.lib.util import str_to_dpid
 import ipaddress as IP
 import multiprocessing #multiprocess
+import datetime
 
 import my_topo_SDN as topo #new class
 import graphUpdater as gu
@@ -41,17 +42,39 @@ log = core.getLogger()
 # SDN_network = "10.0.0.0/24"
 # SDN_NETMASK = "255.255.255.0"
 SDN_network = ""#"192.168.0.0/16"
+
 PCK_ERROR_OPT = 1
 DELAY_OPT     = 2
 DEFAULT_OPT   = 3
-
-EXTERNAL = (None,None)
+hasGateway    = False
+hasNetwork    = False
+EXTERNAL = None
+unknown_link = []
+recent_packets = []
 
 def _handle_LinkEvent(event):
     """
     handle event ("LinkEvent") from openflow.discovery
     """
     l = event.link
+    #if there's a link, the controller knows the link between switches
+    for i,c in enumerate(unknown_link):
+        if c[0] == l.dpid1:
+            try:
+                dpid2 = switch[c[0]].port_dpid[c[1]]
+                topo.link_capacity(c[0],dpid2, c[2])
+                del unknown_link[i]# remove unknown link from the list
+            except:
+                break # doesn't have the other link yet
+        elif c[0] == l.dpid2:
+            try:
+                dpid = switch[c[0]].port_dpid[c[1]]
+                topo.link_capacity(c[0],dpid, c[2])
+                del unknown_link[i] # remove unknown link from the list
+            except:
+                break # doesn't have the other link yet
+        else:
+            continue
     if event.added:
         log.debug('LinkAdd dpid1: {0} porta {1}, dpid2: {2} porta {3}'.format(l.dpid1, l.port1, l.dpid2, l.port2))
         topo.add_link(l.dpid1, l.port1, l.dpid2, l.port2)
@@ -71,25 +94,37 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
         if(port.curr & of.ofp_port_features_rev_map["OFPPF_10MB_HD"] or
                 port.curr & of.ofp_port_features_rev_map["OFPPF_10MB_FD"]):
             topo.switch[event.connection.dpid].port_capacity[port.port_no] = 10; #TODO constants
-            topo.capacity_gf(event.dpid,topo.switch[event.dpid].port_dpid[port.port_no],10)
+            try:
[…]

### 2017-02-01 pox/ext/my_topo_SDN.py

@@ -24,11 +24,11 @@ ip_to_switch = {} #dizionario in cui l'ip solo le chiavi e i valori gli elementi
 mac_to_ip = {} #per gli host
 hosts=[] # list of object of type Host
 
-__DEFAULT_RULES_PRIORITY = 500
-__DEFAULT_ARP_PATH = 50
-__DEFAULT_IP_PATH = 1000
-__DEFAULT_EXT_NET_RULE=10
-__DEFAULT_INT_NET_RULE= 30
+DEFAULT_RULES_PRIORITY = 500
+DEFAULT_ARP_PATH = 50
+DEFAULT_IP_PATH = 1000
+DEFAULT_EXT_NET_RULE=10
+DEFAULT_INT_NET_RULE= 30
 
 IP_TIMEOUT = 60 #seconds
 PCK_ERROR_MIN_OPT = 1
@@ -106,7 +106,6 @@ def save_graph():
     plt.axis('off')
     plt.savefig("pck_error_max_gf.png")   #salva l'immagine
     plt.clf()                        #elimina l'immagine corrente dalla libreria
-
 
     edge_labels2=nx.draw_networkx_edge_labels(delay_gf,pos,font_size=12)
     nx.draw_networkx(delay_gf,pos, with_labels=True,node_color='blue',node_size=700, width=6,font_size=20,font_family='sans-serif')    #stampa anche il grafo
@@ -222,6 +221,7 @@ def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
     #devo fare in modo di riconoscere uno switch come gateway
     #capire nella realta' come arrivano i dpid
 
+
 def add_path(ip_src, ip_dst, option, isExt=False):
     if isExt:
         newPath=get_path(ip_int,hosts[0].ip,option)
@@ -233,89 +233,56 @@ def add_path(ip_src, ip_dst, option, isExt=False):
     else:
         h.addConnection(ip_ext,newPath,TCP)
     oldPath = h.isConnected(ip_ext)
-        if oldPath is False:
+    if oldPath is False:
         #add path as it's the first time:
-            for i in range (1, len(newPath) - 2):
-                #install fluxes from ip_src to ip_dst
-                msg = of.ofp_flow_mod()
-                msg.priority = __DEFAULT_IP_PATH
-                msg.match.nw_dst = IPAddr(str(ip_dst))
-                msg.match.nw_src = IPAddr(str(ip_src))
-                msg.match.nw_proto = 17 # UDP
-                msg.match.dl_type = 0x800 #ip
[…]

### 2017-02-01 pox/ext/stats_handler.py

@@ -212,14 +212,14 @@ def _setTraffic(flow_stat,dpid):
         myTopo.link_load(dpid, dpid2, utilization)
         if(utilization>TRAFFIC_THRES): #byte for packet
             log.debug("on dpid %s there's an host who is congestioning. Throughput/Capacity=%.3f",dpid_to_str(dpid),utilization)
-        myTopo.switch[dpid].host_traffic[port["Pnum"]] = True
+        myTopo.switch[dpid].host_traffic[table["actions"]] = True
         for h in myTopo.hosts:
-            if h.ip == myTopo.switch[dpid].port_dpid[port["Pnum"]]:
+            if h.ip == myTopo.switch[dpid].port_dpid[table["actions"]]:
                 h.traffic = True
             else:
-                myTopo.switch[dpid].host_traffic[port["Pnum"]] = False
+                myTopo.switch[dpid].host_traffic[table["actions"]] = False
                 for h in myTopo.hosts:
-                    if h.ip == myTopo.switch[dpid].port_dpid[port["Pnum"]]:
+                    if h.ip == myTopo.switch[dpid].port_dpid[table["actions"]]:
                         h.traffic = False
 
 def launch():

### 2017-02-02 pox/ext/my_controller.py
modifiche 02 feb

@@ -47,6 +47,9 @@ DEFAULT_OPT   = 3
 
 EXTERNAL = (None,None)
 
+counter = None
+
+
 def _handle_LinkEvent(event):
     """
     handle event ("LinkEvent") from openflow.discovery
@@ -104,7 +107,6 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
         #default rules
         if SDN_network != "": # I set the default network
             topo.add_default_rules(event.connection.dpid, SDN_network)
-
 
     #verificare che sua uno switch openflow
     log.debug("Add switch: %s", dpid_to_str(event.connection.dpid))
@@ -219,7 +221,7 @@ def _handle_ip_packet(event):
         """ pacchetti di controllo """
         #non esegue nulla in quanto se ne occupa network_performance
         return
-    elif (ip_src.in_network(SDN_network):#, netmask=SDN_NETMASK)):
+    elif (ip_src.in_network(SDN_network)): # netmask=SDN_NETMASK)):
         """ sorgente e' nella sotto rete SDN """
         #log.debug("is_src %s, il_log: %s", ip_src, topo.is_logged(ip_src) )
         if (topo.is_logged(ip_src)): #se non e' presente lo aggiungo
@@ -230,7 +232,7 @@ def _handle_ip_packet(event):
             h.addConnection(ip_dst)
             topo.hosts.append(h)
             log.debug("\n %s aggiunto nella rete", ip_src)
-    
+
         if (ip_dst.inNetwork(SDN_network, netmask=SDN_NETMASK)):
         #log.debug("sorgente nella rete SDN 3")
             """ destinatario nella sotto rete SDN """
@@ -265,7 +267,12 @@ def _show_topo():
     """
     function to show the graph on a separate process
     """
-    job_for_another_core = multiprocessing.Process(target=topo.save_graph,args=()) #chiama la funzione save_graph in un processo separato
+    global counter
+    if counter == None:
+        counter = 1
+    else:
+        counter = counter + 1
+    job_for_another_core = multiprocessing.Process(target=topo.save_graph(counter),args=()) #chiama la funzione save_graph in un processo separato
     job_for_another_core.start()
 
 
@@ -289,7 +296,7 @@ def launch(__INSTANCE__=None, **kw):
                 topo.hosts.append(h) # add the host as the first one
                 # print("added ext host" )
                 log.info("added the external host and ready for finding the switch/port to which forward all external packets")
[…]

### 2017-02-02 pox/ext/my_topo_SDN.py

@@ -11,6 +11,7 @@ import matplotlib.pyplot as plt  #for ploting graph
 import datetime #for connection expiration in host tuples
 import random #debug
 
+
 log = core.getLogger()
 switch = {} #dizionario di switch dpid e' la chiave
 grafo = nx.Graph()          #grafo con vari attributi
@@ -88,48 +89,51 @@ def rm_switch(dpid):
     load_gf.remove_node(dpid)
     del switch[dpid] #remove the disconnected switch
 
-def save_graph():
+def save_graph(counter):
+
+
     pos=nx.spring_layout(grafo) # positions for all nodes
     nx.draw_networkx(grafo,pos, with_labels=True, node_size=700, width=6, font_size=20,font_family='sans-serif')    #stampa anche il grafo
     plt.axis('off')
-    plt.savefig("grafo.png")      #salva l'immagine
+
+    plt.savefig("grafo%i.png" % (counter))      #salva l'immagine
     plt.clf()                     #elimina l'immagine corrente dalla libreria
 
     edge_labels=nx.draw_networkx_edge_labels(pck_error_min_gf,pos,font_size=12)
     nx.draw_networkx(pck_error_min_gf,pos, with_labels=True,node_color='green',node_size=700, width=6,font_size=20,font_family='sans-serif')    #stampa anche il grafo
     plt.axis('off')
-    plt.savefig("pck_error_min_gf.png")   #salva l'immagine
+    plt.savefig("pck_error_min_gf%i.png" % (counter))   #salva l'immagine
     plt.clf()                        #elimina l'immagine corrente dalla libreria
 
     edge_labels=nx.draw_networkx_edge_labels(pck_error_max_gf,pos,font_size=12)
     nx.draw_networkx(pck_error_max_gf,pos, with_labels=True,node_color='green',node_size=700, width=6,font_size=20,font_family='sans-serif')    #stampa anche il grafo
     plt.axis('off')
-    plt.savefig("pck_error_max_gf.png")   #salva l'immagine
+    plt.savefig("pck_error_max_gf%i.png" % (counter))   #salva l'immagine
     plt.clf()                        #elimina l'immagine corrente dalla libreria
 
 
     edge_labels2=nx.draw_networkx_edge_labels(delay_gf,pos,font_size=12)
[…]

### 2017-02-02 pox/ext/graphUpdater.py
controller working again with real switches   - TODO avoid first packet loss

@@ -9,7 +9,8 @@ def checkChanges():
         if h.isGaming: # redirect tcp through the least delay
             for i,val in enumerate(h.connectedToTCP):
                 oldPath=h.connectedToTCP[h.connectedToTCP[val]][1]
-                if isinstance(oldPath,datetime.datetime):                    continue
+                if isinstance(oldPath,datetime.datetime):
+                    continue
                 dstH=val
                 newPath = nx.dijkstra_path(get_gf(DELAY_OPT), source=h.ip, target=dstH.ip, weight='weight')
                 if len(oldPath)!= len(newPath):
@@ -24,7 +25,8 @@ def checkChanges():
                             h.addConnection(dstH,newPath)
             for i,val in enumerate(h.connectedToUDP):
                 oldPath=h.connectedToUDP[h.connectedToUDP[val]][1]
-                if isinstance(oldPath,datetime.datetime):                    continue
+                if isinstance(oldPath,datetime.datetime):
+                    continue
                 dstH=val
                 newPath = nx.dijkstra_path(get_gf(DELAY_OPT), source=h.ip, target=dstH.ip, weight='weight')
                 if len(oldPath)!= len(newPath):
@@ -42,7 +44,8 @@ def checkChanges():
             for i,val in enumerate(h.connectedToUDP):
                 oldPath=h.connectedToUDP[h.connectedToUDP[val]][1]
                 dstH=val
-                if isinstance(oldPath,datetime.datetime):                    continue
+                if isinstance(oldPath,datetime.datetime):
+                    continue
                 newPath = nx.dijkstra_path(get_gf(PCK_ERROR_MAX_OPT), source=h.ip, target=dstH.ip, weight='weight')
                 if len(oldPath)!= len(newPath):
                     change_path(oldPath,newPath,topo.UDP)
@@ -58,7 +61,8 @@ def checkChanges():
         else: # default rule
             for i,val in enumerate(h.connectedToTCP):
                 oldPath=h.connectedToTCP[h.connectedToTCP[val]][1]
-                if isinstance(oldPath,datetime.datetime):                    continue
+                if isinstance(oldPath,datetime.datetime):
+                    continue
                 dstH=val
                 newPath = nx.dijkstra_path(get_gf(PCK_ERROR_MIN_OPT), source=h.ip, target=dstH.ip, weight='weight')
                 if len(oldPath)!= len(newPath):

### 2017-02-02 pox/ext/my_controller.py

@@ -46,6 +46,7 @@ SDN_network = ""#"192.168.0.0/16"
 PCK_ERROR_OPT = 1
 DELAY_OPT     = 2
 DEFAULT_OPT   = 3
+
 hasGateway    = False
 hasNetwork    = False
 EXTERNAL = None
@@ -58,23 +59,23 @@ def _handle_LinkEvent(event):
     """
     l = event.link
     #if there's a link, the controller knows the link between switches
-    for i,c in enumerate(unknown_link):
-        if c[0] == l.dpid1:
-            try:
-                dpid2 = switch[c[0]].port_dpid[c[1]]
-                topo.link_capacity(c[0],dpid2, c[2])
-                del unknown_link[i]# remove unknown link from the list
-            except:
-                break # doesn't have the other link yet
-        elif c[0] == l.dpid2:
-            try:
-                dpid = switch[c[0]].port_dpid[c[1]]
-                topo.link_capacity(c[0],dpid, c[2])
-                del unknown_link[i] # remove unknown link from the list
-            except:
-                break # doesn't have the other link yet
-        else:
-            continue
+    # for i,c in enumerate(unknown_link):
+    #     if c[0] == l.dpid1:
+    #         try:
+    #             dpid2 = switch[c[0]].port_dpid[c[1]]
+    #             topo.link_capacity(c[0],dpid2, c[2])
+    #             del unknown_link[i]# remove unknown link from the list
+    #         except:
+    #             break # doesn't have the other link yet
+    #     elif c[0] == l.dpid2:
+    #         try:
+    #             dpid = switch[c[0]].port_dpid[c[1]]
+    #             topo.link_capacity(c[0],dpid, c[2])
+    #             del unknown_link[i] # remove unknown link from the list
+    #         except:
+    #             break # doesn't have the other link yet
+    #     else:
+    #         continue
     if event.added:
         log.debug('LinkAdd dpid1: {0} porta {1}, dpid2: {2} porta {3}'.format(l.dpid1, l.port1, l.dpid2, l.port2))
         topo.add_link(l.dpid1, l.port1, l.dpid2, l.port2)
@@ -94,6 +95,7 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
         if(port.curr & of.ofp_port_features_rev_map["OFPPF_10MB_HD"] or
[…]

### 2017-02-02 pox/ext/my_topo_SDN.py

@@ -37,6 +37,9 @@ DELAY_OPT     = 2
 DEFAULT_OPT   = 3
 LOAD_OPT      = 5
 
+TCP = 1 # same as Host
+UDP = 2
+
 def add_host(dpid, mac, port, ip):
     """
     add host on the graph and the port of switch
@@ -61,6 +64,19 @@ def add_host(dpid, mac, port, ip):
         log.debug("add host %s", ip)
     else:
         log.warning("Add host to switch that don't exist")
+
+        # msg = of.ofp_flow_mod()
+        # msg.dl_type = 0x800
+        # msg.match.dl_dst = mac
+        # # msg.match.nw_dst = IPAddr(ip,24)
+        # msg.actions.append(of.ofp_action_output(port=port))
+        # core.openflow.sendToDPID(dpid,msg)
+        #
+        # msg = of.ofp_flow_mod()
+        # msg.match.dl_dst = mac
+        # msg.dl_type = 0x806
+        # msg.actions.append(of.ofp_action_output(port=port))
+        # core.openflow.sendToDPID(dpid,msg)
 
 def add_switch(dpid):
     """
@@ -206,10 +222,13 @@ def get_gf(option):
     elif option == DEFAULT_OPT:
         return grafo
 
-def get_path(ip_int, ip_ext, option):
-    return nx.dijkstra_path(get_gf(option), source=ip_int, target=ip_ext, weight='weight')
+def get_path(ip_int, ip_dst, option):
+    if option == DEFAULT_OPT:
+        return nx.dijkstra_path(grafo, source=ip_int, target=ip_dst)
+    else:
+        return nx.dijkstra_path(grafo, source=ip_int, target=ip_dst)#, weight='weight')
 
-def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
+def add_path_through_gw(ip_int, ip_dst, option,isDpid=False):
     if isDpid:
         #TODO dpid to internet
 
@@ -223,16 +242,21 @@ def add_path_through_gw(ip_int, ip_ext, option,isDpid=False):
 
 
 def add_path(ip_src, ip_dst, option, isExt=False):
+    h=None
+    for hst in hosts:
+        if hst.ip == ip_src:
+            h=hst
+            break
     if isExt:
         newPath=get_path(ip_int,hosts[0].ip,option)
     else:
         newPath=get_path(ip_src,ip_dst,option)
 
[…]

### 2017-02-02 pox/ext/my_controller.py
trying to handle connection to external network

@@ -103,7 +103,7 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
             log.info("port %i is a 10Mbps",port.port_no)
         elif(port.curr & of.ofp_port_features_rev_map["OFPPF_100MB_HD"] or
             port.curr & of.ofp_port_features_rev_map["OFPPF_100MB_FD"]):
-            topo.switch[event.connection.dpid].port_capacity[port.port_no] = 100; #TODO constants
+            topo.switch[event.connection.dpid].port_cpacity[port.port_no] = 100; #TODO constants
             log.debug("port %i has capacity %i", port.port_no, topo.switch[event.connection.dpid].port_capacity[port.port_no])
             try:
                 topo.link_capacity(event.connection.dpid,topo.switch[event.connection.dpid].port_dpid[port.port_no],100)
@@ -138,8 +138,8 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
                 if topo.switch[dpid].dpid == event.connection.dpid:
                     sw = topo.switch[dpid]
                     break
-            sw.port_dpid[port.port_no] = IPAddr("0.0.0.0",0)
-            sw.dpid_port[IPAddr("0.0.0.0",0)]=port.port_no
+            sw.port_dpid[port.port_no] = IPAddr("",0)
+            sw.dpid_port[IPAddr("192.168.10.254",24)]=port.port_no
             #probably useless
             sw.port_mac[port.port_no]=port.hw_addr
             sw.mac_port[port.hw_addr]=port.port_no
@@ -229,6 +229,11 @@ def _handle_ip_packet(event):
     #             src_mac=dst_mac
     #             dst_mac=a
 
+    if type(EXTERNAL) is tuple and event.dpid == EXTERNAL[0] and event.port == EXTERNAL[1]:
+        host[0].mac=src_mac
+        if ip_src.inNetwork(SDN_network[0],SDN_network[1]):
+            host[0].ip=ip_src
+
     for i,pkt in enumerate(recent_packets):
         if pkt[0] == ip_src and pkt[1]==ip_dst and pkt[2] == event.dpid:
             return # it's a flooded packet that reached controller
@@ -281,7 +286,7 @@ def _handle_ip_packet(event):
             log.debug ("ip src and ip dst are logged. now add_path")
             topo.add_path(ip_src,ip_dst,topo.LOAD_OPT)
         elif hasNetwork and not ip_dst.inNetwork(SDN_network):
[…]

### 2017-02-02 pox/ext/my_topo_SDN.py

@@ -82,19 +82,35 @@ def add_switch(dpid):
     """
     Add a switch if it wasn't already added
     """
+    # global grafo
+    # global pck_error_min_gf
+    # global pck_error_max_gf
+    # global capacity_gf
+    # global delay_gf
+    # global switch_gf
+    # global load_gf
     if switch.has_key(dpid):
         pass
     else:
         switch[dpid] = my_Switch(dpid)
         grafo.add_node(dpid)
         pck_error_min_gf.add_node(dpid)
+        pck_error_max_gf.add_node(dpid)
         delay_gf.add_node(dpid)
         capacity_gf.add_node(dpid)
         load_gf.add_node(dpid)
         switch_gf.add_node(dpid)
+
         log.debug("Add switch: %s", dpid_to_str(dpid))
         add_default_rules(dpid)
 def rm_switch(dpid):
+    # global grafo
+    # global pck_error_min_gf
+    # global pck_error_max_gf
+    # global capacity_gf
+    # global delay_gf
+    # global switch_gf
+    # global load_gf
     grafo.remove_node(dpid)
     pck_error_min_gf.remove_node(dpid)
     pck_error_max_gf.remove_node(dpid)
@@ -105,6 +121,13 @@ def rm_switch(dpid):
     del switch[dpid] #remove the disconnected switch
 
 def save_graph():
+    # global grafo
+    # global pck_error_min_gf
+    # global pck_error_max_gf
+    # global capacity_gf
+    # global delay_gf
+    # global switch_gf
+    # global load_gf
     pos=nx.spring_layout(grafo) # positions for all nodes
     nx.draw_networkx(grafo,pos, with_labels=True, node_size=700, width=6, font_size=20,font_family='sans-serif')    #stampa anche il grafo
     plt.axis('off')
@@ -152,6 +175,13 @@ def add_link(dpid1, port1, dpid2, port2, isHost=False):
     """
     Inside add_link function, add switch and link
     """
+    # global grafo
+    # global pck_error_min_gf
+    # global pck_error_max_gf
+    # global capacity_gf
+    # global delay_gf
+    # global switch_gf
+    # global load_gf
     add_switch(dpid1)
     add_switch(dpid2)
     switch[dpid1].port_dpid[port1] = dpid2
@@ -169,6 +199,13 @@ def add_link(dpid1, port1, dpid2, port2, isHost=False):
 
 def rm_link(dpid1, port1, dpid2, port2):
     x = True
+    # global grafo
[…]

### 2017-02-02 pox/ext/stats_handler.py

@@ -207,7 +207,7 @@ def _setTraffic(flow_stat,dpid):
 
         dpid2=myTopo.switch[dpid].port_dpid[table["actions"]]
     #set the traffic load [0;1] 0= no traffic. 1= link is full
-        utilization = avgTrafPort/myTopo.switch[dpid].port_capacity[1]
+        utilization = avgTrafPort/myTopo.switch[dpid].port_capacity[table["actions"]]
         log.debug("Link utilization is %.4f",utilization)
         myTopo.link_load(dpid, dpid2, utilization)
         if(utilization>TRAFFIC_THRES): #byte for packet

### 2017-02-02 pox/ext/my_controller.py
minor modification

@@ -59,23 +59,25 @@ def _handle_LinkEvent(event):
     """
     l = event.link
     #if there's a link, the controller knows the link between switches
-    # for i,c in enumerate(unknown_link):
-    #     if c[0] == l.dpid1:
-    #         try:
-    #             dpid2 = switch[c[0]].port_dpid[c[1]]
-    #             topo.link_capacity(c[0],dpid2, c[2])
-    #             del unknown_link[i]# remove unknown link from the list
-    #         except:
-    #             break # doesn't have the other link yet
-    #     elif c[0] == l.dpid2:
-    #         try:
-    #             dpid = switch[c[0]].port_dpid[c[1]]
-    #             topo.link_capacity(c[0],dpid, c[2])
-    #             del unknown_link[i] # remove unknown link from the list
-    #         except:
-    #             break # doesn't have the other link yet
-    #     else:
-    #         continue
+    for i,c in enumerate(unknown_link):
+        if c[0] == l.dpid1:
+            try:
+                dpid2 = switch[c[0]].port_dpid[c[1]]
+                topo.link_capacity(c[0],dpid2, c[2])
+                topo.switch[dpid].port_capacity[c[1]] = c[2]
+                log.debug ("== ADDED CAPACITY BETWEEN %s AND %s : %i",c[0],dpid2,c[2])
+                del unknown_link[i]# remove unknown link from the list
+            except:
+                break # doesn't have the other link yet
+        elif c[0] == l.dpid2:
+            try:
+                dpid = switch[c[0]].port_dpid[c[1]]
+                topo.link_capacity(c[0],dpid, c[2])
+                del unknown_link[i] # remove unknown link from the list
+            except:
+                break # doesn't have the other link yet
+        else:
+            continue
     if event.added:
         log.debug('LinkAdd dpid1: {0} porta {1}, dpid2: {2} porta {3}'.format(l.dpid1, l.port1, l.dpid2, l.port2))
         topo.add_link(l.dpid1, l.port1, l.dpid2, l.port2)
@@ -98,6 +100,7 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
             log.debug("port %i has capacity %i", port.port_no, topo.switch[event.connection.dpid].port_capacity[port.port_no])
             try:
                 topo.link_capacity(event.connection.dpid,topo.switch[event.connection.dpid].port_dpid[port.port_no],10)
+                topo.switch[dpid].port_capacity[c[1]] = 10
             except:
[…]

### 2017-02-07 pox/ext/my_controller.py
minor correction. added default rule for the external packet

@@ -61,6 +61,12 @@ def _handle_LinkEvent(event):
     handle event ("LinkEvent") from openflow.discovery
     """
     l = event.link
+    if event.added:
+        log.debug('LinkAdd dpid1: {0} porta {1}, dpid2: {2} porta {3}'.format(l.dpid1, l.port1, l.dpid2, l.port2))
+        topo.add_link(l.dpid1, l.port1, l.dpid2, l.port2)
+    elif event.removed:
+        log.debug('LinkRemoved dpid1: {0} porta {1}, dpid2: {2} porta {3}'.format(l.dpid1, l.port1, l.dpid2, l.port2))
+        topo.rm_link(l.dpid1, l.port1, l.dpid2, l.port2)
     #if there's a link, the controller knows the link between switches
     for i,c in enumerate(unknown_link):
         if c[0] == l.dpid1:
@@ -81,12 +87,6 @@ def _handle_LinkEvent(event):
                 break # doesn't have the other link yet
         else:
             continue
-    if event.added:
-        log.debug('LinkAdd dpid1: {0} porta {1}, dpid2: {2} porta {3}'.format(l.dpid1, l.port1, l.dpid2, l.port2))
-        topo.add_link(l.dpid1, l.port1, l.dpid2, l.port2)
-    elif event.removed:
-        log.debug('LinkRemoved dpid1: {0} porta {1}, dpid2: {2} porta {3}'.format(l.dpid1, l.port1, l.dpid2, l.port2))
-        topo.rm_link(l.dpid1, l.port1, l.dpid2, l.port2)
 
 def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo switch legacy
     """
@@ -94,6 +94,12 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
     """
     topo.add_switch(event.connection.dpid)
     for port in event.ofp.ports:
+        topo.switch[event.connection.dpid].port_capacity[port.port_no]=10
+        try:
+            dpid2=topo.switch[event.connection.dpid].port_dpid[port.port_no]
+            topo.link_capacity(event.connection.dpid, dpid2 , 10)
+        except:
+            pass
         if ((port.state & of.ofp_port_state_rev_map["OFPPS_LINK_DOWN"]) and (port.port_no<10000)):
              log.info("port %i  is down",port.port_no)
         #check only current status. ignore maximum status. it's more realistic
[…]

### 2017-02-09 report/paper.tex
Added text on repo part. * ignored svg files * updated stats_handler, network_performance * added default_ext_rule to allow traffic going to internet. I forgot during the test session

@@
 
-\subsection{Traffic analysis}
+\subsection{Traffic analysis}\label{sec:traf}
 We used laptop connected via
@@
 
+\section{Statistics}
+The analysis on the traffic presented in sec \ref{sec:traf} was used to get useful information to the graph and allow
+a wise usage of the network. When the switch connects, the each port gives it's capabilities, so these has been used to
+find the utilization of the link and act a change on the path for some links. We choose to assign a bad link
+to nodes that connect with many other nodes and that generates a lot of traffic. When a link is found to be utilized more of half the
+capacity and other links are less utilized, the path of some traffic has been changed to equally distribute the traffic.
+For the p2p traffic, as shown in paper TODO ref!!** the torrent traffic must not be changed very often, as
+the protocol itself tryies to find different path to reach a destination. This type of traffic, instead, uses the path
+where the packet error rate, calculated with the port statistics, is higher. How the path error rate has been calculated
+can be found in \ref{sec:calculations}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Please, not that you do NOT have to describe the shape of the curves that can be seen in the figures, but the reasons WHY such curves have that shape!
+\subsection{The ARP packets}
+During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
+This type of packets is useful for a multipoint to multipoint connection where the connection is set between a
+HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
+hiher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
+act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
+If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
+send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
+the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
+For an IP network, it will check the path to reach a destination and then send to the defined output port. The mac address in this case
+became useless, as the connection is point to point. So, knowing the destination layer2 address is useless as the switch can act
+regardless of it. The only thing the switch need to know is the layer2, layer 3 addresses and, at most, the layer 4 address.
+This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
+this is necessary. TODO explain why TODO!!!!**
+
+\subsection{The Gateway}
+The computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
+properly, according to the rules set by the first. The controller must only know which switch ports are connected to the
+internet and then add them to the path. in this way, the controller can set a path to reach it and provide connectivity with
+a traffic shaping policy and/or differentiate the traffic from and to different networks.
 

### 2017-02-09 pox/ext/my_topo_SDN.py

@@ -40,6 +40,7 @@ LOAD_OPT      = 5
 
 TCP = 1 # same as Host
 UDP = 2
+TRANSP_BOTH = 0
 
 def add_host(dpid, mac, port, ip):
     """
@@ -128,43 +129,43 @@ def save_graph(counter):
     nx.draw_networkx(grafo,pos, with_labels=True, node_size=700, width=6, font_size=20,font_family='sans-serif')    #stampa anche il grafo
     plt.axis('off')
 
-    plt.savefig("grafo%i.png" % (counter))      #salva l'immagine
+    plt.savefig("grafo%i.svg" % (counter))      #salva l'immagine
     plt.clf()                     #elimina l'immagine corrente dalla libreria
 
     edge_labels=nx.draw_networkx_edge_labels(pck_error_min_gf,pos,font_size=12)
     nx.draw_networkx(pck_error_min_gf,pos, with_labels=True,node_color='green',node_size=700, width=6,font_size=20,font_family='sans-serif')    #stampa anche il grafo
     plt.axis('off')
-    plt.savefig("pck_error_min_gf%i.png" % (counter))   #salva l'immagine
+    plt.savefig("pck_error_min_gf%i.svg" % (counter))   #salva l'immagine
     plt.clf()                        #elimina l'immagine corrente dalla libreria
 
     edge_labels=nx.draw_networkx_edge_labels(pck_error_max_gf,pos,font_size=12)
     nx.draw_networkx(pck_error_max_gf,pos, with_labels=True,node_color='green',node_size=700, width=6,font_size=20,font_family='sans-serif')    #stampa anche il grafo
     plt.axis('off')
-    plt.savefig("pck_error_max_gf%i.png" % (counter))   #salva l'immagine
+    plt.savefig("pck_error_max_gf%i.svg" % (counter))   #salva l'immagine
     plt.clf()                        #elimina l'immagine corrente dalla libreria
 
     edge_labels2=nx.draw_networkx_edge_labels(delay_gf,pos,font_size=12)
     nx.draw_networkx(delay_gf,pos, with_labels=True,node_color='blue',node_size=700, width=6,font_size=20,font_family='sans-serif')    #stampa anche il grafo
     plt.axis('off')
-    plt.savefig("delay_gf%i.png" % (counter))      #salva l'immagine
+    plt.savefig("delay_gf%i.svg" % (counter))      #salva l'immagine
     plt.clf()                        #elimina l'immagine corrente dalla libreria
 
[…]

### 2017-02-09 pox/ext/network_performance.py

@@ -38,7 +38,7 @@ def send_msg(dpid):
     lista[dpid].add_rtt(identity)
 
 
-def create_msg(ipsrc = "100.100.100.0", ipdst = "100.100.100.1", mac_src = "80:80:80:80:80:80", mac_dst = "50:50:50:50:50:50", port = of.OFPP_CONTROLLER):
+def create_msg(ipsrc = "100.100.100.2", ipdst = "100.100.100.1", mac_src = "80:80:80:80:80:80", mac_dst = "50:50:50:50:50:50", port = of.OFPP_CONTROLLER):
     ipv4_pck = pkt.ipv4()
     ipv4_pck.srcip = IPAddr(ipsrc)
     ipv4_pck.dstip = IPAddr(ipdst)
@@ -77,12 +77,12 @@ def _handle_PacketIn(event):
         ip_packet = packet.find('ipv4')
         ip_dst = ip_packet.dstip #ip destinatario
 
-        if ipaddress.IPv4Address(ip_packet.srcip) == ipaddress.IPv4Address('100.100.100.0'):
+        if IPAddr(ip_packet.srcip) == IPAddr('100.100.100.2'):
             #sicuramente RTT msg
             lista[event.dpid].update_rtt(ip_dst, time)
             lista[event.dpid].calcolate_rtt(ip_dst, event.dpid)
             #log.debug("\n\ndpid= %s rtt %s", event.dpid, lista[event.dpid].av_rtt)
-        elif ipaddress.IPv4Address(ip_packet.srcip) == ipaddress.IPv4Address('100.100.100.1'):
+        elif IPAddr(ip_packet.srcip) == IPAddr('100.100.100.1'):
             #sicuramente delay msg
             dpid2 = topo.switch[event.dpid].port_dpid[event.port]
             rtt1 = -1

### 2017-02-09 pox/ext/stats_handler.py

@@ -147,7 +147,7 @@ def _setPktLoss(stat,dpid):
         if(total == 0):
             pErrRate.append(0)
         else:
-            pErrRate.append(errors/total)
+            pErrRate.append(100*errors/total)
     #get the link connection for a switch/port and update the weight
     for i,PER in enumerate(pErrRate):
         try: #there isn't a link yet
@@ -201,13 +201,17 @@ def _setTraffic(flow_stat,dpid):
         if table["actions"]>10000:
             continue
         log.debug("*** p: %i bc: %i  ts: %i",table["actions"],table["byteCount"],table["Tsecond"])
-        avgTrafPort += 8*10^9*table["byteCount"]/table["Tnanos"] # need a table for each rule
-        log.debug("average pkt size %.4f",avgTrafPort)
+        if(table["Tnanos"]==0):
+            continue
+        else:
+            avgTrafPort += 8*10^9*table["byteCount"]/table["Tnanos"] # need a table for each rule
+            log.debug("average pkt size %.4f",100*avgTrafPort)
     #for port,avg in avgTrafPort:
 
         dpid2=myTopo.switch[dpid].port_dpid[table["actions"]]
     #set the traffic load [0;1] 0= no traffic. 1= link is full
-        utilization = avgTrafPort/myTopo.switch[dpid].port_capacity[table["actions"]]
+
+        utilization = 100*avgTrafPort/myTopo.switch[dpid].port_capacity[table["actions"]]
         log.debug("Link utilization is %.4f",utilization)
         myTopo.link_load(dpid, dpid2, utilization)
         if(utilization>TRAFFIC_THRES): #byte for packet

### 2017-02-16 pox/ext/my_controller.py
sdn network now goes to internet successfully. tested in lab

@@ -145,7 +145,7 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
         #Find the port where is connected the standard router
         if hasGateway  and port.hw_addr == topo.hosts[0].mac: #is the switch with route to internet
             log.info("found port to the internet")
-            EXTERNAL = (event.dpid,port.port_no)
+            EXTERNAL = (event.connection.dpid,port.port_no)
             topo.hosts[0].switch=EXTERNAL
             topo.hosts[0].mac=EthAddr("FF:FF:FF:FF:FF:FF") # broadcast mac
             sw=None
@@ -153,16 +153,28 @@ def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo swi
                 if topo.switch[dpid].dpid == event.connection.dpid:
                     sw = topo.switch[dpid]
                     break
-            sw.port_dpid[port.port_no] = IPAddr("192.168.10.254",24)
-            sw.dpid_port[IPAddr("192.168.10.254",24)]=port.port_no
+            sw.port_dpid[port.port_no] = IPAddr("192.168.11.8",24)
+            sw.dpid_port[IPAddr("192.168.11.8",24)]=port.port_no
             #probably useless
             sw.port_mac[port.port_no]=port.hw_addr
             sw.mac_port[port.hw_addr]=port.port_no
+
+            # msg = of.ofp_flow_mod()
+            # msg.match.dl_type = 0x800
+            # msg.priority = topo.DEFAULT_EXT_NET_RULE
+            # msg.actions.append(of.ofp_action_output(port=port.port_no))
+            # core.openflow.sendToDPID(msg,event.connection.dpid)
+            print EXTERNAL
             msg = of.ofp_flow_mod()
-            msg.match.dl_type = 0x800
-            msg.priority = topo.DEFAULT_EXT_NET_RULE
-            msg.actions.append(of.ofp_action_output(port=port.port_no))
-            core.openflow.sendToDPID(msg,event.connection.dpid)
+            msg.priority = 42
+            msg.match.dl_type = 0x800 #ip
+            act = []
+            act.append(of.ofp_action_output(port=port.port_no))
+            act.append(of.ofp_action_output(port=of.OFPP_CONTROLLER))
+            msg.actions = act
+            core.openflow.sendToDPID(event.connection.dpid, msg) #switch i-esimo
+
+
             #add the link to the external
[…]

### 2017-02-16 pox/ext/my_topo_SDN.py

@@ -284,7 +284,7 @@ def get_gf(option):
 
 def get_path(ip_int, ip_dst, option):
     log.debug("get_path: ip_src=%s  ip_dst=%s",ip_int,ip_dst)
-    print get_gf(option).edges()
+    # print get_gf(option).edges()
     if option == DEFAULT_OPT:
         return nx.dijkstra_path(get_gf(option), source=ip_int, target=ip_dst)
     else:

### 2017-02-16 pox/ext/stats_handler.py

@@ -151,9 +151,10 @@ def _setPktLoss(stat,dpid):
     #get the link connection for a switch/port and update the weight
     for i,PER in enumerate(pErrRate):
         try: #there isn't a link yet
-            dpid2=myTopo.switch[dpid].port_dpid[i+1] # get the dpid connected to that port
+            dpid2=myTopo.switch[dpid].port_dpid[i] # get the dpid connected to that port
             myTopo.link_pck_err(dpid,dpid2,PER) # Update the packet error rate
         except:
+            log.debug("no link on port i")
             pass;
 
 # def _setLinkLoad(stat,dpid):
@@ -200,22 +201,24 @@ def _setTraffic(flow_stat,dpid):
     for table in flow_stat: # fill the throughput of a rule for that port
         if table["actions"]>10000:
             continue
-        log.debug("*** p: %i bc: %i  ts: %i",table["actions"],table["byteCount"],table["Tsecond"])
+        # log.debug("*** p: %i bc: %i  ts: %i",table["actions"],table["byteCount"],table["Tsecond"])
         if(table["Tnanos"]==0):
             continue
         else:
             avgTrafPort += 8*10^9*table["byteCount"]/table["Tnanos"] # need a table for each rule
-            log.debug("average pkt size %.4f",100*avgTrafPort)
+            # log.debug("average pkt size %.4f",100*avgTrafPort)
     #for port,avg in avgTrafPort:
 
         dpid2=myTopo.switch[dpid].port_dpid[table["actions"]]
     #set the traffic load [0;1] 0= no traffic. 1= link is full
-
-        utilization = 100*avgTrafPort/myTopo.switch[dpid].port_capacity[table["actions"]]
-        log.debug("Link utilization is %.4f",utilization)
-        myTopo.link_load(dpid, dpid2, utilization)
-        if(utilization>TRAFFIC_THRES): #byte for packet
-            log.debug("on dpid %s there's an host who is congestioning. Throughput/Capacity=%.3f",dpid_to_str(dpid),utilization)
+        try:
+            utilization = 100*avgTrafPort/myTopo.switch[dpid].port_capacity[table["actions"]]
+            # log.debug("Link utilization is %.4f",utilization)
+            myTopo.link_load(dpid, dpid2, utilization)
+        except:
+            utilization = 0.5
+        # if(utilization>TRAFFIC_THRES): #byte for packet
[…]

### 2017-02-16 report/paper.tex

@@
 we decided to focus on that. In this paper we will show how we handled the contemporary traffic between multiple host, both inside SDN network
-nd between the SDN one and the external internet. Starting from traffic analysis and going through the SDN state of the art, we implemented some function
+and between the SDN one and the external internet. Starting from traffic analysis and going through the SDN state of the art, we implemented some function
 to exploit better the SDN POX controller and we reach the end of this paper showing how we found three different types of traffic and
@@
 in the paper called "Getting traffic statistics from network devices in an
-SDN environment using OpenFlow" by D.J. Hamad & others. [ put note here]
+SDN environment using OpenFlow" by D.J. Hamad \& others. [ put note here]
 For traffic analysis we haven't found any paper describing the characteristics of
 some types of flows, so we used our PCs to sniff the traffic. Traffic analysis will be discussed in section \ref{sec:symo}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%parte scritta da Matteo da tradure e forse da spostare
+%\section{Controller Implementation}\label{sec:Implementation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%We developed a Pox based controller using Python2.
+%The main function that we need and that we had developed are:
+%\begin{itemize}
+% \item Gloal knowledg of our network topology with link performance metrics
+% \item Ensure host to host connectivity and external networ reachability using different performance path
+% \item Proces packetIn request
+%\end{itemize}
+
+%\subsection{Network topology}
+%To storage our network topology we use a specific python library called \emph{NetworkX} this library 
+%\url{https://networkx.github.io/}
+%give us the possibility
+%to create a graph. We have create this graph with two types of nodes: Switch and Hosts, all of them connected with link.
+%For now we use a single-edge graph but a real application have to use double-edge graph in orther to measure both link's performances.
+%\newline There is a principal function implemented by Pox called dyscovery.py that give us the event regarding ( check ) host, switch
+%and link connection/disconnection so handling this event we can build our graph step-by-step.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The System model is a description of your operating assumptions with related motivation and justification
+(The System model is a description of your operating assumptions with related motivation and justification)
 
@@
 We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
-#Mb for the gaming traffic, #Mb for the VoIP traffic and 3GB for the torrent traffic.
+\#Mb for the gaming traffic, \#Mb for the VoIP traffic and 3GB for the torrent traffic.
 The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP)
@@
 On another test, we tried the VoIP while a download was ongoing, and we saw that the quality
-decreased to allow communication. The traffic is between the UDP port 80 ##TODO## TOCHECK ##
+decreased to allow communication. The traffic is between the UDP port 80 \#\#TODO\#\# TOCHECK \#\#
 
 The last traffic analysis we made was with a DOTA[explain what it is] web match.
-We saw that for that type of gaming session, the packet size is about # bytes and involves both TCP and UDP ports.
+We saw that for that type of gaming session, the packet size is about \# bytes and involves both TCP and UDP ports.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%da vedere come distribuire
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{matteo}
+
+\section{Program structure}
+
+On our controller in orther two offer principals function we have two storage two principal things: the logical topology of the SDN network and the 
+proprirty of every nodes.
+
+\subsection{Topology}
+E' necessario memorizzare la topologia della rete in una struttura dati che permetta la facile ricerca di cammini minimi tra due host. 
+Abbiamo utilizzato quindi una struttura a grafo semplice in cui i nodi interni sono gli switch e le foglie sono gli host della nostra rete.
+Nella realtà le connessioni logiche tra due switch possono prevedere più di un link fisico che può essere mono o bdirezionale.
+Nel nostro sistema abbiamo preferito rappresentare la connessione tra due switch con un unico edge bidirezionale, da cui poi è facilmente estendibile
+in vista di un' applicazione pratica.
+
+Per gestire il grafo viene utilizzata la libreria scritta in python NeworkX
+
+The scope of our structure is storage the logically connection of network's nodes in orther to quikly discovery the minimum path
+
+\subsection{Network knowledge}\label{sec:topo}
+Our controller program need to know SDN all network topology in orther to guided ( instradare ) the packet along a better path or along a path to ensure 
+a better ( funzionamento ) on SDN network.
+We use two type of date structur:
+\begin{itemize}
+ \item \emph{Switch and Node}, we create ad-hoc object to save the main information of every single Node and Switch. Inside switch opject there are many 
+ dictionary that ( permettono ) to find quikly port to switch association, ip to port association and other. [da sistemare con piu' precisione]
+ \item \emph{Logically connection} to have a complete knowledg of the topology we use graph structured, despitly ( diversamente ) reality in
+ whitch  double-edge used we use single-edge graph to avoid evil implementation difficult that esule our aim. The NetworkX python library give as method
+ to create a graph in witch every node is a pointer to real Host or Switch object and the link conceptually have various metrics attribute.
+ ``Conceptually'' because NetworkX library don't give us the possibility to save more than one single weight attribute for each link, so 
+ we need to mantain multy graph, each of every metrics used with the same topology atribute but different weigth on theyr links.
+\end{itemize}
+
 \end{document}

### 2017-03-04 report/paper.tex
v 2 paper da sistemare molto

--- 2017-02-09 report/paper.tex
@@
 \begin{abstract}
-In this paper we present an application of the Software Defined Networks, focusing on the integration of the actual network with the SDN one.
-We will show how it's possible to implement a QoS mechanism to keep the network up and to grant the users the possibility of using it outside
-a laboratory.
+Quality of Service(QoS) in existing network architectures is an ongoing problem.
+A lot of researcher and industries try to solve this problem with many solution but a lot of them are too 
+expensive other failed or are not implemented.
+Software Defined Networking(SDN) paradigm has emerged in response to limitations of traditional networking architectures.
+The main advantages of SDN are centralized global network view. This particular features have got attention of researchers to improve QoS.
+In questo report mostreremo quali sono stati i principali risutati raggiunti e i problemi incontrati nella realizzazione di 
+un applicativo SDN in grado di supportare un servizio di tipo QoS.
+Lo standard OpenFlow e il controller python POX largamente diffuso sono stati gli strumenti che hanno reso possibile la
+realizzazione di quest'esperienza pur mostrando i limiti dovuti all'utilizzo di una tecnologia nuova e in fase di ricerca.
+%translate
 \end{abstract}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}\label{sec:intro}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The \textit{\textbf{SDN}, Software Defined Network} is a new approach to create a network of computer and networks, which is based on the
-most famous OpenFlow protocol. The key concepts of this type of networks are the hierarchy, the flow control and the statistics.
-The hierarchy allows a network administrator to set up a controller which will choose how to make the network evolve setting up
-how the routing will switch the traffic and doing so, how the single switch will behave.
-The flow control is done by the controller that, running the algorithm the network administrator created, choose the path for a
-type of traffic that occurs between a source and a destinaion host.
-The most important and interesting thing about those type of networks is the statistics: every switching unit collects data from its
-interfaces (e.g. the number of packets going in or out through a specific port or interface). This allow the controller to instantly
-how the network is evolving and, if necessary, switch a traffic from congestioned links to other that, in that moment, are almost fully loaded.
-Exploiting the statistic capabilities is the key to manage wisely a network and, since there aren't many papers about these potentialities,
-we decided to focus on that. In this paper we will show how we handled the contemporary traffic between multiple host, both inside SDN network
-nd between the SDN one and the external internet. Starting from traffic analysis and going through the SDN state of the art, we implemented some function
-to exploit better the SDN POX controller and we reach the end of this paper showing how we found three different types of traffic and
-for each one, we provided different metrics and different path to optimize the load. We realized the controller using POX, a python-based controller
-and we tested it with the switches that were provided us in the Communication Networks Lab at DEI.
-\subsection{Known Limitations}
-SDN networks are known to be underperforming with low traffic, which means that if nodes tries to connect with each other
-to exchange little traffic, the network will have a higher delay and will perform worst than a
-traditional network. We tried to reduce this underperformance with the usage of the wildcards:
-Every rule needs at least the source ip or the destination ip. If one of them is in a network
-(e.g. the internet traffic is in the network 0.0.0.0/0 excluded 192.168.10.0/24 ),
-we can redirect all the traffic from an internal host to the internet with a predefined
-path. This means that only one PacketIn will be sent to the controller if a host is surfing the network.
+Software-defined networking (SDN) is way to manage networks that separates the control plane from the forwarding plane. 
+SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brains” of the network. 
+The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs. 
+One of the most well-known protocols used by SDN Controllers is OpenFlow, however, it isn’t the only SDN standard, but it is widely accepted.
+Centralized, programmable SDN environments can easily adjust to the rapidly changing needs of businesses. SDN can lower costs and limit
+wasteful provisioning, as well as provide flexibility and innovation for networks.
+\#TODO da completare 
+  \subsection{OpenFlow}
+  During 2006 thanks to M. Casado was developed Ethane a new network architecture on this conncept during 2011 was bord the Open Network Foundatin.
+  This is a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
+  The first version of OpenFlow is 1.0.0 release on 2009 but the last standardized version is v1.5.1 approved on march 2015 \cite{ONF_report}.
+  \#TODO da sistemare
+  \subsection{The state of the art}
+  \#TODO stato dell arte e paper di riferimento  matteo
 
-Another limitation is the no default routing from a source to a destination: when an host
-is plugged to a network of SDN switches, it won't know the destinatio. The only way it can
-discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
-slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
-the network disposition, but doesn't know each host how much traffic is doing. For that the administrator need to put a firewall.
-The last one isn't needed, as the controller can set a switch to drop all the packets from/to a destination.
+\section{Needs and Gools}\label{sec:obb} %obbiettivi ed esigenze controllare 
+The main gool of this experience is to obtain statistics on realtime traffic in order to reroute traffic based on QoS or other reasons. 
+%correggere
+  \subsection{Adaptability of the network}
+  Made Network Adaptable means poter dirottare il traffico in real time to ensura a lot of demands. For example we have to avoid link congestion, 
+  but we can prefer reroute some hosting traffic to peggiorare it's performance because of it's evil traffic.
+  To reach this gools we need to have a global knowledge on the network and we have to modify the routing table on every routerboard with a 
+  global scope.
+  The performance of a single connection and entire network can ensure various type of metrics.
+  In order to assicurare queste ci serve una struttura dati che si aggiorna costantemente nella quale immagazzinare le statistiche dei link 
+  in modo all'occorrenza di trovare il percordo migliore per ogni singolo flusso dati.
+  \#TODO MATTEO
+  \subsection{Exterior comunication}
+  \#TODO pitt
 
-\subsection{Openflow 1.0 messages}
-Openflow defines a number of messages between a controller and a switch. Now there will be presented some messages we exploited:
-\begin{itemize}
-	\item[Hello] This message is sent both from the controller and the switch who is looking for a controller after it
-	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
-	\item[FeatureReq] this is the message that the controller sends to a switch to know its abilities. The switch
-	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
-	With the featureRes the controller will know how many ports a switch has and their properties (e.g. if they are connected, link speed,...), the flow table capabilities and the supported statistics.
-	\item[PortStatus] this message is sent from a switch to the controller, to notify that some ports has changed their status.
-		It's useful to check if a port has someone connected
-	\item[StatsReq] The controller will send this message to request the statistics. The payload of the message contains the type of statistic the controller want to know.
-		The switch then will answer with the Stats reply, if it supports that type of stat. It will send a Bad_Request message if not.
-		\item[PacketIn] This message is sent from the switch to the controller whenever a packet arrives and it doesn't match
-		any rule. The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
-		a flow rule for packets like this using a FlowMod.
-		\item[PacketOut] This message is sent from the controller to forward a packet, which is the payload of the PacketOut message, to some
-		port(s).
-		\item[FlowMod] This message set, delete or modify a rule inside a switch. Is the message that allows routing.
-\end{itemize}
+\section{Instrument}\label{sec:instrument}
+\#TODO matteo
+  \subsection{Controller}
+  The controller that we use is POX, it's repository is at this link \url{https://github.com/noxrepo/pox}. 
+  POX is an opensource python controller, it is an evolution of NOX a controller developed on C language.
+  The documentation about the progect is present at this link \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
+  The only problem to use this controller is that it support only the first version of OpenFlow1.0.
 
-\begin{itemize}
-	\item What are we talking about: description of the addressed problem. - done
-\item Motivation: why the problem is important. done
-\item Novelty: how you contribute to advance the state of the art. done
-\item Results: summary of the main findings  done
-\end{itemize}
+  The controller run inside a VirtualMachine on our personal computer, offered by SDN hub \cite{VM}. This VM use UBUNTU like OS and version 2.7.6 of
+  Python that is sufficient for POX that's required version 2.7.
+  % \#TODO matteo da correggere e completare
+  \subsection{Router Board}
+  The router board that we used is MikroTik miniROUTER RouterBOARD 450G. This Router Board virtualize an OpenFlow switch that support 
+  OpenFlow v1.0. The main caraptheristics is that it has only 5 ethernet port this is a limitation because consider that one of this port it is 
+  used to connect conntroller and with other port we have made host and other routerboard connection.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}\label{sec:sota}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-A good introduction about the SDN can be the  "Simulation in an SDN network scenario using the
-POX Controller" where has been presented the SDN environment and the mininet simulation
-tool that creates a topology where to simulate the switches, the traffic between hosts and
-allow the controller to communicate with the switches. The work presented in
-"Efficient topology discovery in OpenFlow-based Software Defined Networks" by F. Pakzad [put cit]
-has been implemented to improve the controller speed and efficiency as it make it more reactive.
-The introduction to the statistics power in a SDN network using the OpenFlow protocol has been found
-in the paper called "Getting traffic statistics from network devices in an
-SDN environment using OpenFlow" by D.J. Hamad & others. [ put note here]
-For traffic analysis we haven't found any paper describing the characteristics of
-some types of flows, so we used our PCs to sniff the traffic. Traffic analysis will be discussed in section \ref{sec:symo}.
+  %\#TODO matteo ( parlare del firmware utilizzato nella routerboard) correggere e completare
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{System Model}\label{sec:symo}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The System model is a description of your operating assumptions with related motivation and justification
+\section{Results}\label{sec:results}
+\#TODO %qui non so se serve un'introduzione o meno
+  \subsection{Graph and Network structures}
+  In order to achieve the realtime adaptability and statistics storage we have develop ad-hoc structure.
+  We use a specific python library called NetworkX \cite{networkx} to create a graph in witch we can storage ad-hoc node that represent 
+  resplectively switch or host element.
+  This topological structure is updated realtime thanks to connection event and link discover offered by various library. 
+  ( mettere i vari eventi che sfrutto )
+  The link besides have many parameters in witch we can save network performance. To update every link parameter we have to creat ad hoc function 
+  because there isn't some function constructed yet. ( sistemare frase) .
+  This type of structure allow us to run Dijkstra algorithm that return the minimum path ( in base a vari parametri ) on a nlog n time ( verificare).
+  (aggiungere qualcosa e correggere)
+  \#TODO matteo
 
-We developed a controller that can handle a mesh network, which has loops, that avoid
-the traffic flooding. The controller stores the switch and the connection between them;
-knows which hosts are connected to each switch port, with the hypothesys that only one
-host is connected at each port, unless the port is the one connected to the internet.
-With the use of the statistics, the controller get the status of the network and, if necessary,
-tries to adjust the traffic and put in the link with the higher path loss the torrent traffic,
-make the connection between two gaming host the one with less delay and tries to redistribute all the traffic in a way
-such that all the nodes have an average throughput as lowest as possible.
+  \subsection{Traffics analisis}
+  For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
+  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed.
+  Il che ha portato a dover utilizzare stratagemmi.
+  \#TODO %riformulare
+    \subsubsection{Delay}
+    Per molte applicazioni odierne il delay e' un parametro importantissimo. 
+    La rete attuale puo' conoscere il delay di un link semplicemente inviando un pachetto ICMP di tipo TimeStamp ( verificare ) 
+    supportato dal protocollo openflow, questo permette di ottenere una precisione nell'ordine dei microsecondi ( verificare ).
 
-\subsection{Traffic analysis}\label{sec:traf}
-We used laptop connected via
-GigabitEthernet interface and Ethernet cable cat.5a to receive the traffic from the destination host.
-Our laptop has been connected via Wi-Fi to our access point via protocol 802.11n.
-No other traffic sources were active during the testing, so the router was handling only
-the traffic from laptop when sniffing traffic. The connection used was an ADSL2+ via copper.
-We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
-#Mb for the gaming traffic, #Mb for the VoIP traffic and 3GB for the torrent traffic.
-The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP)
+    \#TODO MATTEO
+    \subsubsection{Trougthput}
+    \#TODO andrea
+    \subsubsection{Link Capacity}
+    \#TODO andrea
+    \subsubsection{altre}
+    \#TODO andrea
 
-The traffic analysis of the torrent protocol shows that the average packet size is the MTU
-(i.e. 1500bytes) and the protocol tryies to exploit all the channel capabilities. The traffic
-is made via UDP packets with the port range in the ephimeral ports (2000_- 65535).
-Moreover, the node is linked to a lot of IPs with multiple connection: this means that
-if someone wants to find a p2p traffic, it can look to the statistics and see if the traffic is loading all the channel,
-to how many IPs, an host is connected to and if it uses UDP.
-For the VoIP we saw that the protocol uses UDP and the traffic, after a transition time, goes only between two hosts.
-The average packet size is the MTU and the protocol tryies to exploit all the channel for a better QoE.
-On another test, we tried the VoIP while a download was ongoing, and we saw that the quality
-decreased to allow communication. The traffic is between the UDP port 80 ##TODO## TOCHECK ##
+  \subsection{Route changing} %cambio rotte
+  \#TODO andrea
 
-The last traffic analysis we made was with a DOTA[explain what it is] web match.
-We saw that for that type of gaming session, the packet size is about # bytes and involves both TCP and UDP ports.
-
-\section{Statistics}
-The analysis on the traffic presented in sec \ref{sec:traf} was used to get useful information to the graph and allow
-a wise usage of the network. When the switch connects, the each port gives it's capabilities, so these has been used to
-find the utilization of the link and act a change on the path for some links. We choose to assign a bad link
-to nodes that connect with many other nodes and that generates a lot of traffic. When a link is found to be utilized more of half the
-capacity and other links are less utilized, the path of some traffic has been changed to equally distribute the traffic.
-For the p2p traffic, as shown in paper TODO ref!!** the torrent traffic must not be changed very often, as
-the protocol itself tryies to find different path to reach a destination. This type of traffic, instead, uses the path
-where the packet error rate, calculated with the port statistics, is higher. How the path error rate has been calculated
-can be found in \ref{sec:calculations}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}\label{sec:res}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The Results section contains a selection of the most relevant results with the explanation of their meaning.
-Please, not that you do NOT have to describe the shape of the curves that can be seen in the figures, but the reasons WHY such curves have that shape!
-\subsection{The ARP packets}
+  \subsection{ARP solution} %TODO questione dell'arp inutile
   During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
@@
   this is necessary. TODO explain why TODO!!!!**
+  \#TODO andrea
 
-\subsection{The Gateway}
+  \subsection{Default Gw problem} 
   The computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
@@
   a traffic shaping policy and/or differentiate the traffic from and to different networks. 
+  \#TODO andrea
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \subsection{Ip switching in pox}
+  \#TODO andrea
+
 \section{Conclusions}\label{sec:conclusion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
 
+\#TODO
+
+\begin{thebibliography}{99}
+	\bibitem{pox} POX \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}
+	\bibitem{openflow} OpenFlow \url{http://flowgrammable.org/sdn/openflow/}
+	\bibitem{pox_repo} POX's repository \url{https://github.com/noxrepo/pox}
+	\bibitem{VM} Virtual Machine \url{http://sdnhub.org/tutorials/pox/}
+	\bibitem{router_board} Router Board \url{https://www.senetic.it/product/RB450G}
+	\bibitem{netoworkx} NetworkX python's library \url{https://networkx.github.io/}
+	\bibitem{ONF} Open Networking Foundation \url{https://www.opennetworking.org/sdn-resources/technical-library}
+	\bibitem{ONF_report} v1.5.1 OpenFlow report \url{https://www.opennetworking.org/sdn-resources/technical-library}
+
+\end{thebibliography}
+
 \end{document}

### 2017-03-04 pox/ext/my_controller.py

@@ -39,9 +39,7 @@ import my_topo_SDN as topo #new class
 import graphUpdater as gu
 log = core.getLogger()
 
-# SDN_network = "10.0.0.0/24"
-# SDN_NETMASK = "255.255.255.0"
-SDN_network = ""#"192.168.0.0/16"
+SDN_network = ""
 
 PCK_ERROR_OPT = 1
 DELAY_OPT     = 2
@@ -56,15 +54,14 @@ recent_packets = []
 counter = None
 
 def _addCapacity():
-    # print unknown_link
     if len(unknown_link)<1:
         return;
     for l in topo.grafo.edges():
         for i,c in enumerate(unknown_link):
-            log.debug("going on unknown_link")
+            #log.debug("going on unknown_link")
             if c[0] == l[0]:
                 try:
-                    log.debug("inside first try")
+                    #log.debug("inside first try")
                     dpid2 = switch[c[0]].port_dpid[c[1]]
                     topo.link_capacity(c[0],dpid2, c[2])
                     topo.switch[dpid].port_capacity[c[1]] = c[2]
@@ -74,7 +71,7 @@ def _addCapacity():
                     break # doesn't have the other link yet
             elif c[0] == l[1]:
                 try:
-                    log.debug("inside second try")
+                    #log.debug("inside second try")
                     dpid2 = switch[c[0]].port_dpid[c[1]]
                     topo.link_capacity(c[0],dpid2, c[2])
                     topo.switch[dpid].port_capacity[c[1]] = c[2]
@@ -84,8 +81,6 @@ def _addCapacity():
                     break # doesn't have the other link yet
             else:
                 continue
-
-
 
 def _handle_LinkEvent(event):
     """
@@ -98,9 +93,8 @@ def _handle_LinkEvent(event):
     elif event.removed:
         log.debug('LinkRemoved dpid1: {0} porta {1}, dpid2: {2} porta {3}'.format(l.dpid1, l.port1, l.dpid2, l.port2))
         topo.rm_link(l.dpid1, l.port1, l.dpid2, l.port2)
-    #if there's a link, the controller knows the link between switches
 
-def _handle_ConnectionUp (event): #capire se nella pratica si logga anche lo switch legacy
+def _handle_ConnectionUp (event):
     """
     handle connection up from switch
     """

### 2017-03-04 pox/ext/my_topo_SDN.py

@@ -84,13 +84,6 @@ def add_switch(dpid):
         log.debug("Add switch: %s", dpid_to_str(dpid))
         add_default_rules(dpid)
 def rm_switch(dpid):
-    # global grafo
-    # global pck_error_min_gf
-    # global pck_error_max_gf
-    # global capacity_gf
-    # global delay_gf
-    # global switch_gf
-    # global load_gf
     grafo.remove_node(dpid)
     pck_error_min_gf.remove_node(dpid)
     pck_error_max_gf.remove_node(dpid)
@@ -203,9 +196,7 @@ def link_delay(dpid1, dpid2, value):
     """
     modifica il peso del link del grafo delay_gf
     """
-    # global delay_gf
     delay_gf[dpid1][dpid2]['weight']=value
-    #log.debug("update delay with weight %.2f",delay_gf[dpid1][dpid2]['weight'])
 
 def link_pck_error(dpid1, dpid2, value):
     """

### 2017-03-04 pox/ext/network_performance.py

@@ -17,10 +17,10 @@ log = core.getLogger()
 lista = {} #lista dpid rtt
 
 def init_lista():
+    """ insert my_topo's dpid on lista """
     ls = topo.switch.keys() #lista chiavi dpid della my topo
     for i in range (0, len(ls)):
         lista[ls[i]] = my_rtt(ls[i]) #assegno un oggetto rtt ad ogni dpid
-    #log.debug(lista)
 
 def get_rtt():
     """deve aggiornare tutti gli rtt"""
@@ -36,8 +36,6 @@ def send_msg(dpid):
     msg = create_msg(ipdst = ip_dst)
     core.openflow.sendToDPID(dpid, msg)
     lista[dpid].add_rtt(identity)
-
-    #ciaop pino
 
 def create_msg(ipsrc = "100.100.100.2", ipdst = "100.100.100.1", mac_src = "80:80:80:80:80:80", mac_dst = "50:50:50:50:50:50", port = of.OFPP_CONTROLLER):
     ipv4_pck = pkt.ipv4()
@@ -63,7 +61,7 @@ def delay_discovery():
 
 def get_delay(dpid1, dpid2):
     out_port = topo.switch[dpid1].dpid_port[dpid2] #porta del dpid1 a cui e' collegato il dpid2
-    ip_dst = "100.%d.%d.%d" % (random.randint(0,250),random.randint(0,250), random.randint(0,250)) #docro' crearlo univoco
+    ip_dst = "100.%d.%d.%d" % (random.randint(0,250),random.randint(0,250), random.randint(0,250))
     identity = IPAddr(ip_dst)
     msg = create_msg(ipsrc = "100.100.100.1", ipdst = ip_dst, port = out_port)
     core.openflow.sendToDPID(dpid1, msg)
@@ -98,9 +96,14 @@ def _handle_PacketIn(event):
                 rtt2 = lista[dpid2].av_rtt
             except:
                 pass
-            if ((rtt1 != -1) and (rtt2 != -1)):
-                delay = lista[event.dpid].update_delay(ip_dst, time) - rtt1/2 -rtt2/2
-                topo.link_delay(event.dpid, dpid2 , delay)
+            if ((rtt1 != -1) and (rtt2 != -1) and (rtt1 < 10) and (rtt2 < 10) ):
+                delay = lista[event.dpid].update_delay(ip_dst, time)
[…]

### 2017-03-04 pox/ext/stats_handler.py

@@ -7,8 +7,6 @@ import my_topo_SDN as myTopo
 from pox.lib.addresses import IPAddr
 log = core.getLogger()
 
-# from enum import Enum
-# from graphUpdater import GraphUpdater
 UPD_GRAPH = 10 # every 10 seconds update the graph weight
 GAMING_THRES = 200 #bytes per packet
 TRAFFIC_THRES = 0.8 #

### 2017-03-04 report/paper.tex
create image folder

@@
 \section{Needs and Gools}\label{sec:obb} %obbiettivi ed esigenze controllare 
-The main gool of this experience is to obtain statistics on realtime traffic in order to reroute traffic based on QoS or other reasons. 
+The main gool of this experience is to develop a small SDN network in a lab with real hardware
+in orther to understood come a livello pratico si possa implementare  a QoS service in a SDN scenarois.
+Our network have to manage traffic with a global scope, in order to evolve in base on different parameters.
+Inoltre the SDN network need to interact with world wide web efficently.
 %correggere
   \subsection{Adaptability of the network}
-  Made Network Adaptable means poter dirottare il traffico in real time to ensura a lot of demands. For example we have to avoid link congestion, 
+  Made Network Adaptable means poter dirottare il traffico in real time to ensure a lot of demands. For example we have to avoid link congestion, 
   but we can prefer reroute some hosting traffic to peggiorare it's performance because of it's evil traffic.
@@
   \subsection{Controller}
-  The controller that we use is POX, it's repository is at this link \url{https://github.com/noxrepo/pox}. 
-  POX is an opensource python controller, it is an evolution of NOX a controller developed on C language.
-  The documentation about the progect is present at this link \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
+  The controller that we use is POX, the software is opensource and it is write with Python, it is based on SDN and support OpenFlow.
+  The software use the version 2.7 of Python.
+  It's repository is at this link \url{https://github.com/noxrepo/pox} and a lot of documentation about the progect 
+  is present at this link \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
   The only problem to use this controller is that it support only the first version of OpenFlow1.0.
-
-  The controller run inside a VirtualMachine on our personal computer, offered by SDN hub \cite{VM}. This VM use UBUNTU like OS and version 2.7.6 of
-  Python that is sufficient for POX that's required version 2.7.
+  \\
+  The controller run inside a VirtualMachine on our personal computer, 
+  the VM is offered by SDN hub on this site: \url{http://sdnhub.org/tutorials/sdn-tutorial-vm/}. 
+  This VM is a 64-bit Ubuntu 14.04 that has a number of SDN software and tools installed.
   % \#TODO matteo da correggere e completare
@@
   The router board that we used is MikroTik miniROUTER RouterBOARD 450G. This Router Board virtualize an OpenFlow switch that support 
-  OpenFlow v1.0. The main caraptheristics is that it has only 5 ethernet port this is a limitation because consider that one of this port it is 
+  OpenFlow v1.0\cite{routerboard_doc}. 
+  \begin{itemize}
+   \item CPU		AR7161 680MHz
+   \item Memory		256MB DDR SDRAM onboard memory
+   \item Data storage	512MB onboard NAND memory chip, microSD card slot (on reverse)
+   \item Ethernet 	Three 10/100/1000 Mbit/s Gigabit Ethernet ports supporting Auto-MDI/X
+   \item Serial port 	One DB9 RS232C asynchronous serial port
+  \end{itemize}
+  \textbf{firmware} the current version is: '' da gercare col comando: system routerboard print ''
+  \newline
+  The \textbf{Operating System} that they use is MikroTik RouterOS ( cercare la versione installata) \cite{routeboard_software}.
+  \newline
+  One limitation of this routerboard is that it has only 5 ethernet port this is a limitation because consider that one of this port it is 
   used to connect conntroller and with other port we have made host and other routerboard connection.
 
-  %\#TODO matteo ( parlare del firmware utilizzato nella routerboard) correggere e completare
+  %correggere e completare Matteo
 
@@
 	\bibitem{ONF_report} v1.5.1 OpenFlow report \url{https://www.opennetworking.org/sdn-resources/technical-library}
-
+	\bibitem{routerboard_doc} RouterBOARD 450G MikroTik documentation \url{http://www.routerboard.sk/files/pdf/rb450g_manual.pdf}
+	\bibitem{routeboard_software} RouterBOARD MikroTik Software \url{https://mikrotik.com/software}
 \end{thebibliography}

### 2017-03-05 report/paper.tex
Continued report

@@
     \#TODO MATTEO
-    \subsubsection{Trougthput}
-    \#TODO andrea
 		\subsubsection{Link Capacity}
 		\#TODO andrea
+		The first data that was collected was the actual link capacity, which depends both on the cable used to connect
+		different devices and the effective capacity for each ethernet port linked by the cable itself. Everytime a device supporting
+		openflow connects to the controller, it gives to the latter information about its status and its capabilities. One of them is
+		the link capacity in three ways: the first is the \textit{supported} capacity, which is the phisical maximum capacity, the second is the
+		\textit{advised} capacity, which name describes itself and the latter is the \textit{current} one. We choose to save the last one,
+		as it gives the most useful about the link state. This information, moreover, is the one that non-SDN switches use to choose the \textit{metric}
+		parameter. The simulation with Mininet showed us that this type of information can be used, but the Routeboards didn't support it.
+		As a matter of fact, we weren't able to exploit this capacity in our network realization.
+    \subsubsection{Throughput}
+    \#TODO andrea
+		Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
+		We exploited the information given by the switches with the match structure: this means that we used the throughput for the TCP traffic,
+		the one for the UDP and for each of them we could know toward which IP the traffic is directed or is from. This was exploited with mininet
+		simulations, but not in the set up in the laboratory. With the latter, we used the statistics as an indicator of the bitrate of a link
+		and divided the host to categories based on it.
+
+		\subsubsection{Packet error rate}
+		The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
+		reliability, we divided the number of packets with erroneous symbols that couldn't be corrected with a FEC by the total number of packets that were sent,
+		getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
+		heavy traffic was directed to links with highest packet error rate.
+		\#TODO andrea
+
     \subsubsection{altre}
-    \#TODO andrea
+    \#TODO andrea ?!?!?
 
@@
   \#TODO andrea
+	For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
+	Dijkstra algorithm for a weighted graph gave different minimum path between two hosts. From the parameter we saved, we choose to change the
+	route to reach the destination and viceversa. The traffic follow the same path, in both the directions and depends only on the type of traffic:
+	we choose to set the minimum unweighted path if no statistics were available, the minimum throughput path for the standard traffic, the minimum
+	delay for the gaming one and the maximum packet error rate for the heavy traffic generated by a host.
+	The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
+	algortithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
+	load in the network will be heavy and last longer.
 
@@
   \#TODO andrea
+	One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
+	switches the source and the destination IP and MAC address with the source one. The problem was partially solved, checking out the
+	parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
+	destination address, we can switch it back; the result is that the PacketIn event now has correct informations.
 

### 2017-03-05 pox/ext/stats_handler.py

@@ -199,28 +199,26 @@ def _setTraffic(flow_stat,dpid):
     for table in flow_stat: # fill the throughput of a rule for that port
         if table["actions"]>10000:
             continue
-        # log.debug("*** p: %i bc: %i  ts: %i",table["actions"],table["byteCount"],table["Tsecond"])
         if(table["Tnanos"]==0):
             continue
         else:
             avgTrafPort += 8*10^9*table["byteCount"]/table["Tnanos"] # need a table for each rule
-            # log.debug("average pkt size %.4f",100*avgTrafPort)
-    #for port,avg in avgTrafPort:
-
-        dpid2=myTopo.switch[dpid].port_dpid[table["actions"]]
-    #set the traffic load [0;1] 0= no traffic. 1= link is full
-        try:
-            utilization = 100*avgTrafPort/myTopo.switch[dpid].port_capacity[table["actions"]]
-            # log.debug("Link utilization is %.4f",utilization)
-            myTopo.link_load(dpid, dpid2, utilization)
-        except:
-            utilization = 0.5
-        # if(utilization>TRAFFIC_THRES): #byte for packet
-            # log.debug("on dpid %s there's an host who is congestioning. Throughput/Capacity=%.3f",dpid_to_str(dpid),utilization)
-        myTopo.switch[dpid].host_traffic[table["actions"]] = True
-        for h in myTopo.hosts:
-            if h.ip == myTopo.switch[dpid].port_dpid[table["actions"]]:
-                h.traffic = True
+        if table["match"].dltype == 0x800 and table["match"].nw_proto == 0x06: #TCP:
+            continue; #I don't care about tcp traffic. UDP traffic is more important
+        elif table["match"].dltype == 0x800 and table["match"].nw_proto == 0x017: #UDP
+            dpid2=myTopo.switch[dpid].port_dpid[table["actions"]]
+            try:
+                utilization = avgTrafPort/myTopo.switch[dpid].port_capacity[table["actions"]]
+                # log.debug("Link utilization is %.4f",utilization)
+                myTopo.link_load(dpid, dpid2, utilization)
+            except:
+                utilization = 0.5
+            if(utilization>TRAFFIC_THRES): #byte for packet
+                # log.debug("on dpid %s there's an host who is congestioning. Throughput/Capacity=%.3f",dpid_to_str(dpid),utilization)
+                myTopo.switch[dpid].host_traffic[table["actions"]] = True
[…]

### 2017-03-05 report/paper.tex
corrected some parts written by matteo and added some other parts

@@
 
-\section{Needs and Gools}\label{sec:obb} %obbiettivi ed esigenze controllare
+\section{Needs and Goals}\label{sec:obb} %obbiettivi ed esigenze controllare
 The main gool of this experience is to develop a small SDN network in a lab with real hardware
-in orther to understood come a livello pratico si possa implementare  a QoS service in a SDN scenarois.
+in orther to understand in a practical level how to implement a \textit{Quality of Service, \textbf{QoS}} service in a SDN scenarios.
 Our network have to manage traffic with a global scope, in order to evolve in base on different parameters.
-Inoltre the SDN network need to interact with world wide web efficently.
+Moreover the SDN network need to interact with world wide web efficently.
 %correggere
@@
   Made Network Adaptable means poter dirottare il traffico in real time to ensure a lot of demands. For example we have to avoid link congestion,
-  but we can prefer reroute some hosting traffic to peggiorare it's performance because of it's evil traffic.
+  but we can prefer reroute some hosting traffic to worsen it's performance because  it is the traffic we don't want to encourage.
   To reach this gools we need to have a global knowledge on the network and we have to modify the routing table on every routerboard with a
@@
   The performance of a single connection and entire network can ensure various type of metrics.
-  In order to assicurare queste ci serve una struttura dati che si aggiorna costantemente nella quale immagazzinare le statistiche dei link
-  in modo all'occorrenza di trovare il percordo migliore per ogni singolo flusso dati.
+  In order to ensure the traffic shaping based on network informations, there is a need of a structure, automatically updatable, where the link statistics
+	can be saved to allow the best path for a data stream. The best metric will be described below.
+	%  assicurare queste ci serve una struttura dati che si aggiorna costantemente nella quale immagazzinare le statistiche dei link
+  % in modo all'occorrenza di trovare il percordo migliore per ogni singolo flusso dati.
   \#TODO MATTEO
@@
   \#TODO pitt
+	Our network needed to be able to communicate with the external internet to be useful for a demo which demonstrate the capabilities of a SDN network.
+	The first thing we needed to accomplish was to make the controller aware of what switch and which port of the switch is connected to the external internet:
+	to do so, we allowed our controller to be able to receive in-line parameters. The two parameters we needed the controller to be aware were the mac address of the
+	port of a switch that is connected to the link and the network address of our network, to let the controller know if a request from a host was meant to another host
+	within the network or the data stream was intended to the internet.
+	In this way the host connected to the inner network can surf the net regardless of which type of devices are switching and routing the traffic they are generating.
 
@@
 
-  \subsection{Traffics analisis}
+  \subsection{Traffics analisis}\label{subsec:traf}
   For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
@@
     \subsubsection{Delay}
+
 	  Per molte applicazioni odierne il delay e' un parametro importantissimo.
@@
 		\subsubsection{Link Capacity}
+
 		\#TODO andrea
@@
     \subsubsection{Throughput}
+
 	  \#TODO andrea
@@
 		\subsubsection{Packet error rate}
+
 		The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
@@
 		getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
-		heavy traffic was directed to links with highest packet error rate.
+		heavy traffic was directed to links with highest packet error rate.  As the switches exchanges a packet before recognizing most of the error, both because the
+		phisical error are corrected using blocks coding and other error are found using a \textit{Frame Check Sequence} like CRC, the byte error were considered worse as
+		those are indirect measures themself and the error derived from the calculations could grow up.
 		\#TODO andrea
@@
   \subsection{Route changing} %cambio rotte
+
   \#TODO andrea
@@
 	delay for the gaming one and the maximum packet error rate for the heavy traffic generated by a host.
+	Two threshold were chosen based on the traffic analysis described on Sec. \ref{subsec:traf} which are 200 bytes/s for the gaming traffic and one megabyte/s
+	when the node is considered in heavy traffic. Those two threshold were made as they are average values and considering that our network is made of few really fast link,
+	so on one side the lower limit is sufficient not to consider gaming traffic to be a normal traffic, wherease it's easy to reach the high level of traffic load between two machines.
 	The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
@@
   \subsection{ARP solution} %TODO questione dell'arp inutile
+
   During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
@@
   \subsection{Default Gw problem}
+
   The computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
@@
   \subsection{Ip switching in pox}
+
   \#TODO andrea

### 2017-03-06 report/paper.tex
merge with andrea

--- 2017-03-04 report/paper.tex
@@
   In order to assicurare queste ci serve una struttura dati che si aggiorna costantemente nella quale immagazzinare le statistiche dei link 
-  in modo all'occorrenza di trovare il percordo migliore per ogni singolo flusso dati.
+  in modo all'occorrenza di trovare il percorso migliore per ogni singolo flusso dati.
   \#TODO MATTEO
@@
 Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
-
+ehi
 \#TODO

### 2017-03-06 report/paper.tex

--- 2017-03-05 report/paper.tex
@@
   \#TODO da sistemare
-  \subsection{The state of the art}
+  \\
+  \\
+  \textbf{Openflow 1.0 messages}
+Openflow defines a number of messages between a controller and a switch. Now there will be presented some messages we exploited:\cite{openflow}
+\begin{itemize}
+
+\item \emph{Hello} \newline This message is sent both from the controller and the switch who is looking for a controller after it
+	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
+\item \emph{FeatureReq} \newline this is the message that the controller sends to a switch to know its abilities. The switch
+	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
+	With the featureRes the controller will know how many ports a switch has and their properties (e.g. if they are connected, link speed,...), the flow table capabilities and the supported statistics.
+\item \emph{PortStatus} \newline this message is sent from a switch to the controller, to notify that some ports has changed their status.
+		It's useful to check if a port has someone connected
+\item \emph{StatsReq} \newline The controller will send this message to request the statistics. The payload of the message contains the type of statistic the controller want to know.
+		The switch then will answer with the Stats reply, if it supports that type of stat. It will send a Bad_Request message if not.
+\item \emph{PacketIn} \newline This message is sent from the switch to the controller whenever a packet arrives and it doesn't match
+		any rule. The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
+		a flow rule for packets like this using a FlowMod.
+\item \emph{PacketOut} \newline This message is sent from the controller to forward a packet, which is the payload of the PacketOut message, to some
+		port(s).
+\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch. Is the message that allows routing.
+\\
+\\
+\end{itemize}
+
+\textbf{Forwarding table:}f
+\begin{itemize}
+ \item \texttt{match fields:} sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze. Contengono la porta d'ingresso, l'intestazione del pacchetto, 
+ e altri campi opzionali;
+ \item \texttt{priority:} precedenza nel caso di voci con pi\`u corrispondenze;
+ \item \texttt{counter:} incrementato ogni qualvolta che viene riscontrata la corrispondenza con un pacchetto;
+ \item \texttt{istructions:} istruzioni da aggiungere all'\emph{Action-Set} o modifiche al flusso della pipeline;
+ \item \texttt{timeouts:} massimo tempo per cui lo switch tiene in memoria questa voce, altre il quale la voce viene scartata;
+ \item \texttt{cookie:} valori utilizzati per statistiche riguardo la singola voce, quindi non vengono considerati nell'elaborazione dei pacchetti dati;
+ \item \texttt{flags:} utilizzate per differenziare la gestione della voce;
+ \\
+ \\
+\end{itemize}
+
+\textbf{Match Constraints}
+principal match method that we can used with OpenFlow v1.0:
+\begin{itemize}
+ \item \texttt{in_port}: the port of swith;
+ \item \texttt{dl_src/dl_dst}: 802.3 MAC source and destination of packet;
+ \item \texttt{dl_type}: 802.3 EtherType;
+ \item \texttt{nw_tos}: IPv4 ToS;
+ \item \texttt{nw_proto}: IPv4 Protocol;
+ \item \texttt{nw_src/nw_dst}: IPv4 Source/Destination;
+ \item \texttt{tp_src/tp_dst}: TCP/UDP Source/Destination Port;
+ \\
+ \\
+\end{itemize}
+
+\textbf{Action}
+With the OpenFlow v1.0 in practice we can do simply things:
+\begin{itemize}
+ \item \texttt{Output}: set output port;
+ \item \texttt{SetDLSrc/Dst}: modify packet MAC address;
+ \item \texttt{SetNWSrc/Dst}: modify packet IPv4 address;
+ \item \texttt{SetNWTos}
+ \item \texttt{SetTPSrc/Dst}
+\end{itemize}
+With from OpenFlow v1.1 in poi we can modify Time To Live and other things.
+
+  \subsection{Related Work}
+  The work presented in 
+"Efficient topology discovery in OpenFlow-based Software Defined Networks" by F. Pakzad \cite{farzaneh} 
+has been implemented to improve the controller speed and efficiency as it make it more reactive.
+The introduction to the statistics power in a SDN network using the OpenFlow protocol has been found 
+in the paper called "Getting traffic statistics from network devices in an SDN environment using OpenFlow" by D.J. Hamad \& others. [ put note here]
+For traffic analysis we haven't found any paper describing the characteristics of
+some types of flows, so we used our PCs to sniff the traffic. Traffic analysis will be discussed in section.
   \#TODO stato dell arte e paper di riferimento  matteo
@@
   The performance of a single connection and entire network can ensure various type of metrics.
-  In order to ensure the traffic shaping based on network informations, there is a need of a structure, automatically updatable, where the link statistics
-	can be saved to allow the best path for a data stream. The best metric will be described below.
+  In order to ensure the traffic shaping based on network informations, there is a need of a structure, 
+  automatically updatable, where the link statistics can be saved to allow the best path for a data stream. The best metric will be described below.
   %  assicurare queste ci serve una struttura dati che si aggiorna costantemente nella quale immagazzinare le statistiche dei link
@@
 
-    \subsubsection{altre}
-    \#TODO andrea ?!?!?
-
   \subsection{Route changing} %cambio rotte
@@
 	\bibitem{routeboard_software} RouterBOARD MikroTik Software \url{https://mikrotik.com/software}
+	\bibitem{farzaneh}  Efficient Topology Discovery in Software Defined Networks F. Pakzad, M. Portmann, W. L. Tan and J. Indulska 
+	School of ITEE, The University of Queensland Brisbane, Australia \url{https://www.researchgate.net/publication/271508903_Efficient_Topology_Discovery_in_OpenFlow-based_Software_Defined_Networks}
 \end{thebibliography}

### 2017-03-07 report/paper.tex
delay extension

@@
 
-\textbf{Forwarding table:}f
+\textbf{Forwarding table:}
 \begin{itemize}
@@
   \subsection{Router Board}
+  \begin{figure}[!h]
+ \centering
+ \includegraphics[scale=0.30]{images/rboard.png}
+ \caption{\emph{Experience scenario}}
+ \label{fig:topo}
+  \end{figure}
+
   The router board that we used is MikroTik miniROUTER RouterBOARD 450G. This Router Board virtualize an OpenFlow switch that support
@@
 \section{Results}\label{sec:results}
+\subsection{Real Lab Scenario}
+The topological scenario in which we have set up experience is the next:
+\begin{figure}[!h]
+ \centering
+ \includegraphics[scale=0.70]{images/topo.png}
+ \caption{\emph{Experience scenario}}
+ \label{fig:topo}
+\end{figure}
+
+We have used some principal function released with POX controller:
+\begin{itemize}
+ %\item \texttt{pox.topology.launch()} and \texttt{pox.openflow.topology.launch()} descrivere a che ci serve\cite{pox}.
+ \item \texttt{pox.openflow.discovery.launch()} this function run algorithm in order to discovery the network and every time in witch some new 
+ topological change occour it return an event with ( con cui ) we can update our topological knowledge.
+ It does not provide the host event. For host event\cite{pox}.
+ \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology 
+ because with a spanning_tree algorithm it avoid the broadcast storm of message. When we set up \texttt{OFPP_FLOOD} like output port action in a table's 
+ switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service 
+ on the switch in orther to avoid loop\cite{pox}.
+\end{itemize}
+
 \#TODO %qui non so se serve un'introduzione o meno
@@
   For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
-  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed.
+  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
   Il che ha portato a dover utilizzare stratagemmi.
@@
     \subsubsection{Delay}
+Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
+The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  (hosts  or  middleboxes).
+In traditional  networks, the control over the network routing is
+limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  sim-plifies 
+network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
+specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
+\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from 
+an other switch.
+\\
+\newline First we have to discovery round trip time from controller to every single switch on our topology.
+In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
+option and when the message return to the controller we stop the timer and calculare RTT for the switch.
+\begin{figure}[!h]
+ \centering
+ \includegraphics[scale=0.70]{images/rtt2.png}
+ \caption{\emph{RTT between controller and OF switch}}
+ \label{fig:topo}
+\end{figure}
+\newline
+If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
+\\
+\newline 
+Now we create a proper message with unique identifier number, we start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
+Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
+\newline This part required a particolar flow entry on the second switch in orther to send to the controller only some particulare packet.
+\begin{figure}[!h]
+ \centering
+ \includegraphics[scale=0.70]{images/rtt1.png}
+ \caption{\emph{Pck from controller to itself following a particular way.}}
+ \label{fig:topo}
+\end{figure}
+\\
+\newline At this point we can calculate the delay of the link.
+\\
+\\
+ delay = (TM2 - TM1) - RTT1/2 - RTT2/2
+\\
+\\
+Teorethically this method must works but in practice there are a lot of delay introduce from computational concestion on the virtualized controller.
+What we observe is a estimated delay like some seconds that is impossible, infact if we made some \texttt{ping test} from one host to another host 
+in this SDN network we realize delay around 3ms.
+\\
 
-	  Per molte applicazioni odierne il delay e' un parametro importantissimo.
-    La rete attuale puo' conoscere il delay di un link semplicemente inviando un pachetto ICMP di tipo TimeStamp ( verificare )
-    supportato dal protocollo openflow, questo permette di ottenere una precisione nell'ordine dei microsecondi ( verificare ).
-
-    \#TODO MATTEO
 		\subsubsection{Link Capacity}
@@
 		As a matter of fact, we weren't able to exploit this capacity in our network realization.
+		\\
+		\\
     \subsubsection{Throughput}
-
 	  \#TODO andrea
@@
 		and divided the host to categories based on it.
-
+\\
 		\subsubsection{Packet error rate}
@@
 		\#TODO andrea
-
+\\
   \subsection{Route changing} %cambio rotte
@@
 
-  \subsection{ARP solution} %TODO questione dell'arp inutile
+  \subsection{ARP solution}
 
@@
   This can be seen by a reader as a violation of the layers in the ISO/OSI stack, but to exploit the potentials of the SDN approach,
-  this is necessary. TODO explain why TODO!!!!**
-  \#TODO andrea
+  this is necessary.
 
@@
 	\bibitem{router_board} Router Board \url{https://www.senetic.it/product/RB450G}
-	\bibitem{netoworkx} NetworkX python's library \url{https://networkx.github.io/}
+	\bibitem{networkx} NetworkX python's library \url{https://networkx.github.io/}
 	\bibitem{ONF} Open Networking Foundation \url{https://www.opennetworking.org/sdn-resources/technical-library}
@@
 	School of ITEE, The University of Queensland Brisbane, Australia \url{https://www.researchgate.net/publication/271508903_Efficient_Topology_Discovery_in_OpenFlow-based_Software_Defined_Networks}
+	\bibitem{qos_paper} OpenNetMon: Network Monitoring in OpenFlow Software-Defined Networks Niels L.
+	M. van Adrichem, Christian Doerr and Fernando A. Kuipers Network Architectures and Services, Delft University of Technology 
+	Mekelweg 4, 2628 CD Delft, The Netherlands \url{https://pdfs.semanticscholar.org/ea94/7a9334484ee4c3bb6bcfb4d1590aece1b178.pdf}
+	\bibitem{rtt} Efficient Round-Trip Time Monitoring in OpenFlow Networks A. Atary A. Bremler-Barr Interdisciplinary Center Herzliya \url{http://www.deepness-lab.org/pubs/infocom2016_grami.pdf}
+
 	\end{thebibliography}

### 2017-03-07 report/paper.tex
altre aggiunte

@@
   \#TODO da sistemare
-  \\
+\begin{figure}[!h]
+ \centering
+ \includegraphics[scale=0.58]{images/of.jpg}
+ \caption{\emph{OpenFlow}}
+ \label{fig:topo}
+\end{figure}
   \\
@@
 
+\subsection{Known Limitations}
+SDN networks are known to be underperforming with low traffic, which means that if nodes tries to connect with each other
+to exchange little traffic, the network will have a higher delay and will perform worst than a
+traditional network. We tried to reduce this underperformance with the usage of the wildcards:
+Every rule needs at least the source ip or the destination ip. If one of them is in a network
+(e.g. the internet traffic is in the network 0.0.0.0/0 excluded 192.168.10.0/24 ),
+we can redirect all the traffic from an internal host to the internet with a predefined
+path. This means that only one PacketIn will be sent to the controller if a host is surfing the network.
+
+Another limitation is the no default routing from a source to a destination: when an host
+is plugged to a network of SDN switches, it won't know the destinatio. The only way it can
+discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
+slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
+the network disposition, but doesn't know each host how much traffic is doing. For that the administrator need to put a firewall.
+The last one isn't needed, as the controller can set a switch to drop all the packets from/to a destination.
+
 \subsection{Related Work}
@@
 \section{Instrument}\label{sec:instrument}
-\#TODO matteo
   \subsection{Controller}
@@
 \section{Results}\label{sec:results}
-\subsection{Real Lab Scenario}
-The topological scenario in which we have set up experience is the next:
-\begin{figure}[!h]
- \centering
- \includegraphics[scale=0.70]{images/topo.png}
- \caption{\emph{Experience scenario}}
- \label{fig:topo}
-\end{figure}
-
+\subsection{Controller:}
+We developed a controller that can handle a mesh network, which has loops, that avoid
+the traffic flooding. The controller stores the switch and the connection between them;
+knows which hosts are connected to each switch port, with the hypothesys that only one
+host is connected at each port, unless the port is the one connected to the internet.
+With the use of the statistics, the controller get the status of the network and, if necessary,
+tries to adjust the traffic and put in the link with the higher path loss the torrent traffic,
+make the connection between two gaming host the one with less delay and tries to redistribute all the traffic in a way
+such that all the nodes have an average throughput as lowest as possible.
+\\
+\newline
 We have used some principal function released with POX controller:
@@
 \end{itemize}
+\subsection{Real Lab Scenario}
+The topological scenario in which we have set up experience is the next:
+\begin{figure}[!h]
+ \centering
+ \includegraphics[scale=0.70]{images/topo.png}
+ \caption{\emph{Experience scenario}}
+ \label{fig:topo}
+\end{figure}
 
-\#TODO %qui non so se serve un'introduzione o meno
+We have use 4 MikroTik Router board that were provided us in the Communication Networks Lab at DEI in University of Padua.
+The 4 routerboard is connected each other in order to design a loop. Every sigle board connect two host and is connected with a legacy switch that provide 
+to logically connected with controller.
+
 \subsection{Graph and Network structures}
@@
   \subsection{Traffics analisis}\label{subsec:traf}
+  We used laptop connected via
+GigabitEthernet interface and Ethernet cable cat.5a to receive the traffic from the destination host.
+Our laptop has been connected via Wi-Fi to our access point via protocol 802.11n.
+No other traffic sources were active during the testing, so the router was handling only
+the traffic from laptop when sniffing traffic. The connection used was an ADSL2+ via copper.
+We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
+\#Mb for the gaming traffic, \#Mb for the VoIP traffic and 3GB for the torrent traffic.
+The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP)
+
+The traffic analysis of the torrent protocol shows that the average packet size is the MTU
+(i.e. 1500bytes) and the protocol tryies to exploit all the channel capabilities. The traffic
+is made via UDP packets with the port range in the ephimeral ports (2000_- 65535).
+Moreover, the node is linked to a lot of IPs with multiple connection: this means that
+if someone wants to find a p2p traffic, it can look to the statistics and see if the traffic is loading all the channel,
+to how many IPs, an host is connected to and if it uses UDP.
+For the VoIP we saw that the protocol uses UDP and the traffic, after a transition time, goes only between two hosts.
+The average packet size is the MTU and the protocol tryies to exploit all the channel for a better QoE.
+On another test, we tried the VoIP while a download was ongoing, and we saw that the quality
+decreased to allow communication. The traffic is between the UDP port 80 \#\#TODO\#\# TOCHECK \#\#
+
+The last traffic analysis we made was with a DOTA[explain what it is] web match.
+We saw that for that type of gaming session, the packet size is about \# bytes and involves both TCP and UDP ports.
+  \\
+  \\
   For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software

### 2017-03-11 report/paper.tex
Added content to paper

@@
 \usepackage{enumitem}
-
+\usepackage{fancyhdr}
+\fancyfoot[C]{\thepage}
 %tikz figures
@@
 
+\mainmatter
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Another limitation is the no default routing from a source to a destination: when an host
-is plugged to a network of SDN switches, it won't know the destinatio. The only way it can
+is plugged to a network of SDN switches, it won't know the destination. The only way it can
 discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
@@
 Moreover the SDN network need to interact with world wide web efficently.
+\subsection{Connectivity}\label{sec:connectivity}
+First thing to do is provide connectivity, letting the controller know the connection of a host and avoiding packet to be lost while the first
+add the computer to its graphes and try to add the rules to let the traffic go from the source to the destination. This has been done doing a
+\textit{controlled flooding}: when a PacketIn event arrives at the controller, this look into the data structure that saves the nodes connected.
+\\ If source and destination nodes are already found, it set the flow rules from the source node' switch to the destination node switch. To the latter,
+a PacketOut message is prepared with inside the packet arrived inside the PacketIn event and with destination the port of the destination switch connected
+to the destination node. Then the message is sent to the destination switch, so that the first packet is not lost.
+If, instead, the destination node is not found, the controller sends to the switch that generated the PacketIn event a PacketOut message
+with destination all the ports of the switch itself and with payload, the message that arrived within the PacketIn message.
+To avoid the network to be filled of the only packet sent, we kept the information regarding the switches on which the packet that was flooded:
+if a PacketIn message arrives at the controller and both the packet and the switch got a PacketOut message within 2 seconds, the controller doesn't
+do anything, as this is a loop.
+\\ So this approach allowed us to make the network unaware that it's not a legacy one and allowed connectivity to be provided.
 %correggere
@@
 \subsection{Exterior comunication}
-  \#TODO pitt
 Our network needed to be able to communicate with the external internet to be useful for a demo which demonstrate the capabilities of a SDN network.
 The first thing we needed to accomplish was to make the controller aware of what switch and which port of the switch is connected to the external internet:
-	to do so, we allowed our controller to be able to receive in-line parameters. The two parameters we needed the controller to be aware were the mac address of the
+to do so, we allowed our controller to be able to receive and process in-line parameters. We provided two parametes to the controller, so that
+it could recognize where the traffic to the internet need to be sent. Theese two parameters were the mac address of the
 port of a switch that is connected to the link and the network address of our network, to let the controller know if a request from a host was meant to another host
-	within the network or the data stream was intended to the internet.
-	In this way the host connected to the inner network can surf the net regardless of which type of devices are switching and routing the traffic they are generating.
+within the network or the data stream was intended to the internet. We provided only one gateway, as the Routeboard had few ports available for the packet switching
+and only four SDN-capable switches to be able to simulate a real network that was generating traffic.
 
@@
 \section{Results}\label{sec:results}
-\subsection{Controller:}
+\subsection{Controller}
 We developed a controller that can handle a mesh network, which has loops, that avoid
@@
 Dijkstra algorithm for a weighted graph gave different minimum path between two hosts. From the parameter we saved, we choose to change the
-	route to reach the destination and viceversa. The traffic follow the same path, in both the directions and depends only on the type of traffic:
+route to reach the destination and viceversa. The traffic follow the same path in both the directions and depends only on the type of traffic:
 we choose to set the minimum unweighted path if no statistics were available, the minimum throughput path for the standard traffic, the minimum
@@
 HUB and a set of computers or between a bridge and a computer. The first case is no more in use in common networks as
-  hiher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
+it higher the collision range. The bridges limits their collision range to the port. With openflow each switching unit can
 act as a bridge or as a router. In both cases analyses the header of the L2 packet L3 packet and some of the L4 header.
@@
 
-  \subsection{Default Gw problem}
-
-  The computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
+\subsection{Default Gateway}
+Analyzing the traffic while we were trying the controller on real machines in the lab, we have tried to set the default gateway to
+other machines that were connected to the SDN network, but not to the internet. We proved that the controller can redirect correctly the
+traffic to the internet. This means that
+computers don't need anymore to set a gateway to reach another network as the controller and the switches will redirect the traffic
 properly, according to the rules set by the first. The controller must only know which switch ports are connected to the
-  internet and then add them to the path. in this way, the controller can set a path to reach it and provide connectivity with
+internet and then add them to the path. I this way, the controller can set a path to reach it and provide connectivity with
 a traffic shaping policy and/or differentiate the traffic from and to different networks.
-  \#TODO andrea
 
@@
 
-  \#TODO andrea
-	One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
-	switches the source and the destination IP and MAC address with the source one. The problem was partially solved, checking out the
+During our tests, both with Mininet and in the lab, we found out that the core of POX, when a PacketIn event occoures, switches the source and the destination addresses,
+both MAC and IP, if one of the hosts is already known. We weren't able to understand in which point of the code of POX  and why it happens.
+% One of the problem encountered was that, if an host is already found by pox and there is a PacketIn event, the core of the controller
+% switches the source and the destination IP and MAC address with the source one.
+The problem was partially solved, checking out the
 parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
 destination address, we can switch it back; the result is that the PacketIn event now has correct informations.
+If the two host are in the same switch, the problem is solved using a controlled packet flooding, described in section \ref{sec:connectivity}
 

### 2017-03-11 report/paper.tex
niente fatto
first page translate and correct

@@
 
-\mainmatter
+%\mainmatter
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}
-Quality of Service(QoS) in existing network architectures is an ongoing problem.
-A lot of researcher and industries try to solve this problem with many solution but a lot of them are too
-expensive other failed or are not implemented.
-Software Defined Networking(SDN) paradigm has emerged in response to limitations of traditional networking architectures.
-The main advantages of SDN are centralized global network view. This particular features have got attention of researchers to improve QoS.
-In questo report mostreremo quali sono stati i principali risutati raggiunti e i problemi incontrati nella realizzazione di
-un applicativo SDN in grado di supportare un servizio di tipo QoS.
-Lo standard OpenFlow e il controller python POX largamente diffuso sono stati gli strumenti che hanno reso possibile la
-realizzazione di quest'esperienza pur mostrando i limiti dovuti all'utilizzo di una tecnologia nuova e in fase di ricerca.
-%translate
+Quality of Service (QoS) in existing network architectures is an ongoing problem.
+A lot of researchers and industries tried to solve this problem and found different solutions but some of them are too
+expensive other failed or were not implemented.
+Software Defined Networking (SDN) paradigm has emerged in response to traditional networking architecture's limitations.
+The main advantage of SDN is a centralized global network view. This particular feature pushed researchers to keep on improving QoS.
+
+In this report we will focus on two different aspects; all problems we faced during our attempt to create a based SDN application able to
+support a QoS service and the results we achieved.
+
+OpenFlow standard and python POX controller, which is largely spread, were used as instruments to make this experience possible,
+despite the limit of using a new type of tecnology still in the research phase.
+%corretto
 \end{abstract}
@@
 \section{Introduction}\label{sec:intro}
-Software-defined networking (SDN) is way to manage networks that separates the control plane from the forwarding plane.
-SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brains” of the network.
+Software-defined networking (SDN) is a way to manage networks that separates the control from the forwarding plane.
+SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brain” of the network.
 The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
-One of the most well-known protocols used by SDN Controllers is OpenFlow, however, it isn’t the only SDN standard, but it is widely accepted.
-Centralized, programmable SDN environments can easily adjust to the rapidly changing needs of businesses. SDN can lower costs and limit
+One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it isn’t the only SDN standard. 
+Centralized, programmable SDN environments can easily adjust to the rapidly changing business'needs. SDN can reduce costs and restrict
 wasteful provisioning, as well as provide flexibility and innovation for networks.
-\#TODO da completare
+%corretto
 \subsection{OpenFlow}
-  During 2006 thanks to M. Casado was developed Ethane a new network architecture on this conncept during 2011 was bord the Open Network Foundatin.
-  This is a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
-  The first version of OpenFlow is 1.0.0 release on 2009 but the last standardized version is v1.5.1 approved on march 2015 \cite{ONF_report}.
-  \#TODO da sistemare
+  During 2006 ,thanks to M. Casado, a new network architecture was developed, called Ethane.
+  On this conncept was born, during 2011, the Open Network Foundation, a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
+  The first version of OpenFlow is 1.0.0 was released on 2009 but the last standardized version v1.5.1 was approved only on march 2015 \cite{ONF_report}.
+  %corrects
+  %todo spostare l'immagine correttamente
 \begin{figure}[!h]
@@
   \textbf{Openflow 1.0 messages}
-Openflow defines a number of messages between a controller and a switch. Now there will be presented some messages we exploited:\cite{openflow}
+
+Openflow defines a number of messages between a controller and different switches. Now there will be presented some messages we exploited:\cite{openflow}
 \begin{itemize}
@@
 	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
-\item \emph{FeatureReq} \newline this is the message that the controller sends to a switch to know its abilities. The switch
+\item \emph{FeatureReq} \newline This is the message that the controller sends to a switch to get to know its abilities. The switch
 	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
@@
 \item \emph{PortStatus} \newline this message is sent from a switch to the controller, to notify that some ports has changed their status.
-		It's useful to check if a port has someone connected
-\item \emph{StatsReq} \newline The controller will send this message to request the statistics. The payload of the message contains the type of statistic the controller want to know.
-		The switch then will answer with the Stats reply, if it supports that type of stat. It will send a Bad_Request message if not.
-\item \emph{PacketIn} \newline This message is sent from the switch to the controller whenever a packet arrives and it doesn't match
-		any rule. The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
+		It's useful to check if a port has someone connected to it.
+\item \emph{StatsReq} \newline The controller will send this message to request the statistics. The payload of the message contains the type of 
+statistic the controller wants to know.
+		The switch then will answer with the Stats reply, if it supports that type of stat. If not, The switch will send a Bad_Request
+		message.
+\item \emph{PacketIn} \newline This message is sent from the switch to the controller everytime a packet arrives without matching any rule.
+The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
 		a flow rule for packets like this using a FlowMod.
-\item \emph{PacketOut} \newline This message is sent from the controller to forward a packet, which is the payload of the PacketOut message, to some
-		port(s).
-\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch. Is the message that allows routing.
+\item \emph{PacketOut} \newline This message is sent from the controller to forward a packet, which is the payload of the PacketOut message,
+to some ports.
+\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch.This is the message that allows routing.
 \\
@@
 \begin{itemize}
- \item \texttt{match fields:} sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze. Contengono la porta d'ingresso, l'intestazione del pacchetto,
+ \item \texttt{match fields:} fields that compare packets in order to look fo sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze.
+Contengono la porta d'ingresso, l'intestazione del pacchetto,
  e altri campi opzionali;
@@
   (aggiungere qualcosa e correggere)
+
+  \begin{figure}[!h]
+ \centering
+ \includegraphics[scale=0.30]{images/topology.png}
+ \caption{\emph{Experience scenario}}
+ \label{fig:topo}
+  \end{figure}
+
   \#TODO matteo

### 2017-03-11 report/paper.tex
other corrections

@@
-\documentclass[conference,10pt]{IEEEtran}
+\documentclass[article,10pt]{IEEEtran}
 %\documentclass[conference,draft,onecolumn]{IEEEtran}
@@
 
+\setcounter{page}{1}
 %\mainmatter
@@
 
-Openflow defines a number of messages between a controller and different switches. Now there will be presented some messages we exploited:\cite{openflow}
+Openflow defines a number of messages between a controller and different switches. Now there will be presented some messages we exploited\cite{openflow}:
 \begin{itemize}
@@
 \begin{itemize}
- \item \texttt{match fields:} fields that compare packets in order to look fo sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze.
-Contengono la porta d'ingresso, l'intestazione del pacchetto,
- e altri campi opzionali;
- \item \texttt{priority:} precedenza nel caso di voci con pi\`u corrispondenze;
- \item \texttt{counter:} incrementato ogni qualvolta che viene riscontrata la corrispondenza con un pacchetto;
- \item \texttt{istructions:} istruzioni da aggiungere all'\emph{Action-Set} o modifiche al flusso della pipeline;
- \item \texttt{timeouts:} massimo tempo per cui lo switch tiene in memoria questa voce, altre il quale la voce viene scartata;
- \item \texttt{cookie:} valori utilizzati per statistiche riguardo la singola voce, quindi non vengono considerati nell'elaborazione dei pacchetti dati;
- \item \texttt{flags:} utilizzate per differenziare la gestione della voce;
+ \item \texttt{match fields:} fields that compare packets in order to look for correspondences. They contain the in_port, the packet's header and others optional fields.
+ \item \texttt{priority:} priority in case of rows with more than one correspondence;
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+ \item \texttt{counter:} this field was improved every time a correspondence with a packet was matched;
+ \item \texttt{istructions:} a set of action that the switch add on \emph{Action-set} of packet or modification on pipeline flux;
+ \item \texttt{timeouts:} this is the maximum time afther that the rule exired;
+ %\item \texttt{cookie:} valori utilizzati per statistiche riguardo la singola voce, quindi non vengono considerati nell'elaborazione dei pacchetti dati;
+ \item \texttt{flags:} this field is use to differentiate the management of the voice;
  \\
@@
 \textbf{Match Constraints}
-principal match method that we can used with OpenFlow v1.0:
+
+these are theprincipal match method that we can use with OpenFlow v1.0:
 \begin{itemize}
@@
 \textbf{Action}
-With the OpenFlow v1.0 in practice we can do simply things:
+
+With the OpenFlow v1.0 in practice we can do simply things for example:
 \begin{itemize}
@@
 \end{itemize}
-With from OpenFlow v1.1 in poi we can modify Time To Live and other things.
+
+From OpenFlow v1.1 on we can modify Time To Live and other things that allow us to do something more.
 
@@
 some types of flows, so we used our PCs to sniff the traffic. Traffic analysis will be discussed in section.
-  \#TODO stato dell arte e paper di riferimento  matteo
+  \#TODO stato dell arte e paper di riferimento  prova a darci un occhio adrea please.
 
@@
 The main gool of this experience is to develop a small SDN network in a lab with real hardware
-in orther to understand in a practical level how to implement a \textit{Quality of Service, \textbf{QoS}} service in a SDN scenarios.
-Our network have to manage traffic with a global scope, in order to evolve in base on different parameters.
+in orther to understand in a practical level how to implement a \textit{Quality of Service} in a SDN scenarios.
+Our network have to manage traffic with a global purpose, in order to evolve according to different parameters.
+
 Moreover the SDN network need to interact with world wide web efficently.
@@
 \subsection{Adaptability of the network}
-Made Network Adaptable means poter dirottare il traffico in real time to ensure a lot of demands. For example we have to avoid link congestion,
-but we can prefer reroute some hosting traffic to worsen it's performance because  it is the traffic we don't want to encourage.
+An adaptable Network has to real-time reroute traffic in order to ensure a lot of demands.
+A general efficient Network has to manage some tasks for example: avoid links congestion, reroute traffic when some malfunction accure on a link
+and running some link path algorithm. 
+In addition our network has to be adaptable in orther to ensure a QoS type of service.
+
 To reach this gools we need to have a global knowledge on the network and we have to modify the routing table on every routerboard with a
@@
 In order to ensure the traffic shaping based on network informations, there is a need of a structure,
-automatically updatable, where the link statistics can be saved to allow the best path for a data stream. The best metric will be described below.
-%  assicurare queste ci serve una struttura dati che si aggiorna costantemente nella quale immagazzinare le statistiche dei link
-% in modo all'occorrenza di trovare il percordo migliore per ogni singolo flusso dati.
+automatically updatable, where the link statistics can be saved to allow the best path for a data stream. 
+The best metric will be described below.
 \#TODO MATTEO
@@
 \end{itemize}
-\textbf{firmware} the current version is: '' da gercare col comando: system routerboard print ''
+\textbf{firmware} the current version is: '' da cercare col comando: system routerboard print ''
 \newline
-The \textbf{Operating System} that they use is MikroTik RouterOS ( cercare la versione installata) \cite{routeboard_software}.
+The \textbf{Operating System} that they use is MikroTik RouterOS \cite{routeboard_software}.
 \newline
-One limitation of this routerboard is that it has only 5 ethernet port this is a limitation because consider that one of this port it is
-used to connect conntroller and with other port we have made host and other routerboard connection.
+One limit of this routerboard is that it has only 5 ethernet ports and this is a limit because one of these ports must be
+reserved for the connection with the controller, consequently we have to connect both host and other routerboards using only 4 ports.
 
-%correggere e completare Matteo
+%correggere
 
@@
 \begin{itemize}
- %\item \texttt{pox.topology.launch()} and \texttt{pox.openflow.topology.launch()} descrivere a che ci serve\cite{pox}.
- \item \texttt{pox.openflow.discovery.launch()} this function run algorithm in order to discovery the network and every time in witch some new
- topological change occour it return an event with ( con cui ) we can update our topological knowledge.
+ \item \texttt{pox.openflow.discovery.launch()} this function runs an algorithm in order to discovery the network. Every time a new
+ topological change occur this function returns an event on POX with which we can update our topological knowledge.
  It does not provide the host event. For host event\cite{pox}.
@@
 \subsection{Graph and Network structures}
-  In order to achieve the realtime adaptability and statistics storage we have develop ad-hoc structure.
+  In order to achieve the realtime adaptability and statistics storage we have developed ad-hoc structure.
   We use a specific python library called NetworkX \cite{networkx} to create a graph in witch we can storage ad-hoc node that represent
   resplectively switch or host element.
-  This topological structure is updated realtime thanks to connection event and link discover offered by various library.
-  ( mettere i vari eventi che sfrutto )
-  The link besides have many parameters in witch we can save network performance. To update every link parameter we have to creat ad hoc function
-  because there isn't some function constructed yet. ( sistemare frase) .
-  This type of structure allow us to run Dijkstra algorithm that return the minimum path ( in base a vari parametri ) on a nlog n time ( verificare).
-  (aggiungere qualcosa e correggere)
+  This topological structure is updated live thanks to connection event and link discover offered by POX's libraries:
+  \begin{itemize}
+   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type give us two type of information; a new link was added and a still present link was expired;
+   \item \texttt{_handle_ConnectionUp}: from core.openflow this event join up when a new switch was added on the network and give us
+   some information of the switch, in particular its ports configuration;
+   \item \texttt{_handle_ConnectionUp}: from core.openflow like the event explained above comes up when a switch was disconnected from the network;
+   \item \texttt{_handle_port_status}: from core.openflow this event occure when on a particular switch there is a change on its port, for example
+   some host connect or the link was expired;\\
+  \end{itemize}
+
+  The link on our network besides have many parameters in witch we can save network performance. 
+  In orther to update every link parameters we had to creat ad hoc function
+  because there isn't some function constructed yet.
+
+  This type of structure allow us to run Dijkstra a polynomial algorithm that discover the shorthes path according to various parameters.
+
+  The underlying figure shown the visual output of our laboratory network's topology offered by the controller, we can see 4 different switches and all of the other hosts
+  differentiated with their IP address.
 
@@
 
-  \#TODO matteo
+  %\#TODO matteo
 

### 2017-03-13 report/paper.tex
altre correzioni

@@
 
-\textbf{Forwarding table:}
+\textbf{Forwarding table}
 \begin{itemize}
- \item \texttt{match fields:} fields that compare packets in order to look for correspondences. They contain the in_port, the packet's header and others optional fields.
- \item \texttt{priority:} priority in case of rows with more than one correspondence;
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- \item \texttt{counter:} this field was improved every time a correspondence with a packet was matched;
- \item \texttt{istructions:} a set of action that the switch add on \emph{Action-set} of packet or modification on pipeline flux;
- \item \texttt{timeouts:} this is the maximum time afther that the rule exired;
- %\item \texttt{cookie:} valori utilizzati per statistiche riguardo la singola voce, quindi non vengono considerati nell'elaborazione dei pacchetti dati;
- \item \texttt{flags:} this field is use to differentiate the management of the voice;
+ \item \texttt{match fields:} fields that compare packets in order to look for correspondences. They contain the in_port, the packet's header and others optional fields;
+ \item \texttt{priority:} this field is used when there are rows with more than one correspondence;
+ \item \texttt{counter:} this field was improved every time a correspondence with a packet is matched;
+ \item \texttt{istruction:} a set of actions that the switch add on \emph{Action-set} of the matched packet or some instruction 
+ in order to modify the pipeline flux;
+ \item \texttt{timeout:} this is the maximum time afther which the rule expire;
+ \item \texttt{flag:} with this field we can differenciate similar voices in order to managed multiple table with similar rules;
  \\
@@
 \textbf{Match Constraints}
-
-these are theprincipal match method that we can use with OpenFlow v1.0:
+these are the principal matching method that we can use with OpenFlow v1.0:
 \begin{itemize}
@@
  \item \texttt{nw_proto}: IPv4 Protocol;
- \item \texttt{nw_src/nw_dst}: IPv4 Source/Destination;
- \item \texttt{tp_src/tp_dst}: TCP/UDP Source/Destination Port;
+ \item \texttt{nw_src/nw_dst}: IPv4 Source/Destination addresses;
+ \item \texttt{tp_src/tp_dst}: TCP/UDP Source/Destination ports;
  \\
@@
 
-\textbf{Action}
-
-With the OpenFlow v1.0 in practice we can do simply things for example:
+\textbf{Action} With OpenFlow v1.0 we can do basic tasks in practice, the most important are for example:
 \begin{itemize}
  \item \texttt{Output}: set output port;
- \item \texttt{SetDLSrc/Dst}: modify packet MAC address;
- \item \texttt{SetNWSrc/Dst}: modify packet IPv4 address;
- \item \texttt{SetNWTos}
+ \item \texttt{SetDLSrc/Dst}: modify packet's MAC address;
+ \item \texttt{SetNWSrc/Dst}: modify packet's IPv4 address;
+ \item \texttt{SetNWTos}: change ToS fields of message;
  \item \texttt{SetTPSrc/Dst}
@@
 
-From OpenFlow v1.1 on we can modify Time To Live and other things that allow us to do something more.
+From OpenFlow v1.1 on we can modify Time To Live and other options that allow us to do something more complex.
 
@@
   The work presented in
-"Efficient topology discovery in OpenFlow-based Software Defined Networks" by F. Pakzad \cite{farzaneh}
-has been implemented to improve the controller speed and efficiency as it make it more reactive.
-The introduction to the statistics power in a SDN network using the OpenFlow protocol has been found
-in the paper called "Getting traffic statistics from network devices in an SDN environment using OpenFlow" by D.J. Hamad \& others. [ put note here]
-For traffic analysis we haven't found any paper describing the characteristics of
-some types of flows, so we used our PCs to sniff the traffic. Traffic analysis will be discussed in section.
-  \#TODO stato dell arte e paper di riferimento  prova a darci un occhio adrea please.
+"Efficient topology discovery in OpenFlow-based Software Defined Networks" by F. Pakzad 
+has been implemented to improve the controller speed and efficiency as it make it more reactive \cite{farzaneh}.
+\newline The introduction to the statistics power in a SDN network using the OpenFlow protocol has been found
+in the paper called "Getting traffic statistics from network devices in an SDN environment using OpenFlow" by D.J. Hamad \& others \cite{stat}.
+\newline For traffic analysis we haven't found any paper describing the characteristics of
+some types of flows, so we used our PCs to sniff the traffic. \newline 
+Traffic analysis will be discussed in section \ref{subsec:traf}.
 
-\section{Needs and Goals}\label{sec:obb} %obbiettivi ed esigenze controllare
-The main gool of this experience is to develop a small SDN network in a lab with real hardware
-in orther to understand in a practical level how to implement a \textit{Quality of Service} in a SDN scenarios.
+\section{Needs and Goals}\label{sec:obb}
+The main purpose of this experience is to develop a small SDN network in a lab with physical hardware
+in order to understand how to improve a Quality of Service in a SDN scenery with a real scenario.
 Our network have to manage traffic with a global purpose, in order to evolve according to different parameters.
 
-Moreover the SDN network need to interact with world wide web efficently.
+Moreover the SDN network have to efficiently interact with the world wide web.
+
 \subsection{Connectivity}\label{sec:connectivity}
@@
 \\ So this approach allowed us to make the network unaware that it's not a legacy one and allowed connectivity to be provided.
-%correggere
+
 \subsection{Adaptability of the network}
 An adaptable Network has to real-time reroute traffic in order to ensure a lot of demands.
-A general efficient Network has to manage some tasks for example: avoid links congestion, reroute traffic when some malfunction accure on a link
-and running some link path algorithm. 
-In addition our network has to be adaptable in orther to ensure a QoS type of service.
+A general efficient Network has to manage some tasks, for example avoiding links congestion, reroute traffic when some malfunction accure on a link
+and running a link path algorithm. 
+In addition our network has to be adaptable in order to ensure a QoS type of service.
 
-To reach this gools we need to have a global knowledge on the network and we have to modify the routing table on every routerboard with a
-global scope.
+To reach this gools we needed to have a global knowledge on the network and we had to modify the routing table on every routerboard with a
+global purpose.
 The performance of a single connection and entire network can ensure various type of metrics.
-In order to ensure the traffic shaping based on network informations, there is a need of a structure,
-automatically updatable, where the link statistics can be saved to allow the best path for a data stream. 
+In order to ensure the traffic shaping based on network informations, there is a need of an automatically updatable structure,
+where the link statistics can be saved to allow the best path for a data stream. 
 The best metric will be described below.
-\#TODO MATTEO
+
 \subsection{Exterior comunication}
@@
 \subsection{Controller}
-The controller that we use is POX, the software is opensource and it is write with Python, it is based on SDN and support OpenFlow.
-The software use the version 2.7 of Python.
-It's repository is at this link \url{https://github.com/noxrepo/pox} and a lot of documentation about the progect
-is present at this link \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
-The only problem to use this controller is that it support only the first version of OpenFlow1.0.
+The controller that we use is POX, the software is opensource and it is written with Python, based on SDN and supports OpenFlow.
+The software uses the version 2.7 of Python.
+It's repository, together with lot of documentation regarding the project can be found at these links \url{https://github.com/noxrepo/pox}, \url{https://openflow.stanford.edu/display/ONL/POX+Wiki}.
+The only problem in using this controller is that it supports only the first version of OpenFlow1.0.
 \\
-The controller run inside a VirtualMachine on our personal computer,
-the VM is offered by SDN hub on this site: \url{http://sdnhub.org/tutorials/sdn-tutorial-vm/}.
+The controller runs inside a VirtualMachine on our personal computer and the VM is offered by SDN hub on this site: \url{http://sdnhub.org/tutorials/sdn-tutorial-vm/}.
 This VM is a 64-bit Ubuntu 14.04 that has a number of SDN software and tools installed.
-% \#TODO matteo da correggere e completare
+
 \subsection{Router Board}
@@
 
-The router board that we used is MikroTik miniROUTER RouterBOARD 450G. This Router Board virtualize an OpenFlow switch that support
+The router board that we used is MikroTik miniROUTER RouterBOARD 450G. This Router Board virtualizes an OpenFlow switch that supports
 OpenFlow v1.0\cite{routerboard_doc}.
@@
 \end{itemize}
-\textbf{firmware} the current version is: '' da cercare col comando: system routerboard print ''
-\newline
+%\textbf{firmware} the current version is: '' da cercare col comando: system routerboard print ''
+
 The \textbf{Operating System} that they use is MikroTik RouterOS \cite{routeboard_software}.
@@
 One limit of this routerboard is that it has only 5 ethernet ports and this is a limit because one of these ports must be
-reserved for the connection with the controller, consequently we have to connect both host and other routerboards using only 4 ports.
-
-%correggere
+reserved for the connection with the controller, that is why we have to connect both host and other routerboards using only 4 ports.
 
@@
 \subsection{Controller}
-We developed a controller that can handle a mesh network, which has loops, that avoid
-the traffic flooding. The controller stores the switch and the connection between them;
-knows which hosts are connected to each switch port, with the hypothesys that only one
-host is connected at each port, unless the port is the one connected to the internet.
-With the use of the statistics, the controller get the status of the network and, if necessary,
-tries to adjust the traffic and put in the link with the higher path loss the torrent traffic,
-make the connection between two gaming host the one with less delay and tries to redistribute all the traffic in a way
-such that all the nodes have an average throughput as lowest as possible.
+We developed a controller that can handle a mesh network with loops, avoiding
+packets flooding. The controller stores the switch and the connection between them and knows which hosts are connected to each switch port starting from the hypothesys that only one
+host is connected to each port, unless the port is the one connected to the internet.
+With the use of statistics, the controller gets the status of the network and, if necessary,
+tries to adjust the traffic and to put in the link with the higher path loss the torrent traffic.
+On the other hand it sets up the connection for two gaming hosts going through a link with less delay and tries to redistribute all the 
+traffic in a way that all the nodes have an average throughput as lowest as possible.
 \\
 \newline
-We have used some principal function released with POX controller:
+We have used some principal functions released with POX controller:
 \begin{itemize}
- \item \texttt{pox.openflow.discovery.launch()} this function runs an algorithm in order to discovery the network. Every time a new
- topological change occur this function returns an event on POX with which we can update our topological knowledge.
- It does not provide the host event. For host event\cite{pox}.
+ \item \texttt{pox.openflow.discovery.launch()} this function runs an algorithm in order to discover the network. Every time a new
+ topological change occurs this function returns an event on POX with which we can update our topological knowledge.
+ It does not provide the host event\cite{pox}.
+
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology
@@
  switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service
- on the switch in orther to avoid loop\cite{pox}.
+ on the switch in order to avoid loop\cite{pox}.
 \end{itemize}
@@
 
-We have use 4 MikroTik Router board that were provided us in the Communication Networks Lab at DEI in University of Padua.
-The 4 routerboard is connected each other in order to design a loop. Every sigle board connect two host and is connected with a legacy switch that provide
-to logically connected with controller.
+We have use 4 MikroTik Routerboards that were provided us in the Communication's Network Lab at DEI in University of Padua.
+The 4 routerboards are connected each other in order to design a loop. Every sigle board connects two host and is connected with a 
+legacy switch which provides logically connection with controller.
 
 \subsection{Graph and Network structures}
-  In order to achieve the realtime adaptability and statistics storage we have developed ad-hoc structure.
-  We use a specific python library called NetworkX \cite{networkx} to create a graph in witch we can storage ad-hoc node that represent
-  resplectively switch or host element.
-  This topological structure is updated live thanks to connection event and link discover offered by POX's libraries:
+  In order to achieve the realtime adaptability and statistics storage we developed ad-hoc structure.
+  We used a specific Python's library called NetworkX \cite{networkx}. NetworkX allow us to create a graph in witch we can storage proper nodes
+  that represent resplectively switchs or hosts elements.
+  This topological structure is updated live thanks to connection events and links discover offered by POX's libraries:
   \begin{itemize}
-   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type give us two type of information; a new link was added and a still present link was expired;
-   \item \texttt{_handle_ConnectionUp}: from core.openflow this event join up when a new switch was added on the network and give us
+   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type give us two types of information; a new link is added and 
+   a still present link is expired;
+   \item \texttt{_handle_ConnectionUp}: from core.openflow this event join up when a new switch is added on the network and give us
    some information of the switch, in particular its ports configuration;
-   \item \texttt{_handle_ConnectionUp}: from core.openflow like the event explained above comes up when a switch was disconnected from the network;
-   \item \texttt{_handle_port_status}: from core.openflow this event occure when on a particular switch there is a change on its port, for example
-   some host connect or the link was expired;\\
+   \item \texttt{_handle_ConnectionUp}: from core.openflow like the event explained above comes up when a switch is disconnected from the network;
+   \item \texttt{_handle_port_status}: from core.openflow this event occure when on a particular switch there is a change about its ports, for example
+   some host is connected or the link is expired;\\
   \end{itemize}
 
-  The link on our network besides have many parameters in witch we can save network performance. 
-  In orther to update every link parameters we had to creat ad hoc function
+  The link on our network besides has many parameters in witch we can save network's performance. 
+  In order to update every links parameters we had to create ad hoc functionality
   because there isn't some function constructed yet.
@@
   For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
-  complexity so the routerboard can made only some easy function in orther to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
+  complexity so the routerboard can made only some easy function in order to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
   Il che ha portato a dover utilizzare stratagemmi.
@@
 Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
-\newline This part required a particolar flow entry on the second switch in orther to send to the controller only some particulare packet.
+\newline This part required a particolar flow entry on the second switch in order to send to the controller only some particulare packet.
 \begin{figure}[!h]
@@
 	\bibitem{rtt} Efficient Round-Trip Time Monitoring in OpenFlow Networks A. Atary A. Bremler-Barr Interdisciplinary Center Herzliya \url{http://www.deepness-lab.org/pubs/infocom2016_grami.pdf}
+	\bibitem{stat} Getting traffic statistics from network devices in an SDN environment using OpenFlow Diyar Jamal Hamad , Khirota Gorgees Yalda, and Ibrahim Tanner Okumus
+	\url{http://itas2015.iitp.ru/pdf/1570195931.pdf};
 

### 2017-03-14 report/paper.tex
added conclusions. need review. doing this evening

--- 2017-03-11 report/paper.tex
@@
 
-%\mainmatter
+\mainmatter
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Quality of Service with Software Defined Network.}
+\title{Demo for a Quality of Service networking with Software Defined Network.}
 
@@
 \begin{abstract}
-Quality of Service (QoS) in existing network architectures is an ongoing problem.
-A lot of researchers and industries tried to solve this problem and found different solutions but some of them are too
-expensive other failed or were not implemented.
-Software Defined Networking (SDN) paradigm has emerged in response to traditional networking architecture's limitations.
-The main advantage of SDN is a centralized global network view. This particular feature pushed researchers to keep on improving QoS.
-
-In this report we will focus on two different aspects; all problems we faced during our attempt to create a based SDN application able to
-support a QoS service and the results we achieved.
-
-OpenFlow standard and python POX controller, which is largely spread, were used as instruments to make this experience possible,
-despite the limit of using a new type of tecnology still in the research phase.
-%corretto
+Quality of Service(QoS) in existing network architectures is an ongoing problem.
+A lot of researcher and industries try to solve this problem with many solution but a lot of them are too
+expensive other failed or are not implemented.
+Software Defined Networking(SDN) paradigm has emerged in response to limitations of traditional networking architectures.
+The main advantages of SDN are centralized global network view. This particular features have got attention of researchers to improve QoS.
+In questo report mostreremo quali sono stati i principali risutati raggiunti e i problemi incontrati nella realizzazione di
+un applicativo SDN in grado di supportare un servizio di tipo QoS.
+Lo standard OpenFlow e il controller python POX largamente diffuso sono stati gli strumenti che hanno reso possibile la
+realizzazione di quest'esperienza pur mostrando i limiti dovuti all'utilizzo di una tecnologia nuova e in fase di ricerca.
+%translate
 \end{abstract}
@@
 \section{Introduction}\label{sec:intro}
-Software-defined networking (SDN) is a way to manage networks that separates the control from the forwarding plane.
-SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brain” of the network.
+Software-defined networking (SDN) is way to manage networks that separates the control plane from the forwarding plane.
+SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brains” of the network.
 The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
-One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it isn’t the only SDN standard. 
-Centralized, programmable SDN environments can easily adjust to the rapidly changing business'needs. SDN can reduce costs and restrict
+One of the most well-known protocols used by SDN Controllers is OpenFlow, however, it isn’t the only SDN standard, but it is widely accepted.
+Centralized, programmable SDN environments can easily adjust to the rapidly changing needs of businesses. SDN can lower costs and limit
 wasteful provisioning, as well as provide flexibility and innovation for networks.
-%corretto
+\#TODO da completare
   \subsection{OpenFlow}
-  During 2006 ,thanks to M. Casado, a new network architecture was developed, called Ethane.
-  On this conncept was born, during 2011, the Open Network Foundation, a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
-  The first version of OpenFlow is 1.0.0 was released on 2009 but the last standardized version v1.5.1 was approved only on march 2015 \cite{ONF_report}.
-  %corrects
-  %todo spostare l'immagine correttamente
+  During 2006 thanks to M. Casado was developed Ethane a new network architecture on this conncept during 2011 was bord the Open Network Foundatin.
+  This is a no profit association that would innovate SDN and standardize OpenFlow protocol whit relatives tecnology \cite{ONF}.
+  The first version of OpenFlow is 1.0.0 release on 2009 but the last standardized version is v1.5.1 approved on march 2015 \cite{ONF_report}.
+  \#TODO da sistemare
 \begin{figure}[!h]
@@
   \textbf{Openflow 1.0 messages}
-
-Openflow defines a number of messages between a controller and different switches. Now there will be presented some messages we exploited:\cite{openflow}
+Openflow defines a number of messages between a controller and a switch. Now there will be presented some messages we exploited:\cite{openflow}
 \begin{itemize}
@@
 	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
-\item \emph{FeatureReq} \newline This is the message that the controller sends to a switch to get to know its abilities. The switch
+\item \emph{FeatureReq} \newline this is the message that the controller sends to a switch to know its abilities. The switch
 	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
@@
 \item \emph{PortStatus} \newline this message is sent from a switch to the controller, to notify that some ports has changed their status.
-		It's useful to check if a port has someone connected to it.
-\item \emph{StatsReq} \newline The controller will send this message to request the statistics. The payload of the message contains the type of 
-statistic the controller wants to know.
-		The switch then will answer with the Stats reply, if it supports that type of stat. If not, The switch will send a Bad_Request
-		message.
-\item \emph{PacketIn} \newline This message is sent from the switch to the controller everytime a packet arrives without matching any rule.
-The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
+		It's useful to check if a port has someone connected
+\item \emph{StatsReq} \newline The controller will send this message to request the statistics. The payload of the message contains the type of statistic the controller want to know.
+		The switch then will answer with the Stats reply, if it supports that type of stat. It will send a Bad_Request message if not.
+\item \emph{PacketIn} \newline This message is sent from the switch to the controller whenever a packet arrives and it doesn't match
+		any rule. The controller can decide to ignore that type of message, insert this message in a PacketOut message, or set
 		a flow rule for packets like this using a FlowMod.
-\item \emph{PacketOut} \newline This message is sent from the controller to forward a packet, which is the payload of the PacketOut message,
-to some ports.
-\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch.This is the message that allows routing.
+\item \emph{PacketOut} \newline This message is sent from the controller to forward a packet, which is the payload of the PacketOut message, to some
+		port(s).
+\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch. Is the message that allows routing.
 \\
@@
 \begin{itemize}
- \item \texttt{match fields:} fields that compare packets in order to look fo sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze.
-Contengono la porta d'ingresso, l'intestazione del pacchetto,
+ \item \texttt{match fields:} sono i valori con cui i pacchetti vengono confrontati per cercare corrispondenze. Contengono la porta d'ingresso, l'intestazione del pacchetto,
  e altri campi opzionali;
@@
   (aggiungere qualcosa e correggere)
-
-  \begin{figure}[!h]
- \centering
- \includegraphics[scale=0.30]{images/topology.png}
- \caption{\emph{Experience scenario}}
- \label{fig:topo}
-  \end{figure}
-
   \#TODO matteo
@@
 		\subsubsection{Link Capacity}
-
-	\#TODO andrea
 	The first data that was collected was the actual link capacity, which depends both on the cable used to connect
@@
   \subsubsection{Throughput}
-	\#TODO andrea
 	Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
@@
 \subsubsection{Packet error rate}
-
 The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
@@
 \subsection{Route changing} %cambio rotte
-
-\#TODO andrea
 For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
@@
 \subsection{ARP solution}
-
 During our analysis and in our tests, we prove that in a SDN mesh network, the ARP packets are useless.
@@
 \subsection{Ip switching in pox}
-
 During our tests, both with Mininet and in the lab, we found out that the core of POX, when a PacketIn event occoures, switches the source and the destination addresses,
@@
 \section{Conclusions}\label{sec:conclusion}
-Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
-ehi
-\#TODO
+% Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
+% ehi
+For this course project, the report shows how the demo was made, in particular what problem we encountered
+and what we exploited, showing what protocols could be unused in the future. We showed a possible simple data structure
+that can allow the controller to be able to know the network, its topology and which route to set, based on the traffic.
+Morover, we showed how to exploit the potentialities of the SDN network, giving particular importance on the
+statistics and the path switching capabilities.
+The multi-graph solution allowed us to connect the network and, based on its condition, choose which weight to apply
+in order to give a Quality of Service based network. With some boards that supports the newer version of Openflow
+(i.e. 1.4, the latest), the code can be ported, so it can be used with the Ryu core controller and more statistic can
+be exploited. Moreover, the controller can set more strictive rules for the flow control and can get more informations from the
+switches.
 

### 2017-03-14 report/paper.tex

--- 2017-03-13 report/paper.tex
@@
  It does not provide the host event\cite{pox}.
+ \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology
+ because with a spanning_tree algorithm it avoids the broadcast storm of messages.
 
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- \item \texttt{pox.openflow.spanning_tree.launch()} this function is necessary with a presence of loop on topology
- because with a spanning_tree algorithm it avoid the broadcast storm of message. When we set up \texttt{OFPP_FLOOD} like output port action in a table's
- switch row if some spanning_tree algorithm run the \texttt{OFPP_FLOOD} is only theoretical infact the algorithm close some port for flooding service
- on the switch in order to avoid loop\cite{pox}.
+When we set up this option \texttt{OFPP_FLOOD} like output port's action in a table's switch row if some spanning_tree algorithm is working
+the \texttt{OFPP_FLOOD} is only theoretical in the sense that the algorithm excludes some ports during the flooding option, in order to avoid loop\cite{pox}.
 \end{itemize}
 \subsection{Real Lab Scenario}
-The topological scenario in which we have set up experience is the next:
+The topological scenario in which we have set up experience is as follows:
 \begin{figure}[!h]
@@
 
-We have use 4 MikroTik Routerboards that were provided us in the Communication's Network Lab at DEI in University of Padua.
-The 4 routerboards are connected each other in order to design a loop. Every sigle board connects two host and is connected with a 
+We have use 4 MikroTik Routerboards that were provided us by the Communication's Network Lab in the department of information  engineering to 
+University of Padua.
+The 4 routerboards are connected each other in order to design a topological loop. Then every sigle board connects two hosts and is connected with a 
 legacy switch which provides logically connection with controller.
+The last thing is a host that made gateway's function; we have connected its with both routerboard and legacy network and set up this host like default 
+gateway on the controller.
 
-\subsection{Graph and Network structures}
-  In order to achieve the realtime adaptability and statistics storage we developed ad-hoc structure.
-  We used a specific Python's library called NetworkX \cite{networkx}. NetworkX allow us to create a graph in witch we can storage proper nodes
-  that represent resplectively switchs or hosts elements.
+\subsection{Graph and Network's structures}
+  In order to achieve the realtime adaptability and statistics storage we developed ad-hoc structures.
+  We used a specific Python's library called NetworkX \cite{networkx}. NetworkX allow us to create a graph in witch we can storage suitable nodes
+  that represent resplectively switches or hosts.
   This topological structure is updated live thanks to connection events and links discover offered by POX's libraries:
   \begin{itemize}
-   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type give us two types of information; a new link is added and 
+   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type gives us two types of information; a new link is added and 
    a still present link is expired;
-   \item \texttt{_handle_ConnectionUp}: from core.openflow this event join up when a new switch is added on the network and give us
+   \item \texttt{_handle_ConnectionUp}: from core.openflow this event joins up when a new switch is added on the network and gives us
    some information of the switch, in particular its ports configuration;
    \item \texttt{_handle_ConnectionUp}: from core.openflow like the event explained above comes up when a switch is disconnected from the network;
-   \item \texttt{_handle_port_status}: from core.openflow this event occure when on a particular switch there is a change about its ports, for example
-   some host is connected or the link is expired;\\
+   \item \texttt{_handle_port_status}: from core.openflow this event occures when on a particular switch there is a change about its ports, for example
+   a new host has just been connected or the link is expired;\\
   \end{itemize}
 
-  The link on our network besides has many parameters in witch we can save network's performance. 
-  In order to update every links parameters we had to create ad hoc functionality
-  because there isn't some function constructed yet.
+  The links on our network logically have many parameters in witch we can storage network's performance; in practice a NetworkX's graph support only 
+  one or two link's attributes so we have create a lots of graphs with the same topological information but different link's attributes.
 
-  This type of structure allow us to run Dijkstra a polynomial algorithm that discover the shorthes path according to various parameters.
+  In order to update every link's parameters we had to create ad-hoc functionality
+  since there are not proper functions constructed before.
 
-  The underlying figure shown the visual output of our laboratory network's topology offered by the controller, we can see 4 different switches and all of the other hosts
-  differentiated with their IP address.
+  This type of structure allow us to run Dijkstra's polynomial algorithm that discover the shortest path according to various parameters.
+
+  The following picture shows the visual output of our laboratory network's topology offered by the controller, we can see 4 different switches and all of the other hosts
+  differentiated with their IP addresses.
 
@@
 
-  %\#TODO matteo
-
-  \subsection{Traffics analisis}\label{subsec:traf}
+\subsection{Traffic's Analysis}\label{subsec:traf}
 We used laptop connected via
@@
 \#Mb for the gaming traffic, \#Mb for the VoIP traffic and 3GB for the torrent traffic.
-The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP)
+The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP).
 
@@
 On another test, we tried the VoIP while a download was ongoing, and we saw that the quality
-decreased to allow communication. The traffic is between the UDP port 80 \#\#TODO\#\# TOCHECK \#\#
+decreased to allow communication. The traffic is between the UDP port 80.
 
-The last traffic analysis we made was with a DOTA[explain what it is] web match.
-We saw that for that type of gaming session, the packet size is about \# bytes and involves both TCP and UDP ports.
-  \\
-  \\
   For every networks performance we have tryed to develop a specific solution. The OpenFlow main goal is to centralize controll and software
-  complexity so the routerboard can made only some easy function in order to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
-  Il che ha portato a dover utilizzare stratagemmi.
-  \#TODO %riformulare
+  complexity so the routerboards can made only some easy function in order to contain the hardware complexity and mantainly a good speed\cite{qos_paper}.
+  This led to the necessity of stratagems utilization.
+
 \subsubsection{Delay}
 Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
-The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  (hosts  or  middleboxes).
+The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  hosts  or  middleboxes.
 In traditional  networks, the control over the network routing is
-limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  sim-plifies
+limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  simplifies
 network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
@@
 \newline
-Now we create a proper message with unique identifier number, we start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
+Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
 Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
@@
 \\
-Teorethically this method must works but in practice there are a lot of delay introduce from computational concestion on the virtualized controller.
-What we observe is a estimated delay like some seconds that is impossible, infact if we made some \texttt{ping test} from one host to another host
+Teorethically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
+What we observe is a estimated delay like a few seconds which is not the case, infact if we made some \texttt{ping test} from one host to another host
 in this SDN network we realize delay around 3ms.
@@
 
-	\#TODO andrea
 	The first data that was collected was the actual link capacity, which depends both on the cable used to connect
@@
   \subsubsection{Throughput}
-	\#TODO andrea
+
 	Using the flow statistics, we analyzed the traffic going through a link and, with a simple division, we got the normalized throughput.
@@
 those are indirect measures themself and the error derived from the calculations could grow up.
-\#TODO andrea
-\\
-\subsection{Route changing} %cambio rotte
 
-\#TODO andrea
+\subsection{Route changing}
+
 For each statistic we got and each parameter we obtained, we changed the weight of the graphs. Every graph we used had a specific weight, so the
@@
 \subsection{Default Gateway}
+
 Analyzing the traffic while we were trying the controller on real machines in the lab, we have tried to set the default gateway to
@@
 destination address, we can switch it back; the result is that the PacketIn event now has correct informations.
-If the two host are in the same switch, the problem is solved using a controlled packet flooding, described in section \ref{sec:connectivity}
+If the two host are in the same switch, the problem is solved using a controlled packet flooding, described in section \ref{sec:connectivity}.
 
 \section{Conclusions}\label{sec:conclusion}
-Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
-ehi
-\#TODO
+% Conclusions are a superbrief summary of what has been done and highlighting of the "take home message"
+% ehi
+For this course project, the report shows how the demo was made, in particular what problem we encountered
+and what we exploited, showing what protocols could be unused in the future. We showed a possible simple data structure
+that can allow the controller to be able to know the network, its topology and which route to set, based on the traffic.
+Morover, we showed how to exploit the potentialities of the SDN network, giving particular importance on the
+statistics and the path switching capabilities.
+The multi-graph solution allowed us to connect the network and, based on its condition, choose which weight to apply
+in order to give a Quality of Service based network. With some boards that supports the newer version of Openflow
+(i.e. 1.5, the latest), the code can be ported, so it can be used with the Ryu core controller and more statistic can
+be exploited. Moreover, the controller can set more strictive rules for the flow control and can get more informations from the
+switches.
+
+% mi pare vada bene, l ultima versione di pox dovrebbe essere almeno la 1.5.1
 

### 2017-03-15 report/paper.tex
ultima aggiunta

@@
   The links on our network logically have many parameters in witch we can storage network's performance; in practice a NetworkX's graph support only 
-  one or two link's attributes so we have create a lots of graphs with the same topological information but different link's attributes.
+  one or two link's attributes so we have create a lots of graphs with the same topological information but different link's attributescd .
 
@@
 \\
+\\
+The following picture shows the visual output that we can obtain from delay's graph. In this graph we storage only the core information;
+\begin{figure}[!h]
+ \centering
+ \includegraphics[scale=0.26]{images/delay.png}
+ \caption{\emph{Delay's graph.}}
+ \label{fig:topo}
+\end{figure}
 

### 2017-03-15 report/paper.tex
corrections on english

@@
 Software Defined Networking (SDN) paradigm has emerged in response to traditional networking architecture's limitations.
-The main advantage of SDN is a centralized global network view. This particular feature pushed researchers to keep on improving QoS.
+The main advantage of SDN is a centralized global network view, which pushed researchers to keep on improving QoS exploiting it.
 
-In this report we will focus on two different aspects; all problems we faced during our attempt to create a based SDN application able to
+In this report we will focus on two different aspects, the problems we faced during our attempt to create a based SDN application able to
 support a QoS service and the results we achieved.
@@
 OpenFlow standard and python POX controller, which is largely spread, were used as instruments to make this experience possible,
-despite the limit of using a new type of tecnology still in the research phase.
+despite the limit of using a new type of technology still in the research phase.
 %corretto
@@
 SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brain” of the network.
-The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
-One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it isn’t the only SDN standard. 
-Centralized, programmable SDN environments can easily adjust to the rapidly changing business'needs. SDN can reduce costs and restrict
+The SDN Controller relays on the information given by switches and routers via southbound APIs, and acts with what the applications set via northbound APIs.
+One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it is not the only SDN implementation available. 
+Centralized, programmable SDN environments can easily adjust to the rapidly changing business needs and can reduce costs, restrict
 wasteful provisioning, as well as provide flexibility and innovation for networks.
@@
 
-Openflow defines a number of messages between a controller and different switches. Now there will be presented some messages we exploited\cite{openflow}:
+Openflow defines a number of messages between a controller and different switches; some of the ones we exploited will be shown right after:\cite{openflow}:
 \begin{itemize}
@@
 	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
-\item \emph{FeatureReq} \newline This is the message that the controller sends to a switch to get to know its abilities. The switch
+\item \emph{FeatureReq} \newline This is the message that the controller sends to a switch to get to know its capabilities. The switch
 	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
@@
 to some ports.
-\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch.This is the message that allows routing.
-\\
+\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch, allowing traffic to be routed along specific paths. The structure of this message is explained as follows.
 \\
@@
 \begin{itemize}
- \item \texttt{match fields:} fields that compare packets in order to look for correspondences. They contain the in_port, the packet's header and others optional fields;
- \item \texttt{priority:} this field is used when there are rows with more than one correspondence;
+ \item \texttt{match fields:} compare packets properties in order to look for correspondences. They contain the in_port, the packet's header and others optional fields;
+ \item \texttt{priority:} this field is used when the specific packet matches more than one rule. The higher the number, the higher the priority a rule has;
  \item \texttt{counter:} this field was improved every time a correspondence with a packet is matched;
@@
 \begin{itemize}
- \item \texttt{in_port}: the port of swith;
- \item \texttt{dl_src/dl_dst}: 802.3 MAC source and destination of packet;
- \item \texttt{dl_type}: 802.3 EtherType;
+ \item \texttt{in_port}: the port of swith the packet is arriving;
+ \item \texttt{dl_src/dl_dst}: IEEE 802 MAC source and destination of packet;
+ \item \texttt{dl_type}: IEEE 802 EtherType;
  \item \texttt{nw_tos}: IPv4 ToS;
@@
  \item \texttt{SetNWTos}: change ToS fields of message;
- \item \texttt{SetTPSrc/Dst}
+ \item \texttt{SetTPSrc/Dst}: set the TCP or UDP source and destination port.
 \end{itemize}
@@
 slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
-the network disposition, but doesn't know each host how much traffic is doing. For that the administrator need to put a firewall.
-The last one isn't needed, as the controller can set a switch to drop all the packets from/to a destination.
+the network topology, but doesn't know how much traffic each host is doing. For that the administrator need to put a firewall, which it's not needed in SDN networks,
+as the controller can set a switch to drop all the packets from/to a destination.
 

### 2017-03-15 report/paper.tex
other corrections on english. end to read after, now class

@@
 An adaptable Network has to real-time reroute traffic in order to ensure a lot of demands.
-A general efficient Network has to manage some tasks, for example avoiding links congestion, reroute traffic when some malfunction accure on a link
+A general efficient Network has to manage some tasks, for example avoiding links congestion, reroute traffic when some malfunction occure on a link
 and running a link path algorithm. 
@@
 
-To reach this gools we needed to have a global knowledge on the network and we had to modify the routing table on every routerboard with a
+To reach this goals we needed to have a global knowledge on the network and we had to modify the routing table on every routerboard with a
 global purpose.
@@
 The first thing we needed to accomplish was to make the controller aware of what switch and which port of the switch is connected to the external internet:
-to do so, we allowed our controller to be able to receive and process in-line parameters. We provided two parametes to the controller, so that
-it could recognize where the traffic to the internet need to be sent. Theese two parameters were the mac address of the
+to do so, we allowed our controller to be able to receive and process in-line parameters. We provided two parameters to the controller, so that
+it could recognize where the traffic to the internet need to be sent. These two parameters were the mac address of the
 port of a switch that is connected to the link and the network address of our network, to let the controller know if a request from a host was meant to another host
@@
 We developed a controller that can handle a mesh network with loops, avoiding
-packets flooding. The controller stores the switch and the connection between them and knows which hosts are connected to each switch port starting from the hypothesys that only one
+packets flooding. The controller stores the switch and the connection between them and knows which hosts are connected to each switch port starting from the hypotheses that only one
 host is connected to each port, unless the port is the one connected to the internet.
@@
   The links on our network logically have many parameters in witch we can storage network's performance; in practice a NetworkX's graph support only 
-  one or two link's attributes so we have create a lots of graphs with the same topological information but different link's attributescd .
+  one or two link's attributes so we have create a lots of graphs with the same topological information but different link's attributes .
 
@@
 We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
-\#Mb for the gaming traffic, \#Mb for the VoIP traffic and 3GB for the torrent traffic.
+\#Mb for the gaming traffic, 70 Mb for the VoIP traffic and 3GB for the torrent traffic.
 The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP).
@@
 specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
-\newline The idea is to measure the time that occure from sending a packet from controller to a switch and the time in which the packet return from
+\newline The idea is to measure the time that occur from sending a packet from controller to a switch and the time in which the packet return from
 an other switch.
@@
 \newline First we have to discovery round trip time from controller to every single switch on our topology.
-In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
-option and when the message return to the controller we stop the timer and calculare RTT for the switch.
+In order to do this on the controller we start a timer, we send a proper packet to the switch specifying the \texttt{OFPP_CONTROLLER - Send to the controller}
+option and when the message return to the controller we stop the timer and calculate RTT for the switch.
 \begin{figure}[!h]
@@
 \newline
-Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifing the output port.
+Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifying the output port.
 Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
-\newline This part required a particolar flow entry on the second switch in order to send to the controller only some particulare packet.
+\newline This part required a particular flow entry on the second switch in order to send to the controller only some particular packet.
 \begin{figure}[!h]
@@
 \\
-Teorethically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
-What we observe is a estimated delay like a few seconds which is not the case, infact if we made some \texttt{ping test} from one host to another host
+Theoretically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
+What we observe is a estimated delay like a few seconds which is not the case, in fact if we made some \texttt{ping test} from one host to another host
 in this SDN network we realize delay around 3ms.
@@
 	The first data that was collected was the actual link capacity, which depends both on the cable used to connect
-	different devices and the effective capacity for each ethernet port linked by the cable itself. Everytime a device supporting
+	different devices and the effective capacity for each Ethernet port linked by the cable itself. Every time a device supporting
 	openflow connects to the controller, it gives to the latter information about its status and its capabilities. One of them is
-	the link capacity in three ways: the first is the \textit{supported} capacity, which is the phisical maximum capacity, the second is the
+	the link capacity in three ways: the first is the \textit{supported} capacity, which is the physical maximum capacity, the second is the
 	\textit{advised} capacity, which name describes itself and the latter is the \textit{current} one. We choose to save the last one,
@@
 \subsubsection{Packet error rate}
-The port statistics given by the switches contains different informations about the state of each port in a switch. To classify a link for its
+The port statistics given by the switches contains different information about the state of each port in a switch. To classify a link for its
 reliability, we divided the number of packets with erroneous symbols that couldn't be corrected with a FEC by the total number of packets that were sent,
-getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, wherease the
+getting the packet error rate. The choice we made when exploiting this statistic was privileging the gaming and normal traffic, whereas the
 heavy traffic was directed to links with highest packet error rate.  As the switches exchanges a packet before recognizing most of the error, both because the
-phisical error are corrected using blocks coding and other error are found using a \textit{Frame Check Sequence} like CRC, the byte error were considered worse as
-those are indirect measures themself and the error derived from the calculations could grow up.
+physical error are corrected using blocks coding and other error are found using a \textit{Frame Check Sequence} like CRC, the byte error were considered worse as
+those are indirect measures themselves and the error derived from the calculations could grow up.
 
@@
 when the node is considered in heavy traffic. Those two threshold were made as they are average values and considering that our network is made of few really fast link,
-so on one side the lower limit is sufficient not to consider gaming traffic to be a normal traffic, wherease it's easy to reach the high level of traffic load between two machines.
+so on one side the lower limit is sufficient not to consider gaming traffic to be a normal traffic, whereas it's easy to reach the high level of traffic load between two machines.
 The changing of the path were made only if the last change happened before one minute because the torrent traffic has, in its protocol, a discovery
-algortithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
+algorithm. This means that if a lot of changes were made, most of the traffic generated by the peer to peer, is the control one and the result is that the
 load in the network will be heavy and last longer.
@@
 If it has the rule to forward the packet, it will act as declared in the most specific matching rule. Otherwise it will
-send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host whithin
+send to the controller a PacketIn and the controller will act as programmed. If an host want to communicate to another host within
 the network, it sends the packet to the line and then the switch will check in its tables and execute the actions of the match found.
@@
 \subsection{Ip switching in pox}
-During our tests, both with Mininet and in the lab, we found out that the core of POX, when a PacketIn event occoures, switches the source and the destination addresses,
+During our tests, both with Mininet and in the lab, we found out that the core of POX, when a PacketIn event occures, switches the source and the destination addresses,
 both MAC and IP, if one of the hosts is already known. We weren't able to understand in which point of the code of POX  and why it happens.
@@
 parameters known by the controller, i.e. if the two host are connected to two different switches and the controller switches the source with the
-destination address, we can switch it back; the result is that the PacketIn event now has correct informations.
+destination address, we can switch it back; the result is that the PacketIn event now has correct information.
 If the two host are in the same switch, the problem is solved using a controlled packet flooding, described in section \ref{sec:connectivity}.
@@
 that can allow the controller to be able to know the network, its topology and which route to set, based on the traffic.
-Morover, we showed how to exploit the potentialities of the SDN network, giving particular importance on the
+Moreover, we showed how to exploit the potentialities of the SDN network, giving particular importance on the
 statistics and the path switching capabilities.
@@
 (i.e. 1.5, the latest), the code can be ported, so it can be used with the Ryu core controller and more statistic can
-be exploited. Moreover, the controller can set more strictive rules for the flow control and can get more informations from the
+be exploited. Moreover, the controller can set more strictive rules for the flow control and can get more information from the
 switches.

### 2017-03-15 report/paper.tex
final correction. matteo need to add amount of gaming traffic he analyzed

@@
  in order to modify the pipeline flux;
- \item \texttt{timeout:} this is the maximum time afther which the rule expire;
- \item \texttt{flag:} with this field we can differenciate similar voices in order to managed multiple table with similar rules;
+ \item \texttt{timeout:} this is the maximum time after which the rule expire;
+ \item \texttt{flag:} with this field we can differentiate similar voices in order to managed multiple table with similar rules;
  \\
@@
 \subsection{Known Limitations}
-SDN networks are known to be underperforming with low traffic, which means that if nodes tries to connect with each other
+SDN networks are known to be under performing with low traffic, which means that if nodes tries to connect with each other
 to exchange little traffic, the network will have a higher delay and will perform worst than a
-traditional network. We tried to reduce this underperformance with the usage of the wildcards:
+traditional network. We tried to reduce this under performance with the usage of the wildcards:
 Every rule needs at least the source ip or the destination ip. If one of them is in a network
@@
 
-Another limitation is the no default routing from a source to a destination: when an host
+Another limitation is the absence of the default routing from a source to a destination: when an host
 is plugged to a network of SDN switches, it won't know the destination. The only way it can
@@
 An adaptable Network has to real-time reroute traffic in order to ensure a lot of demands.
-A general efficient Network has to manage some tasks, for example avoiding links congestion, reroute traffic when some malfunction occure on a link
+A general efficient Network has to manage some tasks, for example avoiding links congestion, reroute traffic when some malfunction occur on a link
 and running a link path algorithm. 
@@
 
-\subsection{Exterior comunication}
+\subsection{Exterior communication}
 Our network needed to be able to communicate with the external internet to be useful for a demo which demonstrate the capabilities of a SDN network.
@@
 
-We have use 4 MikroTik Routerboards that were provided us by the Communication's Network Lab in the department of information  engineering to 
+We used 4 MikroTik Routerboards that were provided us by the Communication's Network Lab in the Department of Information Engineering to 
 University of Padua.
-The 4 routerboards are connected each other in order to design a topological loop. Then every sigle board connects two hosts and is connected with a 
+The 4 routerboards are connected each other in order to design a topological loop. Then every single board connects two hosts and is connected with a 
 legacy switch which provides logically connection with controller.
@@
 \subsection{Graph and Network's structures}
-  In order to achieve the realtime adaptability and statistics storage we developed ad-hoc structures.
+  In order to achieve the real time adaptability and statistics storage we developed ad-hoc structures.
   We used a specific Python's library called NetworkX \cite{networkx}. NetworkX allow us to create a graph in witch we can storage suitable nodes
@@
 \newline
-If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
+If we need to calculate the delay on a specific link, the rtt has to be calculated with the headed switch of this link.
 \\

### 2017-03-15 report/paper.tex
final correction. everything should be fine
images sistemate manca da unire alle correzioni di andrea faccio dopo

--- 2017-03-15 report/paper.tex
@@
 Software Defined Networking (SDN) paradigm has emerged in response to traditional networking architecture's limitations.
-The main advantage of SDN is a centralized global network view, which pushed researchers to keep on improving QoS exploiting it.
+The main advantage of SDN is a centralized global network view. This particular feature pushed researchers to keep on improving QoS.
 
-In this report we will focus on two different aspects, the problems we faced during our attempt to create a based SDN application able to
+In this report we will focus on two different aspects; all problems we faced during our attempt to create a based SDN application able to
 support a QoS service and the results we achieved.
@@
 OpenFlow standard and python POX controller, which is largely spread, were used as instruments to make this experience possible,
-despite the limit of using a new type of technology still in the research phase.
+despite the limit of using a new type of tecnology still in the research phase.
 %corretto
@@
 SDN offers a centralized view of the network, giving an SDN Controller the ability to act as the “brain” of the network.
-The SDN Controller relays on the information given by switches and routers via southbound APIs, and acts with what the applications set via northbound APIs.
-One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it is not the only SDN implementation available. 
-Centralized, programmable SDN environments can easily adjust to the rapidly changing business needs and can reduce costs, restrict
+The SDN Controller relays information to switches and routers via southbound APIs, and to the applications with northbound APIs.
+One of the most well-known protocols used by SDN Controllers is OpenFlow, widely accepted even if it isn’t the only SDN standard. 
+Centralized, programmable SDN environments can easily adjust to the rapidly changing business'needs. SDN can reduce costs and restrict
 wasteful provisioning, as well as provide flexibility and innovation for networks.
@@
   %todo spostare l'immagine correttamente
-\begin{figure}[!h]
+\begin{center}
  \centering
  \includegraphics[scale=0.58]{images/of.jpg}
- \caption{\emph{OpenFlow}}
- \label{fig:topo}
-\end{figure}
-  \\
+ {Figure 1: \emph{OpenFlow}}
+ %\label{fig:topo}
+\end{center}
 \textbf{Openflow 1.0 messages}
 
-Openflow defines a number of messages between a controller and different switches; some of the ones we exploited will be shown right after:\cite{openflow}:
+Openflow defines a number of messages between a controller and different switches. Now there will be presented some messages we exploited\cite{openflow}:
 \begin{itemize}
@@
 	is turned on. If the switch sent the message, the controller, after receiving it, will send the FeatureReq packet to that switch.
-\item \emph{FeatureReq} \newline This is the message that the controller sends to a switch to get to know its capabilities. The switch
+\item \emph{FeatureReq} \newline This is the message that the controller sends to a switch to get to know its abilities. The switch
 	respond with a FeatureRes and when this last message reaches the pox controller, the ConnectionUp event is raised.
@@
 to some ports.
-\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch, allowing traffic to be routed along specific paths. The structure of this message is explained as follows.
+\item \emph{FlowMod} \newline This message set, delete or modify a rule inside a switch.This is the message that allows routing.
+\\
 \\
@@
 \begin{itemize}
- \item \texttt{match fields:} compare packets properties in order to look for correspondences. They contain the in_port, the packet's header and others optional fields;
- \item \texttt{priority:} this field is used when the specific packet matches more than one rule. The higher the number, the higher the priority a rule has;
+ \item \texttt{match fields:} fields that compare packets in order to look for correspondences. They contain the in_port, the packet's header and others optional fields;
+ \item \texttt{priority:} this field is used when there are rows with more than one correspondence;
  \item \texttt{counter:} this field was improved every time a correspondence with a packet is matched;
@@
 \begin{itemize}
- \item \texttt{in_port}: the port of swith the packet is arriving;
- \item \texttt{dl_src/dl_dst}: IEEE 802 MAC source and destination of packet;
- \item \texttt{dl_type}: IEEE 802 EtherType;
+ \item \texttt{in_port}: the port of swith;
+ \item \texttt{dl_src/dl_dst}: 802.3 MAC source and destination of packet;
+ \item \texttt{dl_type}: 802.3 EtherType;
  \item \texttt{nw_tos}: IPv4 ToS;
@@
  \item \texttt{SetNWTos}: change ToS fields of message;
- \item \texttt{SetTPSrc/Dst}: set the TCP or UDP source and destination port.
+ \item \texttt{SetTPSrc/Dst}
 \end{itemize}
@@
 slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
-the network topology, but doesn't know how much traffic each host is doing. For that the administrator need to put a firewall, which it's not needed in SDN networks,
-as the controller can set a switch to drop all the packets from/to a destination.
+the network disposition, but doesn't know each host how much traffic is doing. For that the administrator need to put a firewall.
+The last one isn't needed, as the controller can set a switch to drop all the packets from/to a destination.
 
@@
   \includegraphics[scale=0.30]{images/rboard.png}
-  \caption{\emph{Experience scenario}}
+
   \label{fig:topo}
 \end{figure}
+\begin{center}
+ {Figure 2: \emph{Experience scenario}}
+\end{center}
 
@@
  \includegraphics[scale=0.70]{images/topo.png}
- \caption{\emph{Experience scenario}}
+ {Figure 3: \emph{Experience scenario}}
  \label{fig:topo}
@@
  \includegraphics[scale=0.30]{images/topology.png}
- \caption{\emph{Experience scenario}}
+ {Figure 4: \emph{Experience scenario}}
  \label{fig:topo}
@@
 option and when the message return to the controller we stop the timer and calculare RTT for the switch.
-\begin{figure}[!h]
+\begin{center}
  \centering
  \includegraphics[scale=0.70]{images/rtt2.png}
- \caption{\emph{RTT between controller and OF switch}}
+
  \label{fig:topo}
-\end{figure}
-\newline
+\end{center}
+
+\begin{center}
+ {Figure 5: \emph{RTT between controller and OF switch}}
+\end{center}
+
 If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
@@
  \includegraphics[scale=0.70]{images/rtt1.png}
- \caption{\emph{Pck from controller to itself following a particular way.}}
  \label{fig:topo}
 \end{figure}
-\\
-\newline At this point we can calculate the delay of the link.
+
+\begin{center}
+  {Figure 6:\emph{Pck from controller to itself following a particular way.}}
+\end{center}
+
+At this point we can calculate the delay of the link.
 \\
@@
 The following picture shows the visual output that we can obtain from delay's graph. In this graph we storage only the core information;
-\begin{figure}[!h]
+\begin{center}
  \centering
  \includegraphics[scale=0.26]{images/delay.png}
- \caption{\emph{Delay's graph.}}
+ {Figure 7:\emph{ Delay's graph;}}
  \label{fig:topo}
-\end{figure}
+\end{center}
 
+%\end{figure}
 \subsubsection{Link Capacity}

### 2017-03-15 report/paper.tex

--- 2017-03-15 report/paper.tex
@@
   %todo spostare l'immagine correttamente
-\begin{figure}[!h]
+\begin{center}
  \centering
  \includegraphics[scale=0.58]{images/of.jpg}
- \caption{\emph{OpenFlow}}
- \label{fig:topo}
-\end{figure}
-  \\
+ {Figure 1: \emph{OpenFlow.}}
+ %\label{fig:topo}
+\end{center}
 \textbf{Openflow 1.0 messages}
@@
   \includegraphics[scale=0.30]{images/rboard.png}
-  \caption{\emph{Experience scenario}}
+
   \label{fig:topo}
 \end{figure}
+\begin{center}
+ {Figure 2: \emph{Experience scenario.}}
+\end{center}
 
@@
  \includegraphics[scale=0.70]{images/topo.png}
- \caption{\emph{Experience scenario}}
+ {Figure 3: \emph{Experience scenario.}}
  \label{fig:topo}
@@
  \includegraphics[scale=0.30]{images/topology.png}
- \caption{\emph{Experience scenario}}
+ {Figure 4: \emph{Experience scenario.}}
  \label{fig:topo}
@@
 We did 3 type of traffic analysis: Gaming traffic, Skype® and torrent traffic. We analyzed
-\#Mb for the gaming traffic, 70 Mb for the VoIP traffic and 3GB for the torrent traffic.
+300 Mb for the gaming traffic, 70 Mb for the VoIP traffic and 3GB for the torrent traffic.
 The analysis involved the average number of packets per second, the packet size, the protocol (TCP or UDP).
@@
 \newline First we have to discovery round trip time from controller to every single switch on our topology.
-In order to do this on the controller we start a timer, we send a proper packet to the switch specifying the \texttt{OFPP_CONTROLLER - Send to the controller}
-option and when the message return to the controller we stop the timer and calculate RTT for the switch.
-\begin{figure}[!h]
+
+In order to do this on the controller we start a timer, we send a proper packet to the switch specifing the \texttt{OFPP_CONTROLLER - Send to the controller}
+option and when the message return to the controller we stop the timer and calculare RTT for the switch.
+\begin{center}
  \centering
  \includegraphics[scale=0.70]{images/rtt2.png}
- \caption{\emph{RTT between controller and OF switch}}
+
  \label{fig:topo}
-\end{figure}
-\newline
-If we need to calculate the delay on a specific link, the rtt has to be calculated with the headed switch of this link.
-\\
-\newline
+
+\end{center}
+
+\begin{center}
+ {Figure 5: \emph{RTT between controller and OF switch.}}
+\end{center}
+
+If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
+
 Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifying the output port.
@@
  \includegraphics[scale=0.70]{images/rtt1.png}
- \caption{\emph{Pck from controller to itself following a particular way.}}
  \label{fig:topo}
 \end{figure}
-\\
-\newline At this point we can calculate the delay of the link.
+
+\begin{center}
+  {Figure 6: \emph{Pck from controller to itself following a particular way.}}
+\end{center}
+
+At this point we can calculate the delay of the link.
 \\
 \\
+\begin{center}
   delay = (TM2 - TM1) - RTT1/2 - RTT2/2
-\\
-\\
+\end{center}
+
 Theoretically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
@@
 The following picture shows the visual output that we can obtain from delay's graph. In this graph we storage only the core information;
-\begin{figure}[!h]
+\begin{center}
  \centering
  \includegraphics[scale=0.26]{images/delay.png}
- \caption{\emph{Delay's graph.}}
+ {Figure 7:\emph{ Delay's graph.}}
  \label{fig:topo}
-\end{figure}
+\end{center}
 
+%\end{figure}
 \subsubsection{Link Capacity}
